### 2025-04-03 codebook_data_simulated.Rmd
synthetic data

---
title: "Codebook created on 2025-04-03 at 2025-04-03 10:38:35"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis')
```

A codebook contains documentation and metadata describing the contents, structure, and layout of a data file.

## Dataset description
The data contains 1000 cases and 27 variables.

## Codebook

```{r}
summaries <- read.csv("codebook_data_simulated.csv", stringsAsFactors = FALSE)
summaries <- summaries[, !colSums(is.na(summaries)) == nrow(summaries)]
options(knitr.kable.NA = '')
knitr::kable(summaries, row.names = FALSE, digits = 2)
```

### Legend

* __Name__: Variable name
* __type__: Data type of the variable
* __missing__: Proportion of missing values for this variable
* __unique__: Number of unique values
* __mean__: Mean value
* __median__: Median value
* __mode__: Most common value (for categorical variables, this shows the frequency of the most common category)
* **mode_value**: For categorical variables, the value of the most common category
* __sd__: Standard deviation (measure of dispersion for numerical variables
* __v__: Agresti's V (measure of dispersion for categorical variables)
* __min__: Minimum value
* __max__: Maximum value
* __range__: Range between minimum and maximum value
* __skew__: Skewness of the variable
* __skew_2se__: Skewness of the variable divided by 2*SE of the skewness. If this is greater than abs(1), skewness is significant
* __kurt__: Kurtosis (peakedness) of the variable
* __kurt_2se__: Kurtosis of the variable divided by 2*SE of the kurtosis. If this is greater than abs(1), kurtosis is significant.

This codebook was generated using the [Workflow for Open Reproducible Code in Science (WORCS)](https://osf.io/zcvbs/)

### 2025-04-03 prepare_data.R

@@ -4,7 +4,16 @@
 # to store the data.
 
 library(worcs)
-dat_simulated <- read.csv("simulated_data.csv")
+library(here)
+library(tidyverse)
+data_simulated <- read.csv(here("data/simulated_data.csv"))[,-1]
+
+#####################################################################################################################################
+# Read real data in 
+# file <- here("data/Tulokset - xxxxx.xlsx")
+# data <- read_xlsx(file,n_max = XX)
+# summary(data)
+#####################################################################################################################################
 
 f1 <- function (x){
   #as numbers but labelled with words
@@ -17,7 +26,21 @@ f1 <- function (x){
   x
 }
 
-dat_simulated <- dat_simulated %>% 
+#Make question abbreviations
+qabbr <- c(
+  str_c("po0",1:3),
+  str_c("si0",1:3),
+  str_c("it0",1:3),
+  str_c("ie0",1),
+  str_c("ac0",1:4),
+  str_c("ai0",1:4),
+  str_c("iv0",1:3),
+  str_c("bd0",1:4),
+  str_c("de0",1:2)
+)
+
+
+data_simulated <- data_simulated %>% 
   mutate_at(
     vars(
       qabbr[1:21]),
@@ -28,4 +51,27 @@ dat_simulated <- dat_simulated %>%
          de01 = factor(de01)
   )  
 
-open_data(dat_simulated)
+#####################################################################################################################################
+original_name <- here(paste0("data/",deparse(substitute(data_simulated)), ".csv")) # when real data collected this will be changed to it
+#####################################################################################################################################
+
+# Above is your data cleaning script
+# it ends with closed_data(), which for now just saves your synthetic data locally.
+# When you have real data, it will save the real data locally.
+closed_data(
+  #####################################################################################################################################
+  data = data_simulated, # when real data collected this will be changed to it
+  #####################################################################################################################################
+  filename = original_name, 
+  synthetic = FALSE)
+# Then, add your custom synthetic data to this data asset (yes, for now they are both the same, but
+# when you collect the real data, the local data will be real and the remote synthetic
+synthetic_name <- paste0("synthetic_",deparse(substitute(data_simulated)), ".csv")
+#####################################################################################################################################
+original_name <- paste0("data/",deparse(substitute(data_simulated)), ".csv") # when real data collected this will be changed to it
+#####################################################################################################################################
+add_synthetic(
+  data = data_simulated,
+  synthetic_name = synthetic_name,
+  original_name = original_name,
+)

### 2025-04-04 preregistration_code.Rmd
preregistered analysis

---
title: "Supplementary file 1: Preregistered analysis"
author: "Abigail Edwards and Mikko Arvas"
date: "`r Sys.time()`"
output:
  html_document:
    toc: yes
    theme: united
  pdf_document:
    toc: yes
---

TODO:
 -remove superfluous outputs
 -check discussion with Lauri, was there anything else to do?
 -check model saving
 -which format to manuscript file
 -fill out the preregistration form

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(worcs)
library(tidyverse)
library(gtsummary)
library(likert)
library(here)
library(corrplot)
library(ggstats)
library(lavaan)
library(lavaanPlot)
#library(semPlot) # update plots to semPlot later
#library(blavaan) # could not get it to work with this model and synthetic data
#library(tidySEM) # https://github.com/yrosseel/lavaan/issues/359 lavaan does not work if tidySEM is loaded.
```

# Data

```{r}
load_data()
if (exists("data_simulated")) {
  data <- data_simulated
  rm(data_simulated)
}
# sometimes it get strange errors about C stack which can be solved with this
#data <- as.data.frame(data)
fitmodels <- TRUE
if (!fitmodels) {
  print("Loading previously fitted models")
  load(file=here(str_replace(str_c("results/",datafile),"\\.data\\.",".models.")))
  fit.po02ss <- models$fit.po02ss
  fit.po02 <- models$fit.po02
  fit.po03ss <- models$fit.po03ss
  fit.po03 <- models$fit.po03
}

```

Items
<table>
<tr>
<td></td>
<td> Psychological ownership</td>
</tr>
<tr>
<td>po01</td>
<td>
1. I feel a very high degree of personal ownership of donating blood.
</td>
</tr>
<tr>
<td>po02</td>
<td>
2. I feel like donating blood belongs to me.
</td> </tr>
<tr>
<td>po03</td> <td>
3. I feel like I own donating blood.
</td> </tr>
<tr> <td></td>

<td>
Self-identity
</td> </tr>
<tr>
<td>si01</td>
<td>
1. Donating blood is important to me.
</td> <tr>
<td>si02</td> <td>
2. I am like the kind of person who donates blood.
</td> </tr>
<tr> <td>si03</td> 
<td>
3. Donating blood is an important part of who I am.
</td> </tr>
<tr> <td></td>

<td>
Intention
</td></tr>
<tr><td>it01</td>
<td>
1. I would like to donate blood in the future.
</td>   </tr>
<tr><td>it02</td>
<td>
2. I intend to donate blood in the future.
</td> </tr>
<tr><td>it03</td>

<td>
3. I intend to make more than just a one-off blood donation.
</td></tr>
<tr><td></td>
<td>
Intention to donate for emergency
</td></tr>
<tr><td>ie01</td>
<td>
1. I would be willing to donate blood if I received an urgent callout from the Finnish Red Cross Blood Service.
</td></tr>
<tr><td></td> 
<td>
Antecedents of psychological ownership
</td></tr>
<tr><td></td>

<td>
Control
</td></tr>
<tr><td>ac01</td>
<td>
1. I have influence over the things that affect me while donating blood.
</td></tr>
<tr><td>ac02</td>
<td>
2. I control the location and scheduling for donating blood.
</td></tr>
<tr><td>ac03</td>
<td>
3. I influence decisions related to my blood donation.
</td></tr>
<tr><td>ac04</td>
<td>
4. In general, I have control over donating blood.
</td></tr>
<tr><td></td>
<td>
Intimate knowledge
</td></tr>
<tr><td>ai01</td>

<td>
1. I am intimately familiar with what is going on with regard to donating blood.
</td></tr>
<tr><td>ai02</td>
<td>
2. I have a depth of knowledge as it relates to donating blood.
</td></tr>
<tr><td>ai03</td>
<td>
3. I have a comprehensive understanding of donating blood.
</td></tr>
<tr><td>ai04</td>
<td>
4. I have a broad understanding of donating blood.
</td></tr>
<tr><td></td>

<td>
Self-Investment
</td></tr>
<tr><td>iv01</td>
<td>
1.	I have invested a major part of myself into donating blood.
</td></tr>
<tr><td>iv02</td>
<td>
2. I have invested a significant amount of my time into donating blood.
</td></tr>
<tr><td>iv03</td>
<td>
3. Overall, I invested a lot into donating blood.
</td></tr>
<tr><td></td>

<td>
Blood donation history
</td></tr>
<tr><td>bd01</td>
<td>
1. Do you believe that you are currently eligible to donate blood [yes, no, unsure]
</td></tr>
<tr><td>bd02</td>
<td>
2. How many times in total have you donated blood at the Finnish Red Cross Blood Service Finland? [select number]
</td></tr>
<tr><td>bd03</td>
<td>
3. How many times in the last two years have you donated at the Finnish Red Cross Blood Service? [select number]
</td></tr>
<tr><td>bd04</td>
<td>
4. When did you last attend a Finnish Red Cross blood donation centre with the intention of donating blood? [select date]
</td></tr>
<tr><td></td>

<td>
Demographics
</td></tr>
<tr><td>de01</td>
<td>1. What is your gender? [man, woman, non-binary, prefer to not disclose, prefer to self-describe (open response)]
</td></tr>
<tr><td>de02</td>
<td>
2. How old are you today? [select number]
</td></tr>
</table>

```{r}
summary(data)
```

# Plot

## Demographics Table 1

```{r}

table1 <-
  tbl_summary(
    data %>% 
      rename(
      Gender =  de01,
      Age = de02,
      DonationEligibility = bd01,
      DonationCount = bd02,
      DonationCountLast2years = bd03,
    ),
    include = c(
      Age,
      DonationEligibility,
      DonationCount,
      DonationCountLast2years
      ),
    by =  Gender

  ) %>%
  add_n()  # add column with total number of non-missing observations
table1
```

## Age by Gender

```{r}

p <- ggplot(data)
p <- p +  geom_histogram(aes(x=de02),binwidth = 5)
p <- p + facet_wrap(. ~ de01)
p <- p + xlab("Age in years")
p

```

## Psychological ownership

```{r}
gglikert(
  data, 
  include =   starts_with("po") 
)
```

## Self-identity

```{r}
gglikert(
  data, 
  include =   starts_with("si") 
)
```

## Intention

```{r}
gglikert(
  data,
  include = c("it01","it02","it03","ie01")
)
```

## Antecedents of psychological ownership - Control

```{r}
gglikert(
  data, 
  include =   starts_with("ac") 
)
```

## Antecedents of psychological ownership - Intimate knowledge

```{r}
gglikert(
  data, 
  include =   starts_with("ai") 
)
```

## Antecedents of psychological ownership - Self-Investment

```{r}
gglikert(
  data, 
  include =   starts_with("iv") 
)
```

# Model

## Confirmatory factor analysis

```{r}
models <- list()

isordered <- data %>%  summarise(across(everything(), ~ is.ordered(.x))) %>%  as.logical() %>% which()
isfactor <- data %>%  summarise(across(everything(), ~ is.factor(.x))) %>%  as.logical() %>% which()

#lavaan does not understand empty factor levels
data.fit <- data %>%
  #lavaan cannot deal with unordered factor with more then 2 levels
  filter(de01 == 'man' | de01 == "women") %>% # exclusion
  filter(bd01 != 'no') %>% # exclusion
  mutate(
    across(isfactor, ~ fct_drop(.x)
    ) %>% 
      mutate(
        bd04 = as.numeric(bd04), # lavaan cannot handle dates
        bd04 = bd04 / 10^-(2 - str_length(as.numeric(max(bd04)))) # https://groups.google.com/g/lavaan/c/r7w-4HHg5R0 
        #"Your underweight variance seems to be much larger than other variances.  Try dividing it by 10 or 100 to make the SDs more similar across modeled variables." 
        # But this does not seem to necessarily help.
      )
  ) %>% 
  #Original paper used sum scores instead of CFA. Let's add them for comparison
  mutate(
    ss_psycown = as.numeric(po01) + as.numeric(po02) + as.numeric(po03),
    ss_selfid = as.numeric(si01) + as.numeric(si02) + as.numeric(si03),
    ss_intention = as.numeric(it01) + as.numeric(it02) + as.numeric(it03),
    #donation history is on diffent scales, normalise them before summing.
    #others should be positively correlated, but bd04 negatively
    ss_blooddonor = scale(as.numeric(bd02)) 
    + scale(as.numeric(bd01)) 
    + scale(bd03)  
    -scale(bd04) # do we want to include the DATE OF LAST DONATION?

  )

# To make bootstrapping run there cannot be any unordered factors
data.fit2 <- data.fit %>% 
  mutate(
    de01 = as.ordered(de01)
  ) %>% # and neither any factors at all
mutate(
  across(everything(),as.numeric)
)  # but this also good for plotting correlations

summary(data.fit)
```

```{r}
corrplot(cor(data.fit2))
```

```{r}
po01.model <- ' po =~ po01 + po02 + po03
                si =~ si01 + si02 + si03
                it =~ it01 + it02 + it03
                ac =~ ac01 + ac02 + ac03 + ac04
                ai =~ ai01 + ai02 + ai03 + ai04 
                iv =~ iv01 + iv02 + iv03
                bd =~ bd02 + bd03 + bd04 #bd02 is sure correlate positively

'

fit.po01.cfa <- cfa(po01.model,  
                    estimator = "WLSMV", #

                    data = data.fit)
fit.po01.cfa
```

```{r}
lavaan::summary(fit.po01.cfa,standardized = TRUE) # tidySEM is loaded this fails!
```

```{r}
cfastd <- lavaan::standardizedsolution(fit.po01.cfa)
cfastd %>% 
  #"est.std" (or Std.all in above summary output) is "factor loading" and crude rule if thumb is that abs(Std.all) should be above 0.3
  filter(abs(est.std) < 0.3) %>% 
  select(-z,-se) %>% 
  arrange(abs(est.std)) %>% 
  # Just the loadings
  filter(op == '=~')
# with synthetic data bd04 should come up
# as they are just random as can be seen from the corplot
```

```{r}
lavaan::fitmeasures(fit.po01.cfa,c("npar","chisq","df","cfi","rmsea","srmr"))
```

https://stats.stackexchange.com/questions/541707/cfa-in-lavaan-wont-converge 
"For the comparative fit indices CFI and TLI, you need at least >.900. For the absolute fit indices RMSEA and SRMR, you should have <.080."
So ones from simulated data look good enough.

```{r , fig.height=20}
lavaanPlot(
  fit.po01.cfa,
   edge_options = list(color = "grey"),
  coefs = TRUE,
  graph_options = list(rankdir = "LR")
           )
```

```{r}
#Make a model with out prespesified correlation structure
fit.po01.cfa.o <- cfa(po01.model,data.fit,orthogonal=TRUE)
round(
  sapply(list(cfa=fit.po01.cfa,cfa.o=fit.po01.cfa.o),
         function(x) fitmeasures(x,c("npar","chisq","df","cfi","rmsea","srmr"))
         )
,3)

#cfa should be better than cfa.o in all measures. cfi and chisq should be bigger, rmsea and srmr smaller in model cfa
```

## SEM

### Sum score model

```{r}
#https://stats.stackexchange.com/questions/340857/serial-mediation-in-r-how-to-setup-the-model

# N.B. you need to check that first item is positively correlated with your latent factor
# Otherwise you get opposite direction

po02ss.model <- ' 
#mediation
ss_psycown ~ a1 *  bd02 
ss_selfid ~ a2 * bd02 + d21 * ss_psycown

#regression
ss_intention ~  de01 + de02 + ie01 +  cp * bd02 + b1 * ss_psycown + b2 * ss_selfid

ind_eff := a1* d21 * b2 
'

cat(po02ss.model)

```

```{r}
if (fitmodels) {
fit.po02ss <- sem(po02ss.model,  
                    estimator = "MLR", # robust diagonally weighted least squares stable and should work with ordered factors
                                 data = data.fit2,
                auto.fix.first = FALSE, #  If TRUE, the factor loading of the first indicator is set to 1.0 for every latent variable. 
                std.lv = TRUE, 
                  )

models[["fit.po02ss"]] <- fit.po02ss
}

lavaan::summary(fit.po02ss,standardized = TRUE)

```

```{r}
varTable(fit.po02ss)
```

```{r}
lavaan::inspect(fit.po02ss,"r2")
```

```{r}
lavaan::standardizedsolution(fit.po02ss) %>% select(-label)
# Use standardizedSolution to obtain SEs and test statistics for standardized estimates.

```

```{r}
lavaan::standardizedsolution(fit.po02ss) %>% select(-label,se,z) %>%  filter(op == '~')
```

```{r}

fit.po02ss.est  <-standardizedSolution(fit.po02ss)
fit.po02ss.est %>% 
  filter(abs(est.std) < 0.3) %>% 
  select(-z,-se) %>% 
  arrange(abs(est.std)) %>% 
  # Just the loadings
  filter(op != '~1')

```

```{r}

labelsforplottingnull <- c(
  "ss_psycown" = "Psychological ownership",
  "ss_selfid"  = "Self-identity",
  "ss_intention" = "Intention",
  "bd02" = "Blood donation history"
)

lavaanPlot(
  model=fit.po02ss,  
  coefs = TRUE,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  labels = labelsforplottingnull,

  #In simulated data there should be correlations inside
  # bloodonor, psycown, selfid
  # and between
  # bloodonor, psycown
  # and nothing else
  sig = 0.001
  # if sig is < 0.001 one gets significant correlations
  # also from simulated data elsewhere suggesting that 0.05 is not enough

  )
```

```{r}
# Same with standardised coefficients
lavaanPlot(
  model = fit.po02ss, 
   labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  )
```

```{r}
#semPlot::semPaths(fit.po02ss)
```

### SEM

```{r}

po02.model <- ' 
#measurement model
psycown =~ po01 + po02 + po03
selfid =~ si01 + si02 + si03
intention =~ it01 + it02 + it03
blooddonor =~ bd02 + bd03 +bd04 
#bd01 is just used for exclusion of "no"

#mediation
psycown ~ a1 *  blooddonor
selfid ~ a2 * blooddonor + d21 * psycown

#regression
intention ~  de01 + de02 + ie01 +  cp * blooddonor + b1 * psycown + b2 * selfid
ind_eff := a1* d21 * b2
'

cat(po02.model)

```

```{r}

if (fitmodels) {
fit.po02 <- sem(po02.model,  
                    estimator = "MLR", # robust diagonally weighted least squares stable and should work with ordered factors
                                 data = data.fit2,
                auto.fix.first = FALSE,
                std.lv = TRUE 
                  )
models[["fit.po02"]] <- fit.po02
}
lavaan::summary(fit.po02,standardized = TRUE)

```

```{r}
varTable(fit.po02)
```

```{r}
lavaan::inspect(fit.po02,"r2")
```

```{r}
lavaan::standardizedsolution(fit.po02) %>% select(-label)
# Use standardizedSolution to obtain SEs and test statistics for standardized estimates.

```

```{r}
lavaan::standardizedsolution(fit.po02) %>% select(-label,se,z) %>%  filter(op == '~')
```

```{r}

fit.po02.est  <-standardizedSolution(fit.po02)
fit.po02.est %>% 
  filter(abs(est.std) < 0.3) %>% 
  select(-z,-se) %>% 
  arrange(abs(est.std)) %>% 
  # Just the loadings
  filter(op != '~1')

```

```{r}

labelsforplotting <- c(
  "psycown" = "Psychological ownership",
  "selfid"  = "Self-identity",
  "intention" = "Intention",
  "blooddonor" = "Blood donation history"
)

lavaanPlot(
  model=fit.po02,  
  coefs = TRUE,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  labels = labelsforplotting,

  #In simulated data there should be correlations inside
  # bloodonor, psycown, selfid
  # and between
  # bloodonor, psycown
  # and nothing else
  sig = 0.001
  # if sig is < 0.001 one gets significant correlations
  # also from simulated data elsewhere suggesting that 0.05 is not enough

  )
```

```{r}
# Same with standardised coefficients
lavaanPlot(
  model = fit.po02, 
   labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  )
```

### Sum score with bootstrapping

```{r}
nBoots <- 1000

if (fitmodels) {
#fit

# https://groups.google.com/g/lavaan/c/A6KCjXAZl-Q good example code

fit.po03ss <- sem(po02ss.model,  
                    estimator = "ML", # bootstrapping does not work with estimators designed for factors
                    data = data.fit2, # data that is completely numeric
                bootstrap = nBoots,
                se = "boot",
                auto.fix.first = FALSE,
                std.lv = TRUE
                )
models[['fit.po03ss']] <- fit.po03ss
}

lavaan::summary(fit.po03ss,standardized = TRUE)

```

```{r}
lavaanPlot(
  model = fit.po03ss, 
   labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  )

```

```{r}
fit.po03ss.best <- parameterEstimates(fit.po03ss)
fit.po03ss.best %>% select(-se,-z) %>% filter(op == "~")
```

### SEM with bootstrapping

```{r}

if (fitmodels) {
#fit

# https://groups.google.com/g/lavaan/c/A6KCjXAZl-Q good example code

fit.po03 <- sem(po02.model,  
                    estimator = "ML", # bootstrapping does not work with estimators designed for factors
                    data = data.fit2, # data that is completely numeric
                bootstrap = nBoots,
                se = "boot",
                auto.fix.first = FALSE,
                std.lv = TRUE 
                )
models[['fit.po03']] <- fit.po03
}

lavaan::summary(fit.po03,standardized = TRUE)

```

```{r}
lavaanPlot(
  model = fit.po03, 
   labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  )

```

```{r}
fit.po03.best <- parameterEstimates(fit.po03)
fit.po03.best %>% select(-se,-z) %>% filter(op == "~")
```

### SEM with Bayes

```{r}
po02b.model <- ' 
#measurement model
psycown =~ po01 + po02 + po03
selfid =~ si01 + si02 + si03
intention =~ it01 + it02 + it03
blooddonor =~ bd02 + bd03 +bd04

# #mediation
 psycown ~ a1 *  blooddonor
 selfid ~ a2 * blooddonor + d21 * psycown
# 
# #regression
 intention ~  de01 + de02 + ie01 +  cp * blooddonor + b1 * psycown + b2 * selfid
ind_eff := a1* d21 * b2
'

cat(po02.model)

```

```{r}
# This does not work!
# #blavaan
#tmp <- as.data.frame(data.fit2)
 # fit.po04 <- blavaan(po02.model,  
 #         #data = data.fit2,
 #         data = tmp,
 #         auto.fix.first = FALSE,
 #         std.lv = TRUE
 #)
#Error in validObject(.Object) : 
#invalid class “Fit” object: invalid object for slot "fx.group" in class "Fit": got class "NULL", should be or extend class "numeric"
#We will try to fix this with real data
```

```{r}

# #This works!
# data(HolzingerSwineford1939, package = "lavaan")
# # The Holzinger and Swineford (1939) example
# HS.model <- ' visual  =~ x1 + x2 + x3
#               textual =~ x4 + x5 + x6
#               speed   =~ x7 + x8 + x9 '
# 
# fit <- blavaan(HS.model, data = HolzingerSwineford1939,
#                auto.var = TRUE, auto.fix.first = TRUE,
#                auto.cov.lv.x = TRUE)
# summary(fit)
# coef(fit)

```

## Compare models

```{r}
data.forest  <- fit.po03.best %>% 
  mutate(estmet = "Factor analysis\nbootstrapped") %>%  
  filter(op == "~") %>% 
  bind_rows(
    fit.po02.est %>% mutate(estmet = "Factor analysis") %>% 
      rename(est=est.std) %>% 
      filter(op == "~") 
  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs)
    ) %>% 
  arrange(estmet, term)

data.forest  <- fit.po03ss.best %>% 
  mutate(estmet = "Sum score\nbootstrapped") %>%  
  filter(op == "~") %>% 
  bind_rows(
    fit.po02ss.est %>% mutate(estmet = "Sum score") %>% 
      rename(est=est.std) %>% 
      filter(op == "~") 
  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs)
    ) %>% 
  arrange(estmet, term) %>% 
  bind_rows(data.forest) %>% 
  mutate(
    lhs =str_replace_all(lhs,"ss_",""),
    rhs =str_replace_all(rhs,"ss_",""),
    term =str_replace_all(term,"ss_",""),
    estmet = as.factor(estmet),
    hollow_group=factor(if_else(ci.lower<=0 & ci.upper>=0, NA_character_, as.character(estmet)),
                             levels=levels(estmet))
  )

pos <- position_nudge(y=as.numeric(as.factor(data.forest$estmet))/5 -0.5)

data.forest %>% 
  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=estmet))+
  geom_vline(xintercept=0,color="grey") +
  geom_linerange(position=pos)+
  geom_point(shape=21,size=3,position=pos)+
  scale_fill_discrete(na.value=NA, guide ="none") + 
  labs(x="Standardised effect size",y="Relationship") +
  labs(color="Model") +
  geom_stripped_rows(color=NA) + theme_classic()

```

```{r}

round(
  sapply(list(SS=fit.po02ss,SSB=fit.po03ss,FA=fit.po02,FAB=fit.po03 ),
         function(x) fitmeasures(x,c("chisq","df","cfi","rmsea","srmr"))
         )
,3)

```

According to https://easystats.github.io/effectsize/reference/interpret_gfi.html

"For structural equation models (SEM), Kline (2015) suggests that at a minimum the following indices should be reported: The model chi-square, the RMSEA, the CFI and the SRMR."

chisq, Chi-squared, larger is better

cfi Comparative Fit Index, "It should be > .96" , higher is better

rmsea Root Mean Square Error of Approximation, "should be < .08", smaller is better

srmr Standardized Root Mean Square Residual, "Should be < .08", smaller is better

```{r}
if (fitmodels) {
  save(models,file=here(str_replace(str_c("results/",datafile),"\\.data\\.",".models.")))
}
```

### 2025-04-04 manuscript/manuscript.Rmd
manuscript

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["references.bib", "r-references.bib"]

csl               : vox-sanguinis.csl

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
knit              : worcs::cite_all
---

```{r setup, include = FALSE}
library("papaja")
library("worcs")
# We recommend that you prepare your raw data for analysis in 'prepare_data.R',
# and end that file with either open_data(yourdata), or closed_data(yourdata).
# Then, uncomment the line below to load the original or synthetic data
# (whichever is available), to allow anyone to reproduce your code:
# load_data()
r_refs("r-references.bib")
```

This manuscript uses the Workflow for Open Reproducible Code in Science [WORCS version 0.1.1, @vanlissaWORCSWorkflowOpen2021] to ensure reproducibility and transparency. All code <!--and data--> are available at <!--insert repository URL-->.

This is an example of a non-essential citation [@@vanlissaWORCSWorkflowOpen2021]. If you change the rendering function to `worcs::cite_essential`, it will be removed.

In this project we aim to replicate the analysis carried out with Australian blood donors [@edwardsPsychologicalOwnershipIdentity2023] with Finnish blood donors

<!--The function below inserts a notification if the manuscript is knit using synthetic data. Make sure to insert it after load_data().-->
`r notify_synthetic()`

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2025-04-07 preregistration_code.Rmd
indf_eff to forest

@@
  -check model saving
- -which format to manuscript file
  -fill out the preregistration form
@@
   mutate(estmet = "Factor analysis\nbootstrapped") %>%  
-  filter(op == "~") %>% 
+  filter(op == "~" | op == ":=") %>% 
   bind_rows(
@@
       rename(est=est.std) %>% 
-      filter(op == "~") 
+      filter(op == "~"| op == ":=") 
   ) %>% 
@@
   mutate(estmet = "Sum score\nbootstrapped") %>%  
-  filter(op == "~") %>% 
+  filter(op == "~"| op == ":=") %>% 
   bind_rows(
@@
     term =str_replace_all(term,"ss_",""),
+    term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
     estmet = as.factor(estmet),

### 2025-04-07 preregistration.Rmd
include code in prereg

---
title           : "Measuring psychological ownership and identity in a Finnish blood donor context"
shorttitle      : "Donor ownership"
date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"

author: 
  - name        : Abigail Edwards
    affiliation : 1
  - name        : Mikko Arvas
    affiliation : 2
  - name        : Lauri Hietajärvi
    affiliation : 3    
  - name        : Johanna Castrén
    affiliation : 2
  - name        : Barbara Masser
    affiliation : 2

affiliation:
  - id          : 1
    institution : The University of Queensland, Australia 
  - id          : 2
    institution : Finnish Red Cross Blood Service, Finland
  - id          : 3
    institution : University of Helsinki, Finland

header-includes:
  - \usepackage{pdfpages}

output: prereg::prereg_pdf
---

5.2.2025 mail from Abby

# Study Information

## Title
<!-- Provide the working title of your study. It may be the same title that you submit for publication of your final manuscript, but it is not a requirement. The title should be a specific and informative description of a project. Vague titles such as 'Fruit fly preregistration plan' are not appropriate.

Example: Effect of sugar on brownie tastiness. -->

`r rmarkdown::metadata$title`

## Description
<!-- Please give a brief description of your study, including some background, the purpose of the of the study, or broad research questions. The description should be no longer than the length of an abstract. It can give some context for the proposed study, but great detail is not needed here for your preregistration.

Example: Though there is strong evidence to suggest that sugar affects taste preferences, the effect has never been demonstrated in brownies. Therefore, we will measure taste preference for four different levels of sugar concentration in a standard brownie recipe to determine if the effect exists in this pastry. -->

Having a donor self-identity is considered a key motivational factor for sustained blood donation behaviour (Callero, 1985; Ferguson et al., 2020). Within Identity Theory, when a new donor first gives blood they develop a role-identity (I am a blood donor). As a novice donor continues to donate, this role-identity is validated, and thus blood donation becomes part of a donor’s self-identity (Being a blood donor is an important part of who I am; Grube & Piliavin, 2000; Masser, White, Hyde, Terry, & Robinson, 2009). However, despite the link between donor self-identity and donor retention, there are currently no interventions which can foster a sense of identity in donors that do not rely exclusively on repeat donation behaviour. 
Psychological ownership (this is mine) is the sense of ownership experienced over both material (e.g., blood collection agency) and abstract immaterial (e.g., blood donation behaviour) possessions (Pierce et al., 2001). The ownership volunteers experience over a non-profit organisation has been linked to sustained volunteering behaviour (Ainsworth, 2020). Further, psychological ownership is theoretically associated with identity as it has been proposed that through possession identity is established and maintained (Dittmar, 1992; Pierce et al., 2001). 
In an Australian donor context, we have provided initial support for the idea that psychological ownership experienced towards a blood collection agency both increases with past donation experience and is positively associated with blood donor self-identity (Edwards et al., 2023). Further, in a series of semi-structured interviews we have shown that Australian blood donors primarily expressed ownership over the blood donation process (role behaviour), not the blood collection agency (organisation). As such, interventions that enhance the amount of psychological ownership donors experience over their donation have the potential to impact donor retention.

FRCBS

## Hypotheses
<!-- List specific, concise, and testable hypotheses. Please state if the hypotheses are directional or non-directional. If directional, state the direction. A predicted effect is also appropriate here. If a specific interaction or moderation is important to your research, you can list that as a separate hypothesis.

Example: If taste affects preference, then mean preference indices will be higher with higher concentrations of sugar. -->

1. The blood donors reported psychological ownership over FRCBS is positively associated with self-identity, which, in turn, is positively associated with intentions to donate blood. That is the lower bound of effect estimate of 95% standardized confidence or credible interval of the indirect pathway between past behavior and intention via both psychological ownership and self-identity (ind_eff =  a1* d21 * b2) is higher than zero.

2. Psychological ownership over FRCBS increases with donation experience. That is the lower bound of effect estimate of 95% standardized confidence or credible interval of direct effect of past donation behavior to psychological ownership is higher than zero.

# Design Plan
<!-- In this section, you will be asked to describe the overall design of your study. Remember that this research plan is designed to register a single study, so if you have multiple experimental designs, please complete a separate preregistration. -->

## Study type

**Observational Study**. Data is collected from study subjects that are not randomly assigned to a treatment. This includes surveys, natural experiments, and regression discontinuity designs.

## Blinding
<!-- Blinding describes who is aware of the experimental manipulations within a study. Select all that apply. Is there any additional blinding in this study? -->

No blinding is involved in this study.

## Study design
<!-- Describe your study design. Examples include two-group, factorial, randomized block, and repeated measures. Is it a between (unpaired), within-subject (paired), or mixed design? Describe any counterbalancing required. Typical study designs for observation studies include cohort, cross sectional, and case-control studies.

This question has a variety of possible answers. The key is for a researcher to be as detailed as is necessary given the specifics of their design. Be careful to determine if every parameter has been specified in the description of the study design. There may be some overlap between this question and the following questions. That is OK, as long as sufficient detail is given in one of the areas to provide all of the requested information. For example, if the study design describes a complete factorial, 2 X 3 design and the treatments and levels are specified previously, you do not have to repeat that information.

Example: We have a between subjects design with 1 factor (sugar by mass) with 4 levels. -->

Cross sectional survey stud ymodelled with a structural equation model.

## Randomization
<!-- If you are doing a randomized study, how will you randomize, and at what level? Typical randomization techniques include: simple, block, stratified, and adaptive covariate randomization. If randomization is required for the study, the method should be specified here, not simply the source of random numbers.

Example: We will use block randomization, where each participant will be randomly assigned to one of the four equally sized, predetermined blocks. The random number list used to create these four blocks will be created using the web applications available at https://random.org. -->

A set of blood donors will be randomly selected from all active blood donors in FRBS register.

# Sampling Plan
<!-- In this section we’ll ask you to describe how you plan to collect samples, as well as the number of samples you plan to collect and your rationale for this decision. Please keep in mind that the data described in this section should be the actual data used for analysis, so if you are using a subset of a larger dataset, please describe the subset that will actually be used in your study. -->

## Existing data
<!-- Preregistration is designed to make clear the distinction between confirmatory tests, specified prior to seeing the data, and exploratory analyses conducted after observing the data. Therefore, creating a research plan in which existing data will be used presents unique challenges. Please select the description that best describes your situation. Please do not hesitate to contact us if you have questions about how to answer this question (prereg@cos.io). -->

**Registration prior to creation of data**. As of the date of submission of this research plan for preregistration, the data have not yet been collected, created, or realized.

## Explanation of existing data
<!-- If you indicate that you will be using some data that already exist in this study, please describe the steps you have taken to assure that you are unaware of any patterns or summary statistics in the data. This may include an explanation of how access to the data has been limited, who has observed the data, or how you have avoided observing any analysis of the specific data you will use in your study.

An appropriate instance of using existing data would be collecting a sample size much larger than is required for the study, using a small portion of it to conduct exploratory analysis, and then registering one particular analysis that showed promising results. After registration, conduct the specified analysis on that part of the dataset that had not been investigated by the researcher up to that point.

Example: An appropriate instance of using existing data would be collecting a sample size much larger than is required for the study, using a small portion of it to conduct exploratory analysis, and then registering one particular analysis that showed promising results. After registration, conduct the specified analysis on that part of the dataset that had not been investigated by the researcher up to that point. -->

No pre-exisiting data is used.

## Data collection procedures
<!-- Please describe the process by which you will collect your data. If you are using human subjects, this should include the population from which you obtain subjects, recruitment efforts, payment for participation, how subjects will be selected for eligibility from the initial pool (e.g. inclusion and exclusion rules), and your study timeline. For studies that donÍt include human subjects, include information about how you will collect samples, duration of data gathering efforts, source or location of samples, or batch numbers you will use.

The answer to this question requires a specific set of instructions so that another person could repeat the data collection procedures and recreate the study population. Alternatively, if the study population would be unable to be reproduced because it relies on a specific set of circumstances unlikely to be recreated (e.g., a community of people from a specific time and location), the criteria and methods for creating the group and the rationale for this unique set of subjects should be clear.

Example: Participants will be recruited through advertisements at local pastry shops. Participants will be paid $10 for agreeing to participate (raised to $30 if our sample size is not reached within 15 days of beginning recruitment). Participants must be at least 18 years old and be able to eat the ingredients of the pastries. -->

An e-mail inviting to fill an online questionnaire will be sent to the randomly selected blood donors. No recompensation will be offered to participants.Participants must be between 18 and 65 years old.

## Sample size
<!-- Describe the sample size of your study. How many units will be analyzed in the study? This could be the number of people, birds, classrooms, plots, interactions, or countries included. If the units are not individuals, then describe the size requirements for each unit. If you are using a clustered or multilevel design, how many units are you collecting at each level of the analysis? For some studies, this will simply be the number of samples or the number of clusters. For others, this could be an expected range, minimum, or maximum number.

Example: Our target sample size is 280 participants. We will attempt to recruit up to 320, assuming that not all will complete the total task. -->

Our target sample size is 1000 participants.

## Sample size rationale
<!-- This could include a power analysis or an arbitrary constraint such as time, money, or personnel. This gives you an opportunity to specifically state how the sample size will be determined. A wide range of possible answers is acceptable; remember that transparency is more important than principled justifications. If you state any reason for a sample size upfront, it is better than stating no reason and leaving the reader to "fill in the blanks." Acceptable rationales include: a power analysis, an arbitrary number of subjects, or a number based on time or monetary constraints.

Example: We used the software program G*Power to conduct a power analysis. Our goal was to obtain .95 power to detect a medium effect size of .25 at the standard .05 alpha error probability. -->

The blood donor participants (n = 255) and the standardised regression coefficients from our previous serial mediation model (Edwards et al., 2023), were imputed into a Monte Carlo Power analysis for indirect effects (Schoemann, Boulton, & Short, 2017) to estimate sample size. For an alpha set at .05, usable data from 369 participants is required to achieve .9 power for our intended serial mediation model. With a target sample of 1000 participants we hope to carry out sex stratified analysis and further exploratory analysis.

## Stopping rule
<!-- If your data collection procedures do not give you full control over your exact sample size, specify how you will decide when to terminate your data collection.

You may specify a stopping rule based on p-values only in the specific case of sequential analyses with pre-specified checkpoints, alphas levels, and stopping rules. Unacceptable rationales include stopping based on p-values if checkpoints and stopping rules are not specified. If you have control over your sample size, then including a stopping rule is not necessary, though it must be clear in this question or a previous question how an exact sample size is attained.

Example: We will post participant sign-up slots by week on the preceding Friday night, with 20 spots posted per week. We will post 20 new slots each week if, on that Friday night, we are below 320 participants. -->

Enter your response here.

# Variables
<!-- In this section you can describe all variables (both manipulated and measured variables) that will later be used in your confirmatory analysis plan. In your analysis plan, you will have the opportunity to describe how each variable will be used. If you have variables which you are measuring for exploratory analyses, you are not required to list them, though you are permitted to do so. -->

## Manipulated variables
<!-- Describe all variables you plan to manipulate and the levels or treatment arms of each variable. This is not applicable to any observational study. For any experimental manipulation, you should give a precise definition of each manipulated variable. This must include a precise description of the levels at which each variable will be set, or a specific definition for each categorical treatment. For example, “loud or quiet,” should instead give either a precise decibel level or a means of recreating each level. 'Presence/absence' or 'positive/negative' is an acceptable description if the variable is precisely described.

Example: We manipulated the percentage of sugar by mass added to brownies. The four levels of this categorical variable are: 15%, 20%, 25%, or 40% cane sugar by mass. -->

All likert scale measured variables will be treated as numeric as well number of life time donations (bd02) and number of donations in last years (bd03). Last donation date (bd04) will transformed "bd04 / 10^-(2 - str_length(as.numeric(max(bd04))))".

For sum score model latent variables will be formed by summing measured items

Psychological ownership = po01 + po02 + po03

Self-identity           = si01 + si02 + si03

Intention               = it01 + it02 + it03

For structural equation models using latent variables

## Measured variables
<!-- Describe each variable that you will measure. This will include outcome measures, as well as any predictors or covariates that you will measure. You do not need to include any variables that you plan on collecting if they are not going to be included in the confirmatory analyses of this study.

Observational studies and meta-analyses will include only measured variables. As with the previous questions, the answers here must be precise. For example, 'intelligence,' 'accuracy,' 'aggression,' and 'color' are too vague. Acceptable alternatives could be 'IQ as measured by Wechsler Adult Intelligence Scale' 'percent correct,' 'number of threat displays,' and 'percent reflectance at 400 nm.'

Example: The single outcome variable will be the perceived tastiness of the single brownie each participant will eat. We will measure this by asking participants ‘How much did you enjoy eating the brownie’ (on a scale of 1-7, 1 being 'not at all', 7 being 'a great deal') and 'How good did the brownie taste' (on a scale of 1-7, 1 being 'very bad', 7 being 'very good'). -->

Enter your response here.

Items
<table>
<tr>
<td></td>
<td>Psychological ownership</td>
</tr>
<tr>
<td>
po01 1. I feel a very high degree of personal ownership of donating blood.
</td>
</tr>
<td>
po02 2. I feel like donating blood belongs to me.
</td> </tr>
<tr>
<td>
po03 3. I feel like I own donating blood.
</td> </tr>
<tr> <td></td>

<td>
Self-identity
</td> </tr>
<tr>
<td>
si01 1. Donating blood is important to me.
</td> <tr>
si02 2. I am like the kind of person who donates blood.
</td> </tr>
<tr> 
<td>
si03 3. Donating blood is an important part of who I am.
</td> </tr>
<tr> <td></td>

<td>
Intention
</td></tr>
<tr>
<td>
it01 1. I would like to donate blood in the future.
</td>   </tr>
<tr>
<td>
it02 2. I intend to donate blood in the future.
</td> </tr>
<tr>
<td>
it03 3. I intend to make more than just a one-off blood donation.
</td></tr>
<tr><td></td>
<td>
Intention to donate for emergency
</td></tr>
<tr>
<td>
ie01 1. I would be willing to donate blood if I received an urgent callout from the Finnish Red Cross Blood Service.
</td></tr>
<tr><td></td> 
<td>
Antecedents of psychological ownership
</td></tr>
<tr><td></td>

<td>
Control
</td></tr>
<tr>
<td>
ac01  1. I have influence over the things that affect me while donating blood.
</td></tr>
<tr>
<td>
ac02 2. I control the location and scheduling for donating blood.
</td></tr>
<tr>
<td>
ac03 3. I influence decisions related to my blood donation.
</td></tr>
<tr>
<td>
ac04 4. In general, I have control over donating blood.
</td></tr>
<tr><td></td>
<td>
Intimate knowledge
</td></tr>
<tr>
<td>
ai01 1. I am intimately familiar with what is going on with regard to donating blood.
</td></tr>
<tr>
<td>
ai02 2. I have a depth of knowledge as it relates to donating blood.
</td></tr>
<tr>
<td>
ai03 3. I have a comprehensive understanding of donating blood.
</td></tr>
<tr>
<td>
ai04 4. I have a broad understanding of donating blood.
</td></tr>
<tr><td></td>

<td>
Self-Investment
</td></tr>
<td>
iv01 1.	I have invested a major part of myself into donating blood.
</td></tr>
<tr>
<td>
iv02 2. I have invested a significant amount of my time into donating blood.
</td></tr>
<tr>
<td>
iv03 3. Overall, I invested a lot into donating blood.
</td></tr>
<tr><td></td>

<td>
Blood donation history
</td></tr>
<tr>
<td>
bd01 1. Do you believe that you are currently eligible to donate blood [yes, no, unsure]
</td></tr>
<tr>
<td>
bd02 2. How many times in total have you donated blood at the Finnish Red Cross Blood Service Finland? [select number]
</td></tr>
<tr>
<td>
bd03 3. How many times in the last two years have you donated at the Finnish Red Cross Blood Service? [select number]
</td></tr>
<tr>
<td>
bd04 4. When did you last attend a Finnish Red Cross blood donation centre with the intention of donating blood? [select date]
</td></tr>
<tr><td></td>

<td>
Demographics
</td></tr>
<tr>
<td>
de01 1. What is your gender? [man, woman, non-binary, prefer to not disclose, prefer to self-describe (open response)]
</td></tr>
<tr>
<td>
de02 2. How old are you today? [select number]
</td></tr>
</table>

## Indices
<!-- If any measurements are  going to be combined into an index (or even a mean), what measures will you use and how will they be combined? Include either a formula or a precise description of your method. If your are using a more complicated statistical method to combine measures (e.g. a factor analysis), you can note that here but describe the exact method in the analysis plan section.

If you are using multiple pieces of data to construct a single variable, how will this occur? Both the data that are included and the formula or weights for each measure must be specified. Standard summary statistics, such as "means" do not require a formula, though more complicated indices require either the exact formula or, if it is an established index in the field, the index must be unambiguously defined. For example, "biodiversity index" is too broad, whereas "Shannon’s biodiversity index" is appropriate.

Example: We will take the mean of the two questions above to create a single measure of 'brownie enjoyment.'  -->

sum scores and latent variables

# Analysis Plan
<!-- You may describe one or more confirmatory analysis in this preregistration. Please remember that all analyses specified below must be reported in the final article, and any additional analyses must be noted as exploratory or hypothesis generating.

A confirmatory analysis plan must state up front which variables are predictors (independent) and which are the outcomes (dependent), otherwise it is an exploratory analysis. You are allowed to describe any exploratory work here, but a clear confirmatory analysis is required. -->

Bivariate correlations will be calculated between all primary variables included in the survey (past behaviour, psychological ownership, self-identity, intention). Following this, a serial mediation will be conducted. Previous blood donation behaviour (from the last two years) will be entered as the causal variable, psychological ownership and then identity as the serial mediators, and blood donation intentions as the outcome variable.

## Statistical models
<!-- What statistical model will you use to test each hypothesis? Please include the type of model (e.g. ANOVA, multiple regression, SEM, etc) and the specification of the model (this includes each variable that will be included as predictors, outcomes, or covariates). Please specify any interactions, subgroup analyses, pairwise or complex contrasts, or follow-up tests from omnibus tests. If you plan on using any positive controls, negative controls, or manipulation checks you may mention that here. Remember that any test not included here must be noted as an exploratory test in your final article.

This is perhaps the most important and most complicated question within the preregistration. As with all of the other questions, the key is to provide a specific recipe for analyzing the collected data. Ask yourself: is enough detail provided to run the same analysis again with the information provided by the user? Be aware for instances where the statistical models appear specific, but actually leave openings for the precise test. See the following examples:

- If someone specifies a 2x3 ANOVA with both factors within subjects, there is still flexibility with the various types of ANOVAs that could be run. Either a repeated measures ANOVA (RMANOVA) or a multivariate ANOVA (MANOVA) could be used for that design, which are two different tests. 
- If you are going to perform a sequential analysis and check after 50, 100, and 150 samples, you must also specify the p-values you’ll test against at those three points.

Example:  We will use a one-way between subjects ANOVA to analyze our results. The manipulated, categorical independent variable is 'sugar' whereas the dependent variable is our taste index. -->

SEM include R from preregistartion code

## Transformations
<!-- If you plan on transforming, centering, recoding the data, or will require a coding scheme for categorical variables, please describe that process. If any categorical predictors are included in a regression, indicate how those variables will be coded (e.g. dummy coding, summation coding, etc.) and what the reference category will be.

Example: The "Effect of sugar on brownie tastiness" does not require any additional transformations. However, if it were using a regression analysis and each level of sweet had been categorically described (e.g. not sweet, somewhat sweet, sweet, and very sweet), 'sweet' could be dummy coded with 'not sweet' as the reference category. -->

as numeric

## Inference criteria
<!-- What criteria will you use to make inferences? Please describe the information youÍll use (e.g. p-values, bayes factors, specific model fit indices), as well as cut-off criterion, where appropriate. Will you be using one or two tailed tests for each of your analyses? If you are comparing multiple conditions or testing multiple hypotheses, will you account for this?

p-values, confidence intervals, and effect sizes are standard means for making an inference, and any level is acceptable, though some criteria must be specified in this or previous fields. Bayesian analyses should specify a Bayes factor or a credible interval. If you are selecting models, then how will you determine the relative quality of each? In regards to multiple comparisons, this is a question with few "wrong" answers. In other words, transparency is more important than any specific method of controlling the false discovery rate or false error rate. One may state an intention to report all tests conducted or one may conduct a specific correction procedure; either strategy is acceptable.

Example: We will use the standard p<.05 criteria for determining if the ANOVA and the post hoc test suggest that the results are significantly different from those expected if the null hypothesis were correct. The post-hoc Tukey-Kramer test adjusts for multiple comparisons. -->

model selection and then ci

## Data exclusion
<!-- How will you determine what data or samples, if any, to exclude from your analyses? How will outliers be handled? Will you use any awareness check? Any rule for excluding a particular set of data is acceptable. One may describe rules for excluding a participant or for identifying outlier data.

Example: No checks will be performed to determine eligibility for inclusion besides verification that each subject answered each of the three tastiness indices. Outliers will be included in the analysis. -->

All donors answering to question

## Missing data
<!-- How will you deal with incomplete or missing data? Any relevant explanation is acceptable. As a final reminder, remember that the final analysis must follow the specified plan, and deviations must be either strongly justified or included as a separate, exploratory analysis.

Example: If a subject does not complete any of the three indices of tastiness, that subject will not be included in the analysis. -->

Enter your response here.

## Exploratory analyses (optional)
<!-- If you plan to explore your data set to look for unexpected differences or relationships, you may describe those tests here. An exploratory test is any test where a prediction is not made up front, or there are multiple possible tests that you are going to use. A statistically significant finding in an exploratory test is a great way to form a new confirmatory hypothesis, which could be registered at a later time.

Example: We expect that certain demographic traits may be related to taste preferences. Therefore, we will look for relationships between demographic variables (age, gender, income, and marital status) and the primary outcome measures of taste preferences. -->

Qualitative content analysis will be conducted on participants answers to the open-response question (What do you remember about the first time you donated blood?). Bivariate correlations will be calculated to compare donors’ willingness to respond to a call-out to past blood donation behaviour, psychological ownership, self-identity, and intention to donate blood.

# Other

## Other 
<!-- If there is any additional information that you feel needs to be included in your preregistration, please enter it here. Literature cited, disclosures of any related work such as replications or work that uses the same data, or other context that will be helpful for future readers would be appropriate here. -->

\includepdf[pages={-}]{preregistration_code.pdf}

# References
## 
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
\noindent

### 2025-04-07 preregistration_code.Rmd
include code in prereg

@@
 title: "Supplementary file 1: Preregistered analysis"
-author: "Abigail Edwards and Mikko Arvas"
+author: "Abigail Edwards, Lauri Hietajärvi and Mikko Arvas"
 date: "`r Sys.time()`"
 output:
-  html_document:
-    toc: yes
-    theme: united
   pdf_document:
     toc: yes
+
+always_allow_html: true
 ---
@@
 TODO:
+
  -remove superfluous outputs
+
  -check discussion with Lauri, was there anything else to do?
+
  -check model saving
+
  -fill out the preregistration form
+
+ -are Abigail's sum score coefficients bootstrapped?
 
@@
 library(lavaanPlot)
+library(kableExtra)
 #library(semPlot) # update plots to semPlot later
@@
 ```
-
-Items
-<table>
-<tr>
-<td></td>
-<td> Psychological ownership</td>
-</tr>
-<tr>
-<td>po01</td>
-<td>
-1. I feel a very high degree of personal ownership of donating blood.
-</td>
-</tr>
-<tr>
-<td>po02</td>
-<td>
-2. I feel like donating blood belongs to me.
-</td> </tr>
-<tr>
-<td>po03</td> <td>
-3. I feel like I own donating blood.
-</td> </tr>
-<tr> <td></td>
-
-<td>
-Self-identity
-</td> </tr>
-<tr>
-<td>si01</td>
-<td>
-1. Donating blood is important to me.
-</td> <tr>
-<td>si02</td> <td>
-2. I am like the kind of person who donates blood.
-</td> </tr>
-<tr> <td>si03</td> 
-<td>
-3. Donating blood is an important part of who I am.
-</td> </tr>
-<tr> <td></td>
-
-<td>
-Intention
-</td></tr>
-<tr><td>it01</td>
-<td>
-1. I would like to donate blood in the future.
-</td>   </tr>
-<tr><td>it02</td>
-<td>
-2. I intend to donate blood in the future.
-</td> </tr>
-<tr><td>it03</td>
-
-<td>
-3. I intend to make more than just a one-off blood donation.
-</td></tr>
-<tr><td></td>
-<td>
-Intention to donate for emergency
-</td></tr>
-<tr><td>ie01</td>
-<td>
-1. I would be willing to donate blood if I received an urgent callout from the Finnish Red Cross Blood Service.
-</td></tr>
-<tr><td></td> 
-<td>
-Antecedents of psychological ownership
-</td></tr>
-<tr><td></td>
-
-<td>
-Control
-</td></tr>
-<tr><td>ac01</td>
-<td>
-1. I have influence over the things that affect me while donating blood.
-</td></tr>
-<tr><td>ac02</td>
-<td>
-2. I control the location and scheduling for donating blood.
-</td></tr>
-<tr><td>ac03</td>
-<td>
-3. I influence decisions related to my blood donation.
-</td></tr>
-<tr><td>ac04</td>
-<td>
-4. In general, I have control over donating blood.
-</td></tr>
-<tr><td></td>
-<td>
-Intimate knowledge
-</td></tr>
-<tr><td>ai01</td>
-
-<td>
-1. I am intimately familiar with what is going on with regard to donating blood.
-</td></tr>
-<tr><td>ai02</td>
-<td>
-2. I have a depth of knowledge as it relates to donating blood.
-</td></tr>
-<tr><td>ai03</td>
-<td>
-3. I have a comprehensive understanding of donating blood.
-</td></tr>
-<tr><td>ai04</td>
-<td>
-4. I have a broad understanding of donating blood.
-</td></tr>
-<tr><td></td>
-
-<td>
-Self-Investment
-</td></tr>
-<tr><td>iv01</td>
-<td>
-1.	I have invested a major part of myself into donating blood.
-</td></tr>
-<tr><td>iv02</td>
-<td>
-2. I have invested a significant amount of my time into donating blood.
-</td></tr>
-<tr><td>iv03</td>
-<td>
-3. Overall, I invested a lot into donating blood.
-</td></tr>
-<tr><td></td>
-
-<td>
-Blood donation history
-</td></tr>
-<tr><td>bd01</td>
-<td>
-1. Do you believe that you are currently eligible to donate blood [yes, no, unsure]
-</td></tr>
-<tr><td>bd02</td>
-<td>
-2. How many times in total have you donated blood at the Finnish Red Cross Blood Service Finland? [select number]
-</td></tr>
-<tr><td>bd03</td>
-<td>
-3. How many times in the last two years have you donated at the Finnish Red Cross Blood Service? [select number]
-</td></tr>
-<tr><td>bd04</td>
-<td>
-4. When did you last attend a Finnish Red Cross blood donation centre with the intention of donating blood? [select date]
-</td></tr>
-<tr><td></td>
-
-<td>
-Demographics
-</td></tr>
-<tr><td>de01</td>
-<td>1. What is your gender? [man, woman, non-binary, prefer to not disclose, prefer to self-describe (open response)]
-</td></tr>
-<tr><td>de02</td>
-<td>
-2. How old are you today? [select number]
-</td></tr>
-</table>
 
@@
   add_n()  # add column with total number of non-missing observations
-table1
+as_kable_extra(table1, format = "latex")
 ```
@@
   graph_options = list(rankdir = "LR")
-           )
+           ) %>% lavaanPlot::embed_plot_pdf(here("figures/cfa01.pdf"))
 ```
@@
 
-  )
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem01.pdf"))
 ```
@@
   sig=0.001
-  )
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
 ```
@@
 
-  )
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem01.pdf"))
 ```
@@
   sig=0.001
-  )
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllssem02.pdf"))
 ```
@@
   sig=0.001
-  )
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))
 
@@
   sig=0.001
-  )
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem03.pdf"))
 
@@
 if (fitmodels) {
-  save(models,file=here(str_replace(str_c("results/",datafile),"\\.data\\.",".models.")))
+  save(models,file=here("prereg_models.rdata"))
 }

### 2025-04-07 preregistration.Rmd
reorganise models

@@
 
+TODO:
+
+- references
+
 # Study Information
@@
 As such, interventions that enhance the amount of psychological ownership donors experience over their donation have the potential to impact donor retention.
-
-FRCBS
+In this study we try to repeat results from Edwards et al., 2023 in Finland by collecting similar questionnaire data and analysing them with a structural equation model.
+We include pre-registered R analysis code for direct repetition of Edwards et al., 2023 in Appendix 1.
 
@@
 
-model selection and then ci
+We will model the hypothesis through alternative models:
+
+a. Sum score model with robust estimator
+
+b. Sum score model with bootstrapping
+
+c. Latent factor model with robust estimator
+
+d. Latent factor model with boostrapping
+
+Among
 
@@
 
-All donors answering to question
+All donors answering to question de01 other than "man" or "women" will be excluded as donation policies depend on sex.
+All donors answering to question bd01 "no" will be excluded as we are interested in person eligible to donate.
 
@@
 
-Enter your response here.
+The survey platform only reports completed questionnaires and all questions are obligatory. Hence there will be no missing data.
 

### 2025-04-07 preregistration_code.Rmd
reorganise models

@@
 ---
-title: "Supplementary file 1: Preregistered analysis"
+title: "Appendix 1: Preregistered analysis with simulated data"
 author: "Abigail Edwards, Lauri Hietajärvi and Mikko Arvas"
@@
  -are Abigail's sum score coefficients bootstrapped?
+
+ -report effects of exclusions
 
@@
   #lavaan cannot deal with unordered factor with more then 2 levels
-  filter(de01 == 'man' | de01 == "women") %>% # exclusion
-  filter(bd01 != 'no') %>% # exclusion
+  filter(de01 == 'man' | de01 == "women") %>% # EXCLUSION
+  filter(bd01 != 'no') %>% # EXCLUSION
   mutate(
@@
 
-### Sum score model
+### a) Sum score model with robust estimator
 
@@
 
-### SEM
+###  b) Sum score model with bootstrapping
+
+```{r}
+nBoots <- 1000
+
+if (fitmodels) {
+#fit
+
+# https://groups.google.com/g/lavaan/c/A6KCjXAZl-Q good example code
+
+fit.po03ss <- sem(po02ss.model,  
+                    estimator = "ML", # 
+                    data = data.fit2, # data that is completely numeric
+                bootstrap = nBoots,
+                se = "boot",
+                auto.fix.first = FALSE,
+                std.lv = TRUE
+                )
+models[['fit.po03ss']] <- fit.po03ss
+}
+
+lavaan::summary(fit.po03ss,standardized = TRUE)
+
+```
+
+```{r}
+lavaanPlot(
+  model = fit.po03ss, 
+   labels = labelsforplottingnull,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))
+
+```
+
+```{r}
+fit.po03ss.best <- parameterEstimates(fit.po03ss)
+fit.po03ss.best %>% select(-se,-z) %>% filter(op == "~")
+```
+
+### c) Latent factor model with robust estimator
 
@@
 
-### Sum score with bootstrapping
-
-```{r}
-nBoots <- 1000
-
-if (fitmodels) {
-#fit
-
-# https://groups.google.com/g/lavaan/c/A6KCjXAZl-Q good example code
-
-fit.po03ss <- sem(po02ss.model,  
-                    estimator = "ML", # bootstrapping does not work with estimators designed for factors
-                    data = data.fit2, # data that is completely numeric
-                bootstrap = nBoots,
-                se = "boot",
-                auto.fix.first = FALSE,
-                std.lv = TRUE
-                )
-models[['fit.po03ss']] <- fit.po03ss
-}
-
-lavaan::summary(fit.po03ss,standardized = TRUE)
-
-```
-
-```{r}
-lavaanPlot(
-  model = fit.po03ss, 
-   labels = labelsforplottingnull,
-  node_options = list(shape = "box", fontname = "Helvetica"), 
-  edge_options = list(color = "grey"), 
-  coefs = TRUE, 
-  stand = TRUE,
-  sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))
-
-```
-
-```{r}
-fit.po03ss.best <- parameterEstimates(fit.po03ss)
-fit.po03ss.best %>% select(-se,-z) %>% filter(op == "~")
-```
-
-### SEM with bootstrapping
+### d) Latent factor model with boostrapping
 
@@
 fit.po03.best %>% select(-se,-z) %>% filter(op == "~")
-```
-
-### SEM with Bayes
-
-```{r}
-po02b.model <- ' 
-#measurement model
-psycown =~ po01 + po02 + po03
-selfid =~ si01 + si02 + si03
-intention =~ it01 + it02 + it03
-blooddonor =~ bd02 + bd03 +bd04
-
-# #mediation
- psycown ~ a1 *  blooddonor
- selfid ~ a2 * blooddonor + d21 * psycown
-# 
-# #regression
- intention ~  de01 + de02 + ie01 +  cp * blooddonor + b1 * psycown + b2 * selfid
-ind_eff := a1* d21 * b2
-'
-
-cat(po02.model)
-
-```
-
-```{r}
-# This does not work!
-# #blavaan
-#tmp <- as.data.frame(data.fit2)
- # fit.po04 <- blavaan(po02.model,  
- #         #data = data.fit2,
- #         data = tmp,
- #         auto.fix.first = FALSE,
- #         std.lv = TRUE
- #)
-#Error in validObject(.Object) : 
-#invalid class “Fit” object: invalid object for slot "fx.group" in class "Fit": got class "NULL", should be or extend class "numeric"
-#We will try to fix this with real data
-```
-
-```{r}
-
-# #This works!
-# data(HolzingerSwineford1939, package = "lavaan")
-# # The Holzinger and Swineford (1939) example
-# HS.model <- ' visual  =~ x1 + x2 + x3
-#               textual =~ x4 + x5 + x6
-#               speed   =~ x7 + x8 + x9 '
-# 
-# fit <- blavaan(HS.model, data = HolzingerSwineford1939,
-#                auto.var = TRUE, auto.fix.first = TRUE,
-#                auto.cov.lv.x = TRUE)
-# summary(fit)
-# coef(fit)
-
 ```

### 2025-04-08 preregistration.Rmd
sampling

@@
 Psychological ownership over FRCBS increases with donation experience.
-That is the lower bound of effect estimate of 95% standardized confidence or credible interval of direct effect of past donation behavior to psychological ownership is higher than zero.
+That is the lower bound of effect estimate of 95% standardized confidence or credible interval of direct effect of past donation behavior to psychological ownership (a1) is higher than zero.
 
@@
 
-Cross sectional survey stud ymodelled with a structural equation model.
+Cross sectional survey study modelled with a structural equation model.
 
@@
 
-A set of blood donors will be randomly selected from all active blood donors in FRBS register.
+A set of blood donors will be randomly selected from all blood donors in FRCBS register who have donated at least once during the 730 preceding the day the questionnaire e-mail will be sent i.e.
+5.5.2025 .
 
@@
 An e-mail inviting to fill an online questionnaire will be sent to the randomly selected blood donors.
-No recompensation will be offered to participants.Participants must be between 18 and 65 years old.
+No recompensation will be offered to participants.Participants must be between 18 and 72 years old.
 
@@
 
-Our target sample size is 1000 participants.
+Our target sample size is 1000 participants. Based on previous participation rates an email will be sent to 5000 random donors.
 
@@
 
-Enter your response here.
+The online questionnaire will be open for 2 weeks i.e. from 5.5.2025 - to 18.5.2025.
 
@@
 
-All likert scale measured variables will be treated as numeric as well number of life time donations (bd02) and number of donations in last years (bd03).
-Last donation date (bd04) will transformed "bd04 / 10^-(2 - str_length(as.numeric(max(bd04))))".
-
-For sum score model latent variables will be formed by summing measured items
-
-Psychological ownership = po01 + po02 + po03
-
-Self-identity           = si01 + si02 + si03
-
-Intention               = it01 + it02 + it03
-
-For structural equation models using latent variables
-
 ## Measured variables
@@
 
-Enter your response here.
+All Likert items are scored 1 (strongly disagree) to 7 (strongly agree).
 
-Items
+Items:
 <table>
@@
 <td>
-po01 1. I feel a very high degree of personal ownership of donating blood.
+po01 1. I feel a very high degree of personal ownership of donating blood. [Likert]
 </td>
@@
 <td>
-po02 2. I feel like donating blood belongs to me.
+po02 2. I feel like donating blood belongs to me. [Likert]
 </td> </tr>
@@
 <td>
-po03 3. I feel like I own donating blood.
+po03 3. I feel like I own donating blood. [Likert]
 </td> </tr>
@@
 <td>
-si01 1. Donating blood is important to me.
+si01 1. Donating blood is important to me. [Likert]
 </td> <tr>
-si02 2. I am like the kind of person who donates blood.
+si02 2. I am like the kind of person who donates blood. [Likert]
 </td> </tr>
@@
 <td>
-si03 3. Donating blood is an important part of who I am.
+si03 3. Donating blood is an important part of who I am. [Likert]
 </td> </tr>
@@
 <td>
-it01 1. I would like to donate blood in the future.
+it01 1. I would like to donate blood in the future. [Likert]
 </td>   </tr>
@@
 <td>
-it02 2. I intend to donate blood in the future.
+it02 2. I intend to donate blood in the future. [Likert]
 </td> </tr>
@@
 <td>
-it03 3. I intend to make more than just a one-off blood donation.
+it03 3. I intend to make more than just a one-off blood donation. [Likert]
 </td></tr>
@@
 <td>
-ie01 1. I would be willing to donate blood if I received an urgent callout from the Finnish Red Cross Blood Service.
+ie01 1. I would be willing to donate blood if I received an urgent callout from the Finnish Red Cross Blood Service. [Likert]
 </td></tr>
@@
 <td>
-ac01  1. I have influence over the things that affect me while donating blood.
+ac01  1. I have influence over the things that affect me while donating blood. [Likert]
 </td></tr>
@@
 <td>
-ac02 2. I control the location and scheduling for donating blood.
+ac02 2. I control the location and scheduling for donating blood. [Likert]
 </td></tr>
@@
 <td>
-ac03 3. I influence decisions related to my blood donation.
+ac03 3. I influence decisions related to my blood donation. [Likert]
 </td></tr>
@@
 <td>
-ac04 4. In general, I have control over donating blood.
+ac04 4. In general, I have control over donating blood. [Likert]
 </td></tr>
@@
 <td>
-ai01 1. I am intimately familiar with what is going on with regard to donating blood.
+ai01 1. I am intimately familiar with what is going on with regard to donating blood. [Likert]
 </td></tr>
@@
 <td>
-ai02 2. I have a depth of knowledge as it relates to donating blood.
+ai02 2. I have a depth of knowledge as it relates to donating blood. [Likert]
 </td></tr>
@@
 <td>
-ai03 3. I have a comprehensive understanding of donating blood.
+ai03 3. I have a comprehensive understanding of donating blood. [Likert]
 </td></tr>
@@
 <td>
-ai04 4. I have a broad understanding of donating blood.
+ai04 4. I have a broad understanding of donating blood. [Likert]
 </td></tr>
@@
 <td>
-iv01 1.	I have invested a major part of myself into donating blood.
+iv01 1.	I have invested a major part of myself into donating blood. [Likert]
 </td></tr>
@@
 <td>
-iv02 2. I have invested a significant amount of my time into donating blood.
+iv02 2. I have invested a significant amount of my time into donating blood. [Likert]
 </td></tr>
@@
 <td>
-iv03 3. Overall, I invested a lot into donating blood.
+iv03 3. Overall, I invested a lot into donating blood. [Likert]
 </td></tr>
@@
 
-sum scores and latent variables
+For sum score model latent variables will be formed by summing measured items:
+
+Psychological ownership = po01 + po02 + po03
+
+Self-identity           = si01 + si02 + si03
+
+Intention               = it01 + it02 + it03
+
+For structural equation models using latent variables the same variables will be used as constituent items of the latent variable.
+Instead of using just question bd02 as in the original publication a donation history latent variable will be formed as:
+
+Blood donation history = bd02 + bd03 + bd04
 
@@
 
-SEM include R from preregistartion code
+We will model the data with a structural equation model. The exact model, R code to run it and example analysis with simulated data can be found in Appendix 1.
 
@@
 
-as numeric
+Number of life time donations (bd02) and number of donations in last years (bd03) and likert scale measured variables (po01-3, si01-03, it01-03, ac01-04, iv01-03) will be treated as numeric.
+Last donation date (bd04) will transformed "bd04 / 10^-(2 - str_length(as.numeric(max(bd04))))".
 
@@
 
-Among
+In supplement we will report results from all models, but as the main results we will report the model with highest Comparative Fit Index (CFI).
+95 % confidence or credible intervals will be used to decide if an effect is considered significant or not.
+
+Further details of the models can be found in Appendix 1.
 
@@
 All donors answering to question de01 other than "man" or "women" will be excluded as donation policies depend on sex.
-All donors answering to question bd01 "no" will be excluded as we are interested in person eligible to donate.
+All donors answering to question bd01 "no" will be excluded as we are interested in persons eligible to donate.
 

### 2025-04-08 preregistration_code.Rmd
sampling

@@
     term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
+    term =str_replace_all(term,"bd02","blooddonor:"),
+    term =str_replace_all(term,"blooddonor:","blooddonor"),
     estmet = as.factor(estmet),

### 2025-04-09 manuscript/manuscript.Rmd
references

@@
 author: 
-  - name          : "First Author"
+  - name          : "Abigail Edwards"
     affiliation   : "1"
@@
     address       : "Postal address"
-    email         : "my@email.com"
+    email         : "abigail.edwards@uq.edu.au"
     role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
@@
       - "Writing - Review & Editing"
-  - name          : "Ernst-August Doelle"
-    affiliation   : "1,2"
+  - name          : "Mikko Arvas"
+    affiliation   : "2"
     role:
@@
   - id            : "1"
-    institution   : "Wilhelm-Wundt-University"
+    institution   : "The University of Queensland, Australia"
   - id            : "2"
-    institution   : "Konstanz Business School"
+    institution   : "Finnish Red Cross Blood Service, Finland"
 
@@
 bibliography      : ["references.bib", "r-references.bib"]
+#create with zotero using "Better Bibtex" https://retorque.re/zotero-better-bibtex/
 
 csl               : vox-sanguinis.csl
+#https://www.zotero.org/styles?q=vox
 
@@
 
-This manuscript uses the Workflow for Open Reproducible Code in Science [WORCS version 0.1.1, @vanlissaWORCSWorkflowOpen2021] to ensure reproducibility and transparency.
+This manuscript uses the Workflow for Open Reproducible Code in Science [WORCS version 0.1.1, @lissaWORCSWorkflowOpen2021] to ensure reproducibility and transparency.
 All code <!--and data--> are available at <!--insert repository URL-->.
 
-This is an example of a non-essential citation [@@vanlissaWORCSWorkflowOpen2021]. If you change the rendering function to `worcs::cite_essential`, it will be removed.
+This is an example of a non-essential citation [@lissaWORCSWorkflowOpen2021]. If you change the rendering function to `worcs::cite_essential`, it will be removed.
 
@@
 
+<!--https://tex.stackexchange.com/questions/702678/generating-citation-keys-with-zotero-and-better-bibtex-plugin  Zotero library -> right click -> Better BibTex -> Refresh BibTex Key.
+Do this for each entry.
+Then export the collection/library as references.bib file to /manuscript folder for use with this file -->
+
 ::: {#refs custom-style="Bibliography"}

### 2025-04-09 preregistration.Rmd
references

@@
 title           : "Measuring psychological ownership and identity in a Finnish blood donor context"
-shorttitle      : "Donor ownership"
+shorttitle      : "Donation ownership"
 date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
@@
 
+csl             : vox-sanguinis.csl
+bibliography      : ["manuscript/references.bib", "manuscript/r-references.bib"]
+
 output: prereg::prereg_pdf
@@
 
-5.2.2025 mail from Abby
-
-TODO:
-
-- references
+<!-- to be submitted to https://prereg-psych.org/ >
 
@@
 
-Having a donor self-identity is considered a key motivational factor for sustained blood donation behaviour (Callero, 1985; Ferguson et al., 2020).
+Having a donor self-identity is considered a key motivational factor for sustained blood donation behavior [@calleroRoleidentitySalience1985,@fergusonTypologyBloodDonor2020].
 Within Identity Theory, when a new donor first gives blood they develop a role-identity (I am a blood donor).
-As a novice donor continues to donate, this role-identity is validated, and thus blood donation becomes part of a donor’s self-identity (Being a blood donor is an important part of who I am; Grube & Piliavin, 2000; Masser, White, Hyde, Terry, & Robinson, 2009).
-However, despite the link between donor self-identity and donor retention, there are currently no interventions which can foster a sense of identity in donors that do not rely exclusively on repeat donation behaviour.
-Psychological ownership (this is mine) is the sense of ownership experienced over both material (e.g., blood collection agency) and abstract immaterial (e.g., blood donation behaviour) possessions (Pierce et al., 2001).
-The ownership volunteers experience over a non-profit organisation has been linked to sustained volunteering behaviour (Ainsworth, 2020).
-Further, psychological ownership is theoretically associated with identity as it has been proposed that through possession identity is established and maintained (Dittmar, 1992; Pierce et al., 2001).
-In an Australian donor context, we have provided initial support for the idea that psychological ownership experienced towards a blood collection agency both increases with past donation experience and is positively associated with blood donor self-identity (Edwards et al., 2023).
+As a novice donor continues to donate, this role-identity is validated, and thus blood donation becomes part of a donor’s self-identity (Being a blood donor is an important part of who I am; [@grubeRoleIdentityOrganizational2000,@masserPredictingBloodDonation2009].
+However, despite the link between donor self-identity and donor retention, there are currently no interventions which can foster a sense of identity in donors that do not rely exclusively on repeat donation behavior.
+Psychological ownership (this is mine) is the sense of ownership experienced over both material (e.g., blood collection agency) and abstract immaterial (e.g., blood donation behavior) possessions [@pierceTheoryPsychologicalOwnership2001].
+The ownership volunteers experience over a non-profit organisation has been linked to sustained volunteering behavior [@ainsworthFeelingsOwnershipVolunteering2020].
+Further, psychological ownership is theoretically associated with identity as it has been proposed that through possession identity is established and maintained [@dittmarSocialPsychologyMaterial1992, @pierceTheoryPsychologicalOwnership2001].
+In an Australian donor context, we have provided initial support for the idea that psychological ownership experienced towards a blood collection agency both increases with past donation experience and is positively associated with blood donor self-identity [@edwardsPsychologicalOwnershipIdentity2023].
 Further, in a series of semi-structured interviews we have shown that Australian blood donors primarily expressed ownership over the blood donation process (role behaviour), not the blood collection agency (organisation).
 As such, interventions that enhance the amount of psychological ownership donors experience over their donation have the potential to impact donor retention.
-In this study we try to repeat results from Edwards et al., 2023 in Finland by collecting similar questionnaire data and analysing them with a structural equation model.
+In this study we try to repeat results from Edwards et al., 2023 in Finland by collecting similar questionnaire data from donors of Finnish Red Cross Blood Service (FRCBS) and modeling the questionnaire data with a structural equation model.
 We include pre-registered R analysis code for direct repetition of Edwards et al., 2023 in Appendix 1.
@@
 
-Cross sectional survey study modelled with a structural equation model.
+Cross sectional survey study modeled with a structural equation model.
 
@@
 Number of life time donations (bd02) and number of donations in last years (bd03) and likert scale measured variables (po01-3, si01-03, it01-03, ac01-04, iv01-03) will be treated as numeric.
-Last donation date (bd04) will transformed "bd04 / 10^-(2 - str_length(as.numeric(max(bd04))))".
+Last donation date (bd04) will transformed "bd04 / 10^-(1 - str_length(as.numeric(max(bd04))))".
 
@@
 
-In supplement we will report results from all models, but as the main results we will report the model with highest Comparative Fit Index (CFI).
+In supplement we will report results from all models included in preregistered analysis (Appendix 1), but as the main results we will report the model with highest Comparative Fit Index (CFI).
+Among models with identical CFIs we will choose the model which gives most conservative i.e.
+lowest estimate of ind_eff =  a1* d21 * b2.
 95 % confidence or credible intervals will be used to decide if an effect is considered significant or not.
@@
 # References
-## 
+
+::: {#refs custom-style="Bibliography"}
+:::
+
 \vspace{-2pc}

### 2025-04-09 preregistration_code.Rmd
references

--- 2025-04-07 preregistration_code.Rmd
@@
 ---
-
-TODO:
-
- -remove superfluous outputs
-
- -check discussion with Lauri, was there anything else to do?
-
- -check model saving
-
- -fill out the preregistration form
-
- -are Abigail's sum score coefficients bootstrapped?
-
- -report effects of exclusions
 
@@
 #library(semPlot) # update plots to semPlot later
-#library(blavaan) # could not get it to work with this model and synthetic data
-#library(tidySEM) # https://github.com/yrosseel/lavaan/issues/359 lavaan does not work if tidySEM is loaded.
+#library(blavaan) # could not get it to work with this model and synthetic data, try as sensitivity analysis later
+#library(tidySEM) # https://github.com/yrosseel/lavaan/issues/359 lavaan does not work if tidySEM is loaded!
 ```
@@
 }
-# sometimes it get strange errors about C stack which can be solved with this
+# sometimes I get strange errors about C stack which can be solved with this
 #data <- as.data.frame(data)
@@
   fit.po02ss <- models$fit.po02ss
+  fit.po03ss <- models$fit.po03ss
   fit.po02 <- models$fit.po02
-  fit.po03ss <- models$fit.po03ss
   fit.po03 <- models$fit.po03
@@
 
-## Age by Gender
+## Age by sex
 
@@
 
-p <- ggplot(data)
+p <- ggplot(data %>% filter(de01 == 'man' | de01 == "women"))
 p <- p +  geom_histogram(aes(x=de02),binwidth = 5)
@@
 
+# Exclusions
+
+```{r}
+
+data %>%
+  filter(!(de01 == 'man' | de01 == "women"))  %>% nrow()
+cat("are removed because of sex\n")
+
+data %>%
+  filter(bd01 == 'no') %>% nrow()
+cat("are removed because of non-eligibility\n")
+
+data %>%
+  #lavaan cannot deal with unordered factor with more then 2 levels
+  filter(de01 == 'man' | de01 == "women") %>% 
+  filter(bd01 != 'no') %>% nrow()
+cat("in total after exclusions.\n")
+
+```
+
 # Model
@@
         bd04 = as.numeric(bd04), # lavaan cannot handle dates
-        bd04 = bd04 / 10^-(2 - str_length(as.numeric(max(bd04)))) # https://groups.google.com/g/lavaan/c/r7w-4HHg5R0 
+        bd04 = bd04 / 10^-(1 - str_length(as.numeric(max(bd04))))
+        #divide the numeric date which is in 10000s by 10000
+        # https://groups.google.com/g/lavaan/c/r7w-4HHg5R0 
         #"Your underweight variance seems to be much larger than other variances.  Try dividing it by 10 or 100 to make the SDs more similar across modeled variables." 
-        # But this does not seem to necessarily help.
         )
   ) %>% 
-  #Original paper used sum scores instead of CFA. Let's add them for comparison
+  #Original paper used sum scores instead of FA. 
   mutate(
@@
     ss_selfid = as.numeric(si01) + as.numeric(si02) + as.numeric(si03),
-    ss_intention = as.numeric(it01) + as.numeric(it02) + as.numeric(it03),
-    #donation history is on diffent scales, normalise them before summing.
-    #others should be positively correlated, but bd04 negatively
-    ss_blooddonor = scale(as.numeric(bd02)) 
-    + scale(as.numeric(bd01)) 
-    + scale(bd03)  
-    -scale(bd04) # do we want to include the DATE OF LAST DONATION?
-
+    ss_intention = as.numeric(it01) + as.numeric(it02) + as.numeric(it03)
     )
@@
                 iv =~ iv01 + iv02 + iv03
-                bd =~ bd02 + bd03 + bd04 #bd02 is sure correlate positively
+                bd =~ bd02 + bd03 + bd04 #bd02 is sure to correlate positively
 
@@
 fit.po01.cfa <- cfa(po01.model,  
-                    estimator = "WLSMV", #
+                    estimator = "MLR", #
 
-                    data = data.fit)
+                    data = data.fit2)
 fit.po01.cfa
@@
 ```
+
+### CFA metrics
 
@@
   filter(op == '=~')
-# with synthetic data bd04 should come up
-# as they are just random as can be seen from the corplot
 ```
@@
 
-#cfa should be better than cfa.o in all measures. cfi and chisq should be bigger, rmsea and srmr smaller in model cfa
+#cfa should be better than cfa.o in all measures. 
+#cfi and chisq should be bigger, rmsea and srmr smaller in model cfa
 ```
@@
   fit.po02ss <- sem(po02ss.model,  
-                    estimator = "MLR", # robust diagonally weighted least squares stable and should work with ordered factors
+                    estimator = "MLR", 
+  # robust diagonally weighted least squares stable and should work with ordered factors
                     data = data.fit2,
-                auto.fix.first = FALSE, #  If TRUE, the factor loading of the first indicator is set to 1.0 for every latent variable. 
+                    auto.fix.first = FALSE, 
+  #  If TRUE, the factor loading of the first indicator is set to 1.0 for every latent variable. 
                     std.lv = TRUE, 
@@
 ```{r}
-varTable(fit.po02ss)
-```
-
-```{r}
 lavaan::inspect(fit.po02ss,"r2")
@@
 ```{r}
-lavaan::standardizedsolution(fit.po02ss) %>% select(-label)
-# Use standardizedSolution to obtain SEs and test statistics for standardized estimates.
-
-```
-
-```{r}
-lavaan::standardizedsolution(fit.po02ss) %>% select(-label,se,z) %>%  filter(op == '~')
+fit.po02ss.est  <-standardizedSolution(fit.po02ss)
+fit.po02ss.est %>% select(-label,se,z) %>%  filter(op == '~')
 ```
@@
 lavaanPlot(
-  model=fit.po02ss,  
-  coefs = TRUE,
-  node_options = list(shape = "box", fontname = "Helvetica"), 
-  edge_options = list(color = "grey"), 
-  labels = labelsforplottingnull,
-
-  #In simulated data there should be correlations inside
-  # bloodonor, psycown, selfid
-  # and between
-  # bloodonor, psycown
-  # and nothing else
-  sig = 0.001
-  # if sig is < 0.001 one gets significant correlations
-  # also from simulated data elsewhere suggesting that 0.05 is not enough
-
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem01.pdf"))
-```
-
-```{r}
-# Same with standardised coefficients
-lavaanPlot(
   model = fit.po02ss, 
@@
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
-```
-
-```{r}
-#semPlot::semPaths(fit.po02ss)
 ```
@@
 ```{r}
+fit.po03ss.best  <-standardizedSolution(fit.po03ss)
+fit.po03ss.best %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+fit.po03ss.best %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
+  # Just the loadings
+  filter(op != '~1')
+
+```
+
+```{r}
 lavaanPlot(
@@
 
-```
-
-```{r}
-fit.po03ss.best <- parameterEstimates(fit.po03ss)
-fit.po03ss.best %>% select(-se,-z) %>% filter(op == "~")
 ```
@@
   fit.po02 <- sem(po02.model,  
-                    estimator = "MLR", # robust diagonally weighted least squares stable and should work with ordered factors
+                  estimator = "MLR",
+                  # robust diagonally weighted least squares stable and should work with ordered factors
                   data = data.fit2,
@@
 ```{r}
-varTable(fit.po02)
-```
-
-```{r}
 lavaan::inspect(fit.po02,"r2")
@@
 ```{r}
-lavaan::standardizedsolution(fit.po02) %>% select(-label)
-# Use standardizedSolution to obtain SEs and test statistics for standardized estimates.
-
-```
-
-```{r}
-lavaan::standardizedsolution(fit.po02) %>% select(-label,se,z) %>%  filter(op == '~')
+fit.po02.est  <-standardizedSolution(fit.po02)
+fit.po02.est %>% select(-label,se,z) %>%  filter(op == '~')
 ```
@@
 
-fit.po02.est  <-standardizedSolution(fit.po02)
 fit.po02.est %>% 
@@
 
-lavaanPlot(
-  model=fit.po02,  
-  coefs = TRUE,
-  node_options = list(shape = "box", fontname = "Helvetica"), 
-  edge_options = list(color = "grey"), 
-  labels = labelsforplotting,
-
-  #In simulated data there should be correlations inside
-  # bloodonor, psycown, selfid
-  # and between
-  # bloodonor, psycown
-  # and nothing else
-  sig = 0.001
-  # if sig is < 0.001 one gets significant correlations
-  # also from simulated data elsewhere suggesting that 0.05 is not enough
-
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem01.pdf"))
-```
-
-```{r}
-# Same with standardised coefficients
 lavaanPlot(
@@
   fit.po03 <- sem(po02.model,  
-                    estimator = "ML", # bootstrapping does not work with estimators designed for factors
-                    data = data.fit2, # data that is completely numeric
+                  estimator = "ML", 
+                  # bootstrapping does not work with estimators designed for factors
+                  data = data.fit2,
+                  # data that is completely numeric
                   bootstrap = nBoots,
@@
 
+```
+
+```{r}
+fit.po03.best <- parameterEstimates(fit.po03)
+fit.po03.best %>% select(-se,-z) %>% filter(op == "~")
 ```
@@
 
-```{r}
-fit.po03.best <- parameterEstimates(fit.po03)
-fit.po03.best %>% select(-se,-z) %>% filter(op == "~")
-```
-
 ## Compare models
@@
   mutate(estmet = "Factor analysis\nbootstrapped") %>%  
+  #rename(est=est.std) %>% 
   filter(op == "~" | op == ":=") %>% 
@@
 data.forest  <- fit.po03ss.best %>% 
+  rename(est=est.std) %>% 
   mutate(estmet = "Sum score\nbootstrapped") %>%  
@@
     term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
+    term =str_replace_all(term,"bd02","blooddonor:"),
+    term =str_replace_all(term,"blooddonor:","blooddonor"),
+    term =str_replace_all(term,"^:",""),
     estmet = as.factor(estmet),

### 2025-04-14 preregistration.Rmd
time and folder

@@
 shorttitle      : "Donation ownership"
-date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
+date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time())`"
 
@@
 
-csl             : vox-sanguinis.csl
+csl             : manuscript/vox-sanguinis.csl
 bibliography      : ["manuscript/references.bib", "manuscript/r-references.bib"]

### 2025-04-14 preregistration_code.Rmd
time and folder

@@
 author: "Abigail Edwards, Lauri Hietajärvi and Mikko Arvas"
-date: "`r Sys.time()`"
+date: "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time())`"
 output:
@@
 }
+
+dir.create(here("figures"))
 

### 2025-04-14 preregistration.Rmd
clean table

@@
 <table>
+
 <tr>
-<td></td>
 <td>Psychological ownership</td>
@@
 <tr>
-<td>
-po01 1. I feel a very high degree of personal ownership of donating blood. [Likert]
-</td>
+<td>po01 1. I feel a very high degree of personal ownership of donating blood. [Likert]</td>
 </tr>
-<td>
-po02 2. I feel like donating blood belongs to me. [Likert]
-</td> </tr>
+
+<tr> 
+<td>po02 2. I feel like donating blood belongs to me. [Likert] </td> 
+</tr>
+
+<tr>
+<td>po03 3. I feel like I own donating blood. [Likert]</td> 
+</tr>
+
+<tr> 
+<td>Self-identity</td> 
+</tr>
+
+<tr>
+<td>si01 1. Donating blood is important to me. [Likert]</td> 
+<tr>si02 2. I am like the kind of person who donates blood. [Likert]</td> 
+</tr>
+
+<tr> 
+<td>si03 3. Donating blood is an important part of who I am. [Likert]</td> 
+</tr>
+
+<tr>
+<td>Intention</td>
+</tr>
+
+<tr>
+<td>it01 1. I would like to donate blood in the future. [Likert]</td>   
+</tr>
+
+<tr>
+<td>it02 2. I intend to donate blood in the future. [Likert]</td>
+</tr>
+
+<tr>
+<td>it03 3. I intend to make more than just a one-off blood donation. [Likert]</td>
+</tr>
+
 <tr>
 <td>
-po03 3. I feel like I own donating blood. [Likert]
-</td> </tr>
-<tr> <td></td>
+Intention to donate for emergency</td>
+</tr>
 
-<td>
-Self-identity
-</td> </tr>
 <tr>
-<td>
-si01 1. Donating blood is important to me. [Likert]
-</td> <tr>
-si02 2. I am like the kind of person who donates blood. [Likert]
-</td> </tr>
+<td>ie01 1. I would be willing to donate blood if I received an urgent callout from the Finnish Red Cross Blood Service. [Likert]</td>
+</tr>
+
 <tr>
-<td>
-si03 3. Donating blood is an important part of who I am. [Likert]
-</td> </tr>
-<tr> <td></td>
+<td>Antecedents of psychological ownership</td>
+</tr>
 
-<td>
-Intention
-</td></tr>
 <tr>
-<td>
-it01 1. I would like to donate blood in the future. [Likert]
-</td>   </tr>
+<td>Control</td>
+</tr>
+
 <tr>
-<td>
-it02 2. I intend to donate blood in the future. [Likert]
-</td> </tr>
+<td>ac01  1. I have influence over the things that affect me while donating blood. [Likert]</td>
+</tr>
+
 <tr>
-<td>
-it03 3. I intend to make more than just a one-off blood donation. [Likert]
-</td></tr>
-<tr><td></td>
-<td>
-Intention to donate for emergency
-</td></tr>
+<td>ac02 2. I control the location and scheduling for donating blood. [Likert]</td>
+</tr>
 <tr>
-<td>
-ie01 1. I would be willing to donate blood if I received an urgent callout from the Finnish Red Cross Blood Service. [Likert]
-</td></tr>
-<tr><td></td> 
-<td>
-Antecedents of psychological ownership
-</td></tr>
-<tr><td></td>
 
-<td>
-Control
-</td></tr>
+<td>ac03 3. I influence decisions related to my blood donation. [Likert]</td>
+</tr>
+
 <tr>
-<td>
-ac01  1. I have influence over the things that affect me while donating blood. [Likert]
-</td></tr>
+<td>ac04 4. In general, I have control over donating blood. [Likert]</td>
+</tr>
 <tr>
-<td>
-ac02 2. I control the location and scheduling for donating blood. [Likert]
-</td></tr>
+<td>Intimate knowledge</td>
+</tr>
+
 <tr>
-<td>
-ac03 3. I influence decisions related to my blood donation. [Likert]
-</td></tr>
+<td>ai01 1. I am intimately familiar with what is going on with regard to donating blood. [Likert]</td>
+</tr>
+
 <tr>
-<td>
-ac04 4. In general, I have control over donating blood. [Likert]
-</td></tr>
-<tr><td></td>
-<td>
-Intimate knowledge
-</td></tr>
+<td>ai02 2. I have a depth of knowledge as it relates to donating blood. [Likert]</td>
+</tr>
+
 <tr>
-<td>
-ai01 1. I am intimately familiar with what is going on with regard to donating blood. [Likert]
-</td></tr>
+<td>ai03 3. I have a comprehensive understanding of donating blood. [Likert]</td>
+</tr>
+
 <tr>
-<td>
-ai02 2. I have a depth of knowledge as it relates to donating blood. [Likert]
-</td></tr>
+<td>ai04 4. I have a broad understanding of donating blood. [Likert]</td>
+</tr>
+
 <tr>
-<td>
-ai03 3. I have a comprehensive understanding of donating blood. [Likert]
-</td></tr>
+<td>Self-Investment</td>
+</tr>
+
 <tr>
-<td>
-ai04 4. I have a broad understanding of donating blood. [Likert]
-</td></tr>
-<tr><td></td>
+<td>iv01 1.	I have invested a major part of myself into donating blood. [Likert]</td>
+</tr>
 
-<td>
-Self-Investment
-</td></tr>
-<td>
-iv01 1.	I have invested a major part of myself into donating blood. [Likert]
-</td></tr>
 <tr>
-<td>
-iv02 2. I have invested a significant amount of my time into donating blood. [Likert]
-</td></tr>
+<td>iv02 2. I have invested a significant amount of my time into donating blood. [Likert] </td>
+</tr>
+
 <tr>
-<td>
-iv03 3. Overall, I invested a lot into donating blood. [Likert]
-</td></tr>
-<tr><td></td>
+<td>iv03 3. Overall, I invested a lot into donating blood. [Likert]</td>
+</tr>
 
-<td>
-Blood donation history
-</td></tr>
 <tr>
-<td>
-bd01 1. Do you believe that you are currently eligible to donate blood [yes, no, unsure]
-</td></tr>
+<td>Blood donation history</td>
+</tr>
+
 <tr>
-<td>
-bd02 2. How many times in total have you donated blood at the Finnish Red Cross Blood Service Finland? [select number]
-</td></tr>
+<td>bd01 1. Do you believe that you are currently eligible to donate blood [yes, no, unsure]</td>
+</tr>
+
 <tr>
-<td>
-bd03 3. How many times in the last two years have you donated at the Finnish Red Cross Blood Service? [select number]
-</td></tr>
+<td>bd02 2. How many times in total have you donated blood at the Finnish Red Cross Blood Service Finland? [select number]</td>
+</tr>
+
 <tr>
-<td>
-bd04 4. When did you last attend a Finnish Red Cross blood donation centre with the intention of donating blood? [select date]
-</td></tr>
-<tr><td></td>
+<td>bd03 3. How many times in the last two years have you donated at the Finnish Red Cross Blood Service? [select number]</td>
+</tr>
 
-<td>
-Demographics
-</td></tr>
 <tr>
-<td>
-de01 1. What is your gender? [man, woman, non-binary, prefer to not disclose, prefer to self-describe (open response)]
-</td></tr>
+<td>bd04 4. When did you last attend a Finnish Red Cross blood donation centre with the intention of donating blood? [select date]</td>
+</tr>
+
 <tr>
-<td>
-de02 2. How old are you today? [select number]
-</td></tr>
+<td>Demographics</td>
+</tr>
+
+<tr>
+<td>de01 1. What is your gender? [man, woman, non-binary, prefer to not disclose, prefer to self-describe (open response)]</td>
+</tr>
+
+<tr>
+<td>de02 2. How old are you today? [select number]</td>
+</tr>
 </table>

### 2025-04-14 preregistration.Rmd
no main model

@@
 
-In supplement we will report results from all models included in preregistered analysis (Appendix 1), but as the main results we will report the model with highest Comparative Fit Index (CFI).
-Among models with identical CFIs we will choose the model which gives most conservative i.e.
-lowest estimate of ind_eff =  a1* d21 * b2.
+Comparative results from these models will presented and their merits discussed.
 95 % confidence or credible intervals will be used to decide if an effect is considered significant or not.

### 2025-04-22 codebook_data_simulated.Rmd
time stampo of answering to synthetic

@@
 ---
-title: "Codebook created on 2025-04-03 at 2025-04-03 10:38:35"
+title: "Codebook created on 2025-04-22 at 2025-04-22 13:34:49"
 output: github_document
@@
 ## Dataset description
-The data contains 1000 cases and 27 variables.
+The data contains 1000 cases and 28 variables.
 

### 2025-04-22 prepare_data.R

@@ -3,6 +3,8 @@
 # use the function open_data(your_data_frame) or closed_data(your_data_frame)
 # to store the data.
 
+
+#To update data for example with new synthetic data just repalce "simulated_data.csv" and source this file.
 library(worcs)
 library(here)
 library(tidyverse)

### 2025-04-22 prepare_data.R
formatting for missing data

@@ -48,9 +48,17 @@ data_simulated <- data_simulated %>%
       qabbr[1:21]),
     ~ f1(.) 
   ) %>% 
-  mutate(bd04 = as.Date(bd04),
-         bd01 = factor(bd01, levels = c("no","unsure","yes"),ordered = TRUE),
-         de01 = factor(de01)
+  mutate(
+      dttma = as.POSIXct(dttma),
+      bd01 = factor(bd01, levels = c("no","unsure","yes"),ordered = TRUE),
+      bd02 = as.numeric(bd02),
+      bd03 = as.numeric(bd03),
+      bd04 = as.Date(bd04),
+      de01 = factor(de01),
+      de02 = as.numeric(de02),
+      #Days between answering and last donation
+      bd04 = as.numeric(as.Date(dttma) - bd04)
+      
   )  
 
 #####################################################################################################################################

### 2025-04-22 prepare_data.R
type

@@ -4,7 +4,7 @@
 # to store the data.
 
 
-#To update data for example with new synthetic data just repalce "simulated_data.csv" and source this file.
+#To update data for example with new synthetic data just replace "simulated_data.csv" and source this file.
 library(worcs)
 library(here)
 library(tidyverse)

### 2025-04-22 preregistration_code.Rmd
clean table
imputation

@@
 library(kableExtra)
+library(naniar)
 #library(semPlot) # update plots to semPlot later
@@
 #library(tidySEM) # https://github.com/yrosseel/lavaan/issues/359 lavaan does not work if tidySEM is loaded!
+
+# guide on how to setup worcs to reproduce this analysis https://cjvanlissa.github.io/worcs/articles/reproduce.html
 ```
@@
 #data <- as.data.frame(data)
-fitmodels <- TRUE
+fitmodels <- FALSE
 if (!fitmodels) {
   print("Loading previously fitted models")
-  load(file=here(str_replace(str_c("results/",datafile),"\\.data\\.",".models.")))
+  load(file=here("prereg_models.rdata"))
   fit.po02ss <- models$fit.po02ss
   fit.po03ss <- models$fit.po03ss
+  fit.po04ss <- models$fit.po04ss
   fit.po02 <- models$fit.po02
   fit.po03 <- models$fit.po03
+  fit.po04 <- models$fit.po04
 }
@@
       DonationCountLast2years = bd03,
+      DaysFromLastDonation = bd04 
     ),
@@
 
-## Confirmatory factor analysis
+## Format data for modelling
 
 ```{r}
-models <- list()
-
 isordered <- data %>%  summarise(across(everything(), ~ is.ordered(.x))) %>%  as.logical() %>% which()
@@
         bd04 = as.numeric(bd04), # lavaan cannot handle dates
-        bd04 = bd04 / 10^-(1 - str_length(as.numeric(max(bd04))))
-        #divide the numeric date which is in 10000s by 10000
-        # https://groups.google.com/g/lavaan/c/r7w-4HHg5R0 
-        #"Your underweight variance seems to be much larger than other variances.  Try dividing it by 10 or 100 to make the SDs more similar across modeled variables." 
         )
@@
 summary(data.fit)
+
 ```
@@
 ```{r}
-corrplot(cor(data.fit2))
+corrplot(cor(data.fit2,use="pairwise.complete.obs"))
 ```
 
+## Missing data analysis
+
+Check proportion of missing in all variables
+
 ```{r}
+gg_miss_var(
+  data.fit2
+  )
+```
+
+Check major combinations of missingmess.
+
+```{r}
+gg_miss_upset(data.fit2,nsets = 10,nintersects=40)
+```
+
+Check is missingness related to data accumulation.
+
+```{r}
+
+gg_miss_case_cumsum(
+  data.fit2 %>% arrange(dttma)
+)
+
+```
+
+Check is missingness related to major endpoint of the study i.e. intention to donate blood.
+
+```{r}
+
+gg_miss_fct(
+  x = data.fit %>% select(-dttma,
+                          -de01,
+                          -de02,
+                          -bd02,
+                          -bd04,
+                          -ss_psycown,
+                          -ss_selfid,
+                          ), 
+  fct = ss_intention)
+
+```
+
+Carry out a crude overall test of missingness.
+
+```{r}
+mcar_test(data.fit2 %>% select(-dttma))
+# p < 0.05 -> Reject the null hypothesis-> Data is MAR (Missing at random) or MNAR (Missing not at random), meaning missingness follows a pattern.
+# But this assumes normally distributed data and does not prove that data is MCAR (Missing Completely at Random)
+```
+
+## Confirmatory factor analysis
+
+```{r}
+models <- list()
+
 po01.model <- ' po =~ po01 + po02 + po03
@@
                     estimator = "MLR", #
-
+                     missing = "listwise",
                     data = data.fit2)
@@
 ```{r}
-#Make a model with out prespesified correlation structure
+#Make a model with out prespecified correlation structure
 fit.po01.cfa.o <- cfa(po01.model,data.fit,orthogonal=TRUE)
@@
                 auto.fix.first = FALSE,
-                std.lv = TRUE
+                std.lv = TRUE,
+                missing = "listwise"
                 )
@@
 
-### c) Latent factor model with robust estimator
+###  c) Sum score model with missing data imputation
+
+```{r}
+
+if (fitmodels) {
+  fit.po04ss <- sem(
+    po02ss.model,  
+    estimator = "ML", # 
+    data = data.fit2, # data that is completely numeric
+    auto.fix.first = FALSE,
+    std.lv = TRUE,
+    missing = "fiml",
+    fixed.x = F 
+  )
+  models[['fit.po04ss']] <- fit.po04ss
+}
+
+lavaan::summary(fit.po04ss,standardized = TRUE)
+
+```
+
+```{r}
+fit.po04ss.est  <-standardizedSolution(fit.po04ss)
+fit.po04ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+fit.po04ss.est %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
+  # Just the loadings
+  filter(op != '~1')
+
+```
+
+```{r}
+lavaanPlot(
+  model = fit.po04ss, 
+   labels = labelsforplottingnull,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))
+
+```
+
+### d) Latent factor model with robust estimator
 
@@
 
-### d) Latent factor model with boostrapping
+### e) Latent factor model with boostrapping
 
@@
 
+###  f) Latent factor model with missing data imputation
+
+```{r}
+
+if (fitmodels) {
+  fit.po04 <- sem(
+    po02.model,  
+    estimator = "ML", # 
+    data = data.fit2, # data that is completely numeric
+    auto.fix.first = FALSE,
+    std.lv = TRUE,
+    missing = "fiml",
+    fixed.x = F 
+  )
+  models[['fit.po04']] <- fit.po04
+}
+
+lavaan::summary(fit.po04,standardized = TRUE)
+
+```
+
+```{r}
+fit.po04.est  <-standardizedSolution(fit.po04)
+fit.po04.est %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+fit.po04.est %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
+  # Just the loadings
+  filter(op != '~1')
+
+```
+
+```{r}
+lavaanPlot(
+  model = fit.po04, 
+   labels = labelsforplottingnull,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))
+
+```
+
 ## Compare models
@@
     fit.po02.est %>% mutate(estmet = "Factor analysis") %>% 
+      rename(est=est.std) %>% 
+      filter(op == "~"| op == ":="),
+    fit.po04.est %>% mutate(estmet = "Factor analysis\nimputed") %>% 
       rename(est=est.std) %>% 
@@
       rename(est=est.std) %>% 
+      filter(op == "~"| op == ":=") ,
+    fit.po04ss.est %>% mutate(estmet = "Sum score\nimputed") %>% 
+      rename(est=est.std) %>% 
       filter(op == "~"| op == ":=")
+
   ) %>% 
@@
 
-chisq, Chi-squared, larger is better
+chisq, Chi-squared, chisq/df ratio smaller is better
 

### 2025-04-23 preregistration_code.Rmd
missing fixes

@@
 
+# renv::status() to check 
+# renv::snapshot() to add used and install to renv
+
 # guide on how to setup worcs to reproduce this analysis https://cjvanlissa.github.io/worcs/articles/reproduce.html
@@
 #data <- as.data.frame(data)
-fitmodels <- FALSE
+fitmodels <- FALSE # TRUE for fitting models
 if (!fitmodels) {
@@
 
-Check is missingness related to data accumulation.
+Check if missingness is related to data accumulation.
 
@@
 
-Check is missingness related to major endpoint of the study i.e. intention to donate blood.
+Check if missingness is related to major endpoint of the study i.e. intention to donate blood.
 
@@
     missing = "fiml",
-    fixed.x = F 
+    fixed.x = FALSE 
   )
@@
   model = fit.po04, 
-   labels = labelsforplottingnull,
+   labels = labelsforplotting,
   node_options = list(shape = "box", fontname = "Helvetica"), 
@@
 
+### Effect sizes
+
 ```{r}
+
 data.forest  <- fit.po03.best %>% 
-  mutate(estmet = "Factor analysis\nbootstrapped") %>%  
+  mutate(estmet = "Bootstrapped"
+         ) %>%  
   #rename(est=est.std) %>% 
@@
   bind_rows(
-    fit.po02.est %>% mutate(estmet = "Factor analysis") %>% 
+    fit.po02.est %>% mutate(estmet = "Robust") %>% 
       rename(est=est.std) %>% 
       filter(op == "~"| op == ":="),
-    fit.po04.est %>% mutate(estmet = "Factor analysis\nimputed") %>% 
+    fit.po04.est %>% mutate(estmet = "Imputed") %>% 
       rename(est=est.std) %>% 
@@
   mutate(
-    term = str_c(label,":",lhs,op,rhs)
+    term = str_c(label,":",lhs,op,rhs),
+    linetype = "Latent factor"
     ) %>% 
@@
   rename(est=est.std) %>% 
-  mutate(estmet = "Sum score\nbootstrapped") %>%  
+  mutate(estmet = "Bootstrapped") %>%  
   filter(op == "~"| op == ":=") %>% 
   bind_rows(
-    fit.po02ss.est %>% mutate(estmet = "Sum score") %>% 
+    fit.po02ss.est %>% mutate(estmet = "Robust") %>% 
       rename(est=est.std) %>% 
       filter(op == "~"| op == ":=") ,
-    fit.po04ss.est %>% mutate(estmet = "Sum score\nimputed") %>% 
+    fit.po04ss.est %>% mutate(estmet = "Imputed") %>% 
       rename(est=est.std) %>% 
@@
   mutate(
-    term = str_c(label,":",lhs,op,rhs)
+    term = str_c(label,":",lhs,op,rhs),
+    linetype = "Sum score"
     ) %>% 
@@
                              levels=levels(estmet))
+  ) %>% 
+  mutate(
+    mtype = str_c(estmet,"-",linetype)
   )
 
-pos <- position_nudge(y=as.numeric(as.factor(data.forest$estmet))/5 -0.5)
+pos <- position_nudge(y=as.numeric(as.factor(data.forest$mtype))/7 -0.5)
 
-data.forest %>% 
-  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=estmet))+
+pf <- data.forest %>% 
+  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=estmet,linetype = linetype))+
   geom_vline(xintercept=0,color="grey") +
   geom_linerange(position=pos)+
-  geom_point(shape=21,size=3,position=pos)+
+  geom_point(shape=21,size=3,position=pos,alpha=0.7)+
   scale_fill_discrete(na.value=NA, guide ="none") + 
   labs(x="Standardised effect size",y="Relationship") +
-  labs(color="Model") +
+  labs(color="Model type", linetype="Variable type") +
   geom_stripped_rows(color=NA) + theme_classic()
+
+pf
+```
+
+```{r}
+ggsave(pf,
+       filename = here("figures/forest01.png"),
+       height = 140,
+       width = 180,
+       units = "mm"
+         )
+
+ggsave(pf,
+       filename = here("figures/forest01.pdf"),
+       height = 140,
+       width = 180,
+       units = "mm"
+         )
 
@@
 
+## Indices of goodness of fit
+
 ```{r}
 
-round(
-  sapply(list(SS=fit.po02ss,SSB=fit.po03ss,FA=fit.po02,FAB=fit.po03 ),
-         function(x) fitmeasures(x,c("chisq","df","cfi","rmsea","srmr"))
+gfi<- round(
+  sapply(
+    list(
+      SS=fit.po02ss,
+      SSB=fit.po03ss,
+      SSI = fit.po04ss,
+      FA=fit.po02,
+      FAB=fit.po03, 
+      FAI=fit.po04
+      ),
+         function(x) fitmeasures(x,c("chisq","df","pvalue","cfi","rmsea","srmr"))
          )
@@
 
+gfi
 ```
@@
 
-srmr Standardized Root Mean Square Residual, "Should be < .08", smaller is better
+srmr Standardized Root Mean Square Residual, "should be < .08", smaller is better
+
+NB. Sum score and latent factor models use different data and are not nested hence they cannot be compared with these metrics.
 

### 2025-04-23 preregistration.Rmd
missing data

@@
 Number of life time donations (bd02) and number of donations in last years (bd03) and likert scale measured variables (po01-3, si01-03, it01-03, ac01-04, iv01-03) will be treated as numeric.
-Last donation date (bd04) will transformed "bd04 / 10^-(1 - str_length(as.numeric(max(bd04))))".
+Last donation date (bd04) will transformed to days from last donation by subtracting date of questionnaire response from date of last donation.
 
@@
 
-c. Latent factor model with robust estimator
+c. Sum score model with FIML for missing data
 
-d. Latent factor model with boostrapping
+d. Latent factor model with robust estimator
+
+e. Latent factor model with boostrapping
+
+f. Latent factor model with FIML for missing data
 
@@
 
-The survey platform only reports completed questionnaires and all questions are obligatory. Hence there will be no missing data.
+Missing data will be analysed (see Appendix 1 for details) for possible biases and Little’s MCAR test carried out to assess randomness of missing data.
+Models using robust maximum likelihood or bootstrapped maximum likelihood estimation will be fitted with "listwise" deletion of missing data and in addition models using maximum likelihood estimator and "full information maximum likelihood" approach to deal with missing data will be fitted.
+All these results will be reported and discussed, acknowledging that there is no well-accepted method of dealing with data that is not missing at random
 
@@
 Bivariate correlations will be calculated to compare donors’ willingness to respond to a call-out to past blood donation behaviour, psychological ownership, self-identity, and intention to donate blood.
+If there is sufficient data we will stratify models by sex.
+Model improvements will be considered if deemed appropriate.
 

### 2025-04-23 preregistration_code.Rmd
formatting

@@
   tbl_summary(
-    data %>% 
+    data = data %>% 
       rename(
@@
       DonationEligibility,
+      DaysFromLastDonation,
       DonationCount,
@@
       ),
-    by =  Gender
+    by =  Gender,
+     statistic = list(all_continuous() ~ "{median} ({p25}, {p75})", all_categorical() ~
+    "{n} ({p}%)")
 
@@
 mcar_test(data.fit2 %>% select(-dttma))
-# p < 0.05 -> Reject the null hypothesis-> Data is MAR (Missing at random) or MNAR (Missing not at random), meaning missingness follows a pattern.
-# But this assumes normally distributed data and does not prove that data is MCAR (Missing Completely at Random)
+# p < 0.05 -> Reject the null hypothesis-> Data is MAR (Missing at random) 
+# or MNAR (Missing not at random), meaning missingness follows a pattern.
+# But this assumes normally distributed data and 
+# does not prove that data is MCAR (Missing Completely at Random)
 ```
@@
   sapply(list(cfa=fit.po01.cfa,cfa.o=fit.po01.cfa.o),
-         function(x) fitmeasures(x,c("npar","chisq","df","cfi","rmsea","srmr"))
+         function(x) fitmeasures(x,c("npar","chisq","df","pvalue","cfi","rmsea","srmr"))
          )
@@
 #cfa should be better than cfa.o in all measures. 
-#cfi and chisq should be bigger, rmsea and srmr smaller in model cfa
+#cfi and chisq/df should be bigger, rmsea and srmr smaller in model cfa
 ```

### 2025-04-24 preregistration.Rmd
formatting
8 models

@@
 An e-mail inviting to fill an online questionnaire will be sent to the randomly selected blood donors.
-No recompensation will be offered to participants.Participants must be between 18 and 72 years old.
+No recompensation will be offered to participants.
+Participants must be between 18 and 72 years old.
 
@@
 
-The online questionnaire will be open for 2 weeks i.e. from 5.5.2025 - to 18.5.2025.
+The online questionnaire will be open for 2 weeks i.e. from 5.5.2025 to 18.5.2025.
 
@@
 
-a. Sum score model with robust estimator
+a) Sum score model with robust estimator and listwise deletion
 
-b. Sum score model with bootstrapping
+b) Sum score model with bootstrapping and listwise deletion
 
-c. Sum score model with FIML for missing data
+c) Sum score model with robust estimator and fiml
 
-d. Latent factor model with robust estimator
+d) Sum score model with bootstrapping and fiml
 
-e. Latent factor model with boostrapping
+e) Latent factor model with robust estimator and listwise deletion
 
-f. Latent factor model with FIML for missing data
+f) Latent factor model with boostrapping and listwise deletion
+
+g) Latent factor model with robust estimator and fiml
+
+h) Latent factor model with bootstrapping and fiml
 
@@
 
-Missing data will be analysed (see Appendix 1 for details) for possible biases and Little’s MCAR test carried out to assess randomness of missing data.
-Models using robust maximum likelihood or bootstrapped maximum likelihood estimation will be fitted with "listwise" deletion of missing data and in addition models using maximum likelihood estimator and "full information maximum likelihood" approach to deal with missing data will be fitted.
+Missing data will be analysed for possible biases and Little’s MCAR test carried out to assess randomness of missing data.
+Models using "full information maximum likelihood" approach or listwise deletion will be fitted to inspect the effect of these modelling choices (see Appendix 1 for details).
 All these results will be reported and discussed, acknowledging that there is no well-accepted method of dealing with data that is not missing at random

### 2025-04-24 preregistration_code.Rmd
8 models

@@
 #data <- as.data.frame(data)
+#fitmodels <- TRUE # TRUE for fitting models
 fitmodels <- FALSE # TRUE for fitting models
@@
   fit.po04ss <- models$fit.po04ss
+  fit.po05ss <- models$fit.po05ss
   fit.po02 <- models$fit.po02
@@
   fit.po04 <- models$fit.po04
+  fit.po05 <- models$fit.po05
+} else{
+  models <- list()
 }
@@
   mutate(
-    ss_psycown = as.numeric(po01) + as.numeric(po02) + as.numeric(po03),
-    ss_selfid = as.numeric(si01) + as.numeric(si02) + as.numeric(si03),
-    ss_intention = as.numeric(it01) + as.numeric(it02) + as.numeric(it03)
+    ss_psycown = scale(as.numeric(po01) + as.numeric(po02) + as.numeric(po03)),
+    ss_selfid = scale(as.numeric(si01) + as.numeric(si02) + as.numeric(si03)),
+    ss_intention = scale(as.numeric(it01) + as.numeric(it02) + as.numeric(it03))
     )
@@
 ```{r}
-models <- list()
 
@@
 
-## SEM
-
-### a) Sum score model with robust estimator
+## Sum score SEM
 
@@
 
+### a) Sum score model with robust estimator and listwise deletion
+
 ```{r}
@@
                     estimator = "MLR", 
-  # robust diagonally weighted least squares stable and should work with ordered factors
+                    missing = "listwise",
                     data = data.fit2,
                     auto.fix.first = FALSE, 
-  #  If TRUE, the factor loading of the first indicator is set to 1.0 for every latent variable. 
                     std.lv = TRUE,
+                    fixed.x = TRUE
+
   )
@@
 
-###  b) Sum score model with bootstrapping
+###  b) Sum score model with bootstrapping and listwise deletion
 
@@
 if (fitmodels) {
-#fit
-
-# https://groups.google.com/g/lavaan/c/A6KCjXAZl-Q good example code
-
   fit.po03ss <- sem(po02ss.model,  
-                    estimator = "ML", # 
-                    data = data.fit2, # data that is completely numeric
+                    estimator = "ML",  
+                    missing = "listwise",
+                    data = data.fit2, 
                     bootstrap = nBoots,
@@
                     std.lv = TRUE,
-                missing = "listwise"
+                    fixed.x = TRUE
   )
@@
 
-###  c) Sum score model with missing data imputation
+###  c) Sum score model with robust estimator and fiml
 
@@
     po02ss.model,  
-    estimator = "ML", # 
-    data = data.fit2, # data that is completely numeric
+    estimator = "MLR", 
+    data = data.fit2, 
     auto.fix.first = FALSE,
@@
 
-### d) Latent factor model with robust estimator
+###  d) Sum score model with bootstrapping and fiml
+
+```{r}
+
+if (fitmodels) {
+  fit.po05ss <- sem(
+    po02ss.model,  
+    estimator = "ML", 
+    bootstrap = nBoots,
+    se = "boot",
+    data = data.fit2, 
+    auto.fix.first = FALSE,
+    std.lv = TRUE,
+    missing = "fiml",
+    fixed.x = FALSE 
+  )
+  models[['fit.po05ss']] <- fit.po05ss
+}
+
+lavaan::summary(fit.po05ss,standardized = TRUE)
+
+```
+
+```{r}
+fit.po05ss.est  <-standardizedSolution(fit.po05ss)
+fit.po05ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+fit.po05ss.est %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
+  # Just the loadings
+  filter(op != '~1')
+
+```
+
+```{r}
+lavaanPlot(
+  model = fit.po05ss, 
+   labels = labelsforplottingnull,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem05.pdf"))
+
+```
+
+## Latent factor SEM
 
@@
 
+### e) Latent factor model with robust estimator and listwise deletion
+
 ```{r}
@@
                   estimator = "MLR",
-                  # robust diagonally weighted least squares stable and should work with ordered factors
+                  missing = "listwise",
                   data = data.fit2,
                   auto.fix.first = FALSE,
-                  std.lv = TRUE 
+                  std.lv = TRUE,
+                  fixed.x = TRUE
   )
@@
   sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllssem02.pdf"))
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))
 ```
 
-### e) Latent factor model with boostrapping
+### f) Latent factor model with boostrapping and listwise deletion
 
@@
 if (fitmodels) {
-  #fit
-
-  # https://groups.google.com/g/lavaan/c/A6KCjXAZl-Q good example code
-
   fit.po03 <- sem(po02.model,  
                   estimator = "ML", 
-                  # bootstrapping does not work with estimators designed for factors
                   data = data.fit2,
-                  # data that is completely numeric
                   bootstrap = nBoots,
                   se = "boot",
+                  missing = "listwise",
                   auto.fix.first = FALSE,
-                  std.lv = TRUE 
+                  std.lv = TRUE,
+                  fixed.x =TRUE
   )
@@
 
-###  f) Latent factor model with missing data imputation
+###  g) Latent factor model with robust estimator and fiml
 
@@
     po02.model,  
-    estimator = "ML", # 
-    data = data.fit2, # data that is completely numeric
+    estimator = "MLR",  
+    data = data.fit2,
     auto.fix.first = FALSE,
@@
   sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem04.pdf"))
+
+```
+
+###  h) Latent factor model with bootstrapping and fiml
+
+```{r}
+
+if (fitmodels) {
+  fit.po05 <- sem(
+    po02.model,  
+    estimator = "ML",  
+    data = data.fit2, 
+    bootstrap = nBoots,
+    se = "boot",
+    auto.fix.first = FALSE,
+    std.lv = TRUE,
+    missing = "fiml",
+    fixed.x = FALSE 
+  )
+  models[['fit.po05']] <- fit.po05
+}
+
+lavaan::summary(fit.po05,standardized = TRUE)
+
+```
+
+```{r}
+fit.po05.est  <-standardizedSolution(fit.po05)
+fit.po05.est %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+fit.po05.est %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
+  # Just the loadings
+  filter(op != '~1')
+
+```
+
+```{r}
+lavaanPlot(
+  model = fit.po05, 
+   labels = labelsforplotting,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem05.pdf"))
 
@@
 
-```{r}
+```{r , fig.height=8}
 
 data.forest  <- fit.po03.best %>% 
-  mutate(estmet = "Bootstrapped"
+  mutate(
+    estmet = "Bootstrapped",
+    missing = "Listwise"
          ) %>%  
@@
   bind_rows(
-    fit.po02.est %>% mutate(estmet = "Robust") %>% 
+    fit.po02.est %>% mutate(
+      estmet = "Robust",
+      missing = "Listwise"
+      ) %>% 
       rename(est=est.std) %>% 
       filter(op == "~"| op == ":="),
-    fit.po04.est %>% mutate(estmet = "Imputed") %>% 
+    fit.po04.est %>% mutate(
+      estmet = "Robust",
+      missing = "FIML"
+      ) %>% 
+      rename(est=est.std) %>% 
+      filter(op == "~"| op == ":="),
+    fit.po05.est %>% mutate(
+      estmet = "Boostrapped",
+      missing = "FIML"
+      ) %>% 
       rename(est=est.std) %>% 
       filter(op == "~"| op == ":=")
+
   ) %>% 
@@
   rename(est=est.std) %>% 
-  mutate(estmet = "Bootstrapped") %>%  
+  mutate(
+    estmet = "Bootstrapped",
+    missing = "Listwise"
+    ) %>%  
   filter(op == "~"| op == ":=") %>% 
   bind_rows(
-    fit.po02ss.est %>% mutate(estmet = "Robust") %>% 
+    fit.po02ss.est %>% mutate(
+      estmet = "Robust",
+      missing = "Listwise"
+      ) %>% 
       rename(est=est.std) %>% 
       filter(op == "~"| op == ":=") ,
-    fit.po04ss.est %>% mutate(estmet = "Imputed") %>% 
+    fit.po04ss.est %>% mutate(
+      estmet = "Robust",
+      missing = "FIML"
+      ) %>% 
+      rename(est=est.std) %>% 
+      filter(op == "~"| op == ":="),
+    fit.po05.est %>% mutate(
+      estmet = "Boostrapped",
+      missing = "FIML"
+      ) %>% 
       rename(est=est.std) %>% 
@@
     estmet = as.factor(estmet),
-    hollow_group=factor(if_else(ci.lower<=0 & ci.upper>=0, NA_character_, as.character(estmet)),
-                             levels=levels(estmet))
-  ) %>% 
-  mutate(
-    mtype = str_c(estmet,"-",linetype)
+    mtype = as.factor(str_c(estmet,"-",linetype, "-",missing)),
+    color = as.factor(str_c(estmet," and ",missing)),
+    hollow_group=factor(if_else(ci.lower<=0 & ci.upper>=0, NA_character_, as.character(color)),
+                             levels=levels(color))
   )
 
-pos <- position_nudge(y=as.numeric(as.factor(data.forest$mtype))/7 -0.5)
+pos <- position_nudge(y=as.numeric(as.factor(data.forest$mtype))/10 -0.5)
 
 pf <- data.forest %>% 
-  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=estmet,linetype = linetype))+
+  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=color,linetype = linetype))+
   geom_vline(xintercept=0,color="grey") +
@@
        filename = here("figures/forest01.png"),
-       height = 140,
+       height = 250,
        width = 180,
@@
        filename = here("figures/forest01.pdf"),
-       height = 140,
+       height = 250,
        width = 180,
@@
 ```{r}
-
+if (fitmodels) {
   gfi<- round(
@@
       list(
-      SS=fit.po02ss,
-      SSB=fit.po03ss,
-      SSI = fit.po04ss,
-      FA=fit.po02,
-      FAB=fit.po03, 
-      FAI=fit.po04
+        SSRL=fit.po02ss,
+        SSBL=fit.po03ss,
+        SSRF = fit.po04ss,
+        SSBL = fit.po05ss,
+        LFRL=fit.po02,
+        LFBL=fit.po03, 
+        LFRF=fit.po04,
+        LFBF=fit.po05
       ),
@@
     ,3)
-
+  models[['gfi']] <- gfi
+} else {
+  gfi <- models$gfi
+}
 gfi

### 2025-04-28 preregistration.Rmd
dates

@@
 
-A set of blood donors will be randomly selected from all blood donors in FRCBS register who have donated at least once during the 730 preceding the day the questionnaire e-mail will be sent i.e.
-5.5.2025 .
+A set of blood donors will be randomly selected from all blood donors in FRCBS register who have donated at least once during the 730 preceding the day the questionnaire e-mail will be sent in May 2025 .
 
@@
 
-The online questionnaire will be open for 2 weeks i.e. from 5.5.2025 to 18.5.2025.
+The online questionnaire will be open for 2 weeks.
 

### 2025-04-28 preregistration_code.Rmd
ss as in pub

@@
 # sometimes I get strange errors about C stack which can be solved with this
-#data <- as.data.frame(data)
+data <- as.data.frame(data)
 #fitmodels <- TRUE # TRUE for fitting models
@@
   mutate(
-    ss_psycown = scale(as.numeric(po01) + as.numeric(po02) + as.numeric(po03)),
-    ss_selfid = scale(as.numeric(si01) + as.numeric(si02) + as.numeric(si03)),
-    ss_intention = scale(as.numeric(it01) + as.numeric(it02) + as.numeric(it03))
+    ss_psycown = 
+      across(starts_with("po"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) %>% 
+      scale() %>% as.numeric(),
+    ss_selfid = 
+      across(starts_with("si"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) %>% 
+      scale() %>% as.numeric(),
+    ss_intention = 
+      across(starts_with("it"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) %>% 
+      scale() %>% as.numeric()
     )

### 2025-04-29 preregistration.Rmd
Preregistration
PsychArchives request

@@
 In this study we try to repeat results from Edwards et al., 2023 in Finland by collecting similar questionnaire data from donors of Finnish Red Cross Blood Service (FRCBS) and modeling the questionnaire data with a structural equation model.
-We include pre-registered R analysis code for direct repetition of Edwards et al., 2023 in Appendix 1.
+We include pre-registered R analysis code for direct repetition of Edwards et al., 2023 in Appendix 1 (created in https://github.com/FRCBS/bd_psychological_ownership_prereg.git ).
 
@@
 
-\includepdf[pages={-}]{preregistration_code.pdf}
+See Appendix 1 for preregistered analysis code in R with simulated data.
 

### 2025-05-21 codebook_data.Rmd
data

--- 2025-04-22 codebook_data_simulated.Rmd
@@
 ---
-title: "Codebook created on 2025-04-22 at 2025-04-22 13:34:49"
+title: "Codebook created on 2025-05-21 at 2025-05-21 15:18:06"
 output: github_document
@@
 ## Dataset description
-The data contains 1000 cases and 28 variables.
+The data contains 634 cases and 28 variables.
 
@@
 ```{r}
-summaries <- read.csv("codebook_data_simulated.csv", stringsAsFactors = FALSE)
+summaries <- read.csv("codebook_data.csv", stringsAsFactors = FALSE)
 summaries <- summaries[, !colSums(is.na(summaries)) == nrow(summaries)]

### 2025-05-21 prepare_data.R
closed data added

@@ -12,9 +12,9 @@ data_simulated <- read.csv(here("data/simulated_data.csv"))[,-1]
 
 #####################################################################################################################################
 # Read real data in 
-# file <- here("data/Tulokset - xxxxx.xlsx")
-# data <- read_xlsx(file,n_max = XX)
-# summary(data)
+file <- here("data/bdpsyone.csv")
+data <- read.csv(file)
+ summary(data)
 #####################################################################################################################################
 
 f1 <- function (x){
@@ -60,9 +60,30 @@ data_simulated <- data_simulated %>%
       bd04 = as.numeric(as.Date(dttma) - bd04)
       
   )  
+summary(data_simulated)
 
+data <- data %>% 
+  mutate_at(
+    vars(
+      qabbr[1:21]),
+    ~ f1(.) 
+  ) %>% 
+  mutate(
+    dttma = as.POSIXct(dttma),
+    bd01 = factor(bd01, levels = c("no","unsure","yes"),ordered = TRUE),
+    bd02 = as.numeric(bd02),
+    bd03 = as.numeric(bd03),
+    #bd04 = as.Date(bd04),
+    de01 = factor(de01),
+    de02 = as.numeric(de02)#,
+    #Days between answering and last donation, done already
+    #bd04 = as.numeric(as.Date(dttma) - bd04)
+    
+  )  %>% select(-X,-consent,-open)
+
+summary(data)
 #####################################################################################################################################
-original_name <- here(paste0("data/",deparse(substitute(data_simulated)), ".csv")) # when real data collected this will be changed to it
+original_name <- here(paste0("data/",deparse(substitute(data)), ".csv")) # when real data collected this will be changed to it
 #####################################################################################################################################
 
 # Above is your data cleaning script
@@ -70,7 +91,7 @@ original_name <- here(paste0("data/",deparse(substitute(data_simulated)), ".csv"
 # When you have real data, it will save the real data locally.
 closed_data(
   #####################################################################################################################################
-  data = data_simulated, # when real data collected this will be changed to it
+  data = data, # when real data collected this will be changed to it
   #####################################################################################################################################
   filename = original_name, 
   synthetic = FALSE)
@@ -85,3 +106,5 @@ add_synthetic(
   synthetic_name = synthetic_name,
   original_name = original_name,
 )
+
+

### 2025-05-21 preregistration_code.Rmd
closed data

@@
 load_data()
-if (exists("data_simulated")) {
-  data <- data_simulated
-  rm(data_simulated)
-}
+# if (exists("data_simulated")) {
+#   data <- data_simulated
+#   rm(data_simulated)
+# }
+```
+
+```{r}
 # sometimes I get strange errors about C stack which can be solved with this

### 2025-05-26 preregistration_code.Rmd
stop before modelling
abby test
#after data collection changes

--- 2025-04-28 preregistration_code.Rmd
@@
 load_data()
-if (exists("data_simulated")) {
-  data <- data_simulated
-  rm(data_simulated)
-}
+
+# if (exists("data_simulated")) {
+#   data <- data_simulated
+#   rm(data_simulated)
+# }
+```
+
+```{r}
 # sometimes I get strange errors about C stack which can be solved with this
 data <- as.data.frame(data)
-#fitmodels <- TRUE # TRUE for fitting models
-fitmodels <- FALSE # TRUE for fitting models
+fitmodels <- TRUE # TRUE for fitting models
+#fitmodels <- FALSE # TRUE for fitting models
 if (!fitmodels) {
@@
 # Plot
+
+##data cleaning
+
+```{r}
+
+#afterdata changing negative blood donation history to missing data for bd04 from 3 participants
+data %>% filter(bd04 < 0)
+data <- data %>% 
+  mutate(bd04 = ifelse(bd04 < 0, NA, bd04))
+
+```
 
@@
 
-p <- ggplot(data %>% filter(de01 == 'man' | de01 == "women"))
+p <- ggplot(data %>% filter(de01 == 'man' | de01 == "woman"))
 p <- p +  geom_histogram(aes(x=de02),binwidth = 5)
@@
 ```{r}
+
+#how would we seperate out ie01; as it shouldn't sit with it01
+
 gglikert(
@@
 
+#removing non-eligible participants (must be: over 18, have donated blood in past 2 years, believe currently eligible to donate blood)
+
 data %>%
-  filter(!(de01 == 'man' | de01 == "women"))  %>% nrow()
+  filter(!(de01 == 'man' | de01 == "woman"))  %>% nrow()
 cat("are removed because of sex\n")
@@
 
+#afterdata adding additional exclusion criteria in line with preregistered exclusion for bd02 & bd03
+
+data %>%
+  filter(bd02 < 1) %>% nrow()
+cat("are removed because of non-donor\n")
+
+data %>%
+  filter(bd03 < 1) %>% nrow()
+cat("are removed because of 2 year donation\n")
+
 data %>%
   #lavaan cannot deal with unordered factor with more then 2 levels
-  filter(de01 == 'man' | de01 == "women") %>% 
-  filter(bd01 != 'no') %>% nrow()
+  filter(de01 == 'man' | de01 == "woman") %>% 
+  filter(bd01 != 'no') %>% 
+  filter (bd02 >= 1) %>%
+  filter (bd03 >= 1)%>% nrow()
 cat("in total after exclusions.\n")
@@
 isfactor <- data %>%  summarise(across(everything(), ~ is.factor(.x))) %>%  as.logical() %>% which()
+
+#afterdata adding in additional exclusion criteria for bd02 & bd03
+data <- data %>% 
+  filter(bd02 >= 1)
+
+data <- data %>% 
+  filter(bd03 >= 1)
 
@@
   #lavaan cannot deal with unordered factor with more then 2 levels
-  filter(de01 == 'man' | de01 == "women") %>% # EXCLUSION
+  filter(de01 == 'man' | de01 == "woman") %>% # EXCLUSION
   filter(bd01 != 'no') %>% # EXCLUSION
@@
 ```{r}
-mcar_test(data.fit2 %>% select(-dttma))
+
+#afterdata mcar not running, checking if items are too highly correlated
+    #taking out sum scores as these were highly correlated with items that made them (i.e., ss_psyown highly correlated with po01, po02, po03)
+
+data.fit3 <- data.fit2 %>%
+  select(-ss_psycown,
+         -ss_selfid, 
+         -ss_intention)
+
+corrplot(cor(data.fit3,use="pairwise.complete.obs"))
+mcar_test(data.fit3 %>% select(-dttma))
+
+#afterdata curious what mcar looks like without eO1
+#data.fit4 <- data.fit2 %>%
+ # select(-ie01)
+#mcar_test(data.fit4 %>% select(-dttma))
+
+#mcar_test(data.fit2 %>% select(-dttma)) #doesn't run, likely due to high correlations
 # p < 0.05 -> Reject the null hypothesis-> Data is MAR (Missing at random) 
@@
 ## Confirmatory factor analysis
+
+```{r}
+#afterdata model comparison won't run because no participants responded with "1" for ai01 ("I am intimately familiar with what is going on with regard to donating blood")
+
+data.fit2 <- data.fit2 %>%
+  mutate(
+    ai01 = as.numeric(as.character(ai01)),  # convert to numeric if factor
+    ai01 = ai01 - 1,                         # subtract 1 so lowest category is now 1
+    #ai01 = factor(ai01, ordered = TRUE)      # convert back to ordered factor
+  )
+table(data.fit$ai01)
+
+```
 
@@
 #Make a model with out prespecified correlation structure
-fit.po01.cfa.o <- cfa(po01.model,data.fit,orthogonal=TRUE)
+fit.po01.cfa.o <- cfa(po01.model,data.fit2,orthogonal=TRUE)
 round(
@@
 #cfi and chisq/df should be bigger, rmsea and srmr smaller in model cfa
+```
+
+```{r}
+knitr::knit_exit()
+#Stop here with real data and review results before modelling.
 ```

### 2025-05-30 preregistration_code.Rmd
Applying additional cleaning

@@
 ```{r}
+data %>% filter(bd04 < 0)
+data %>% filter(bd03 > 12)
+
+data <- data %>% 
+  mutate(
+    bd04 = ifelse(bd04 < 0, NA, bd04)/7,
+    bd03 = ifelse(bd03 > 12, NA, bd03)
+)
 
 #afterdata changing negative blood donation history to missing data for bd04 from 3 participants
-data %>% filter(bd04 < 0)
-data <- data %>% 
-  mutate(bd04 = ifelse(bd04 < 0, NA, bd04))
-
+#afterdata changing values over 12 (maximum number of whole blood donations for men within 2 years) for bd03 to missing.
+#afterdata rescaling bd04 to number of weeks since last donation
 ```
@@
 
-#afterdata adding additional exclusion criteria in line with preregistered exclusion for bd02 & bd03
-
 data %>%
@@
 
+#afterdata adding additional exclusion criteria in line with preregistered exclusion for bd02 & bd03  
 ```
@@
 
-#afterdata adding in additional exclusion criteria for bd02 & bd03
 data <- data %>% 
@@
 
+#afterdata adding in additional exclusion criteria for bd02 & bd03
 ```
@@
 
-#afterdata mcar not running, checking if items are too highly correlated
-    #taking out sum scores as these were highly correlated with items that made them (i.e., ss_psyown highly correlated with po01, po02, po03)
-
 data.fit3 <- data.fit2 %>%
@@
 mcar_test(data.fit3 %>% select(-dttma))
+
+#afterdata mcar not running, checking if items are too highly correlated
+    #taking out sum scores as these were highly correlated with items that made them (i.e., ss_psyown highly correlated with po01, po02, po03)
 
@@
 ```{r}
+
+data.fit2 <- data.fit2 %>%
+  mutate(
+    ai01 = as.numeric(as.character(ai01)),  # convert to numeric if factor
+    #ai01 = ai01 - 1,                         # subtract 1 so lowest category is now 1
+    #ai01 = factor(ai01, ordered = TRUE)      # convert back to ordered factor
+  )
+table(data.fit$ai01)
+
 #afterdata model comparison won't run because no participants responded with "1" for ai01 ("I am intimately familiar with what is going on with regard to donating blood")
-
-data.fit2 <- data.fit2 %>%
-  mutate(
-    ai01 = as.numeric(as.character(ai01)),  # convert to numeric if factor
-    ai01 = ai01 - 1,                         # subtract 1 so lowest category is now 1
-    #ai01 = factor(ai01, ordered = TRUE)      # convert back to ordered factor
-  )
-table(data.fit$ai01)
-
+  #I'm not sure why but if you just specify that ai01 is numeric it runs without needed to merge level 1 and 2. 
 ```
@@
 ```{r}
-knitr::knit_exit()
+#knitr::knit_exit()
 #Stop here with real data and review results before modelling.

### 2025-06-06 preregistration_code.Rmd
standardising behaviour measures and creating combined donation behaviour measure for sum scores

@@
 
+#standarsing behaviour variables
+data <- data %>%
+mutate(across(c(bd02, bd03, bd04), ~scale(.) %>% as.numeric()))
+
 #lavaan does not understand empty factor levels
@@
 #afterdata adding in additional exclusion criteria for bd02 & bd03
+#afterdata adding in standardisation of bd02, bdo3 and bd04 prior to scale creation
+#afterdata creating sum score variable of donations (bd02 + bd03 + bd04) to be used in analysis as per preregistration
 ```
@@
                           -ss_selfid,
+                          -ss_donation,
                           ), 
@@
          -ss_selfid, 
-         -ss_intention)
+         -ss_intention,
+         -ss_donation)
 
@@
 # does not prove that data is MCAR (Missing Completely at Random)
+
+#afterdata removing sum score donation variable (ss_donation)
 ```
@@
 #mediation
-ss_psycown ~ a1 *  bd02 
-ss_selfid ~ a2 * bd02 + d21 * ss_psycown
+ss_psycown ~ a1 *  ss_donation
+ss_selfid ~ a2 * ss_donation + d21 * ss_psycown
 
 #regression
-ss_intention ~  de01 + de02 + ie01 +  cp * bd02 + b1 * ss_psycown + b2 * ss_selfid
+ss_intention ~  de01 + de02 + ie01 +  cp * ss_donation + b1 * ss_psycown + b2 * ss_selfid
 
@@
 cat(po02ss.model)
+
+#afterdata using sum score donation behaviour variable as per preregistered hypothesis
 
@@
   "ss_intention" = "Intention",
-  "bd02" = "Blood donation history"
+  "ss_donation" = "Blood donation history"
 )
@@
     term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
-    term =str_replace_all(term,"bd02","blooddonor:"),
+    term =str_replace_all(term,"ss_donation","blooddonor:"),
     term =str_replace_all(term,"blooddonor:","blooddonor"),

### 2025-06-06 preregistration_code.Rmd
standardising behaviour measures and creating combined donation behaviour measure for sum scores

@@
       rowMeans(na.rm = TRUE) %>% 
-      scale() %>% as.numeric()
+      scale() %>% as.numeric(),
+    ss_donation = 
+      across(all_of(c("bd02", "bd03", "bd04")), ~as.numeric(.)) %>%
+      rowSums(na.rm = TRUE)
     )

### 2025-06-09 manuscript_code.Rmd
Manuscript & exploratory code

--- 2025-06-06 preregistration_code.Rmd
@@
 ---
-title: "Appendix 1: Preregistered analysis with simulated data"
+title: "Appendix 2: Manuscript and Exploratory Code"
 author: "Abigail Edwards, Lauri Hietajärvi and Mikko Arvas"
@@
 
+#Exploratory Analysis (Latent factor model with robust estimator and fiml)
+#Exploring a1*b1 comparison
+
+```{r}
+
+dim(factor_scores)
+dim(data.fit2)
+
+factor_scores <- lavPredict(fit.po04, type = "lv")
+head(factor_scores)
+summary(factor_scores)
+
+df_scores <- cbind(
+  as.data.frame(factor_scores),
+  de02 = data.fit2$de02,
+  de01 = data.fit2$de01
+)
+
+ggplot(df_scores, aes(x = psycown, y = intention,
+                      color = as.factor(de01))) +
+  geom_point(alpha = 0.7, size = 2.5) +
+  labs(title = "Latent Factor Scores by de01 and de02",
+       x = "Psychological Ownership",
+       y = "Donation Intention",
+       color = "gender") +
+  theme_minimal()
+
+ggplot(df_scores, aes(x = p, y = intention,
+                      color = as.factor(de01))) +
+  geom_point(alpha = 0.7, size = 2.5) +
+  labs(title = "Latent Factor Scores by de01 and de02",
+       x = "Psychological Ownership",
+       y = "Donation Intention",
+       color = "gender") +
+  theme_minimal()
+
+ggplot(df_scores, aes(x = psycown, y = intention,
+                      color = as.factor(de02),
+                      shape = as.factor(de01))) +
+  geom_point(alpha = 0.7, size = 2.5) +
+  labs(title = "Latent Factor Scores by de01 and de02",
+       x = "Psychological Ownership",
+       y = "Donation Intention",
+       color = "age",
+       shape = "gender") +
+  theme_minimal()
+
+```
+
 ###  h) Latent factor model with bootstrapping and fiml
@@
   save(models,file=here("prereg_models.rdata"))
+
 }

### 2025-06-10 manuscript_code.Rmd
changes to factor score graphs

--- 2025-06-06 preregistration_code.Rmd
@@
 library(naniar)
+library(gridExtra)
 #library(semPlot) # update plots to semPlot later
@@
 #library(tidySEM) # https://github.com/yrosseel/lavaan/issues/359 lavaan does not work if tidySEM is loaded!
+#library(griExtra) #used for combining multiple plots
 
@@
       rowMeans(na.rm = TRUE) %>% 
-      scale() %>% as.numeric()
+      scale() %>% as.numeric(),
+    #ss_donation as only measuring bd02
+    ss_donation = 
+      across(all_of(c("bd02")), ~as.numeric(.)) %>%
+      rowSums(na.rm = TRUE)
     )
@@
 ```{r}
+eval=FALSE
 gg_miss_var(
@@
 ```{r}
+eval=FALSE
 gg_miss_upset(data.fit2,nsets = 10,nintersects=40)
@@
 ```{r}
-
+eval=FALSE
 gg_miss_case_cumsum(
@@
 ```{r}
-
+eval=FALSE
 gg_miss_fct(
@@
 ```{r}
-
+eval=FALSE
 data.fit3 <- data.fit2 %>%
@@
 ```{r}
+#making bd02 the only variable in donation history
 
@@
                 iv =~ iv01 + iv02 + iv03
-                bd =~ bd02 + bd03 + bd04 #bd02 is sure to correlate positively
+                bd =~ bd02
 
@@
 ```{r}
+eval=FALSE
 #https://stats.stackexchange.com/questions/340857/serial-mediation-in-r-how-to-setup-the-model
@@
 ```
+```{r}
+#Bivariate correlation table for paper
+
+```
 
@@
 ```{r}
+eval=FALSE
 if (fitmodels) {
@@
 ```{r}
+eval=FALSE
 lavaan::inspect(fit.po02ss,"r2")
@@
 ```{r}
+eval=FALSE
 fit.po02ss.est  <-standardizedSolution(fit.po02ss)
@@
 ```{r}
-
+eval=FALSE
 fit.po02ss.est  <-standardizedSolution(fit.po02ss)
@@
 ```{r}
-
+eval=FALSE
 labelsforplottingnull <- c(
@@
 ```{r}
-nBoots <- 1000
-
-if (fitmodels) {
-  fit.po03ss <- sem(po02ss.model,  
-                    estimator = "ML",  
-                    missing = "listwise",
-                    data = data.fit2, 
-                    bootstrap = nBoots,
-                    se = "boot",
-                    auto.fix.first = FALSE,
-                    std.lv = TRUE,
-                    fixed.x = TRUE
-  )
-  models[['fit.po03ss']] <- fit.po03ss
-}
-
-lavaan::summary(fit.po03ss,standardized = TRUE)
-
-```
-
-```{r}
-fit.po03ss.best  <-standardizedSolution(fit.po03ss)
-fit.po03ss.best %>% select(-label,se,z) %>%  filter(op == '~')
-```
-
-```{r}
-fit.po03ss.best %>% 
-  filter(abs(est.std) < 0.3) %>% 
-  select(-z,-se) %>% 
-  arrange(abs(est.std)) %>% 
+eval=FALSE
+#nBoots <- 1000
+
+#if (fitmodels) {
+  #fit.po03ss <- sem(po02ss.model,  
+                    #estimator = "ML",  
+                    #missing = "listwise",
+                    #data = data.fit2, 
+                    #bootstrap = nBoots,
+                    #se = "boot",
+                    #auto.fix.first = FALSE,
+                    #std.lv = TRUE,
+                    #fixed.x = TRUE
+  #)
+  #models[['fit.po03ss']] <- fit.po03ss
+#}
+
+#lavaan::summary(fit.po03ss,standardized = TRUE)
+
+```
+
+```{r}
+#eval=FALSE
+#fit.po03ss.best  <-standardizedSolution(fit.po03ss)
+#fit.po03ss.best %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+eval=FALSE
+#fit.po03ss.best %>% 
+  #filter(abs(est.std) < 0.3) %>% 
+  #select(-z,-se) %>% 
+  #arrange(abs(est.std)) %>% 
   # Just the loadings
-  filter(op != '~1')
-
-```
-
-```{r}
-lavaanPlot(
-  model = fit.po03ss, 
-   labels = labelsforplottingnull,
-  node_options = list(shape = "box", fontname = "Helvetica"), 
-  edge_options = list(color = "grey"), 
-  coefs = TRUE, 
-  stand = TRUE,
-  sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))
+  #filter(op != '~1')
+
+```
+
+```{r}
+eval=FALSE
+#lavaanPlot(
+  #model = fit.po03ss, 
+   #labels = labelsforplottingnull,
+  #node_options = list(shape = "box", fontname = "Helvetica"), 
+  #edge_options = list(color = "grey"), 
+  #coefs = TRUE, 
+  #stand = TRUE,
+  #sig=0.001
+  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))
 
@@
 ```{r}
-
+eval=FALSE
 if (fitmodels) {
@@
 ```{r}
+eval=FALSE
 fit.po04ss.est  <-standardizedSolution(fit.po04ss)
@@
 ```{r}
+eval=FALSE
 fit.po04ss.est %>% 
@@
 ```{r}
+eval=FALSE
 lavaanPlot(
@@
 ```{r}
-
-if (fitmodels) {
-  fit.po05ss <- sem(
-    po02ss.model,  
-    estimator = "ML", 
-    bootstrap = nBoots,
-    se = "boot",
-    data = data.fit2, 
-    auto.fix.first = FALSE,
-    std.lv = TRUE,
-    missing = "fiml",
-    fixed.x = FALSE 
-  )
-  models[['fit.po05ss']] <- fit.po05ss
-}
-
-lavaan::summary(fit.po05ss,standardized = TRUE)
-
-```
-
-```{r}
-fit.po05ss.est  <-standardizedSolution(fit.po05ss)
-fit.po05ss.est %>% select(-label,se,z) %>%  filter(op == '~')
-```
-
-```{r}
-fit.po05ss.est %>% 
-  filter(abs(est.std) < 0.3) %>% 
-  select(-z,-se) %>% 
-  arrange(abs(est.std)) %>% 
+eval=FALSE
+#if (fitmodels) {
+  #fit.po05ss <- sem(
+    #po02ss.model,  
+    #estimator = "ML", 
+    #bootstrap = nBoots,
+    #se = "boot",
+    #data = data.fit2, 
+    #auto.fix.first = FALSE,
+    #std.lv = TRUE,
+    #missing = "fiml",
+    #fixed.x = FALSE 
+  #)
+  #models[['fit.po05ss']] <- fit.po05ss
+#}
+
+#lavaan::summary(fit.po05ss,standardized = TRUE)
+
+```
+
+```{r}
+eval=FALSE
+#fit.po05ss.est  <-standardizedSolution(fit.po05ss)
+#fit.po05ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+eval=FALSE
+#fit.po05ss.est %>% 
+  #filter(abs(est.std) < 0.3) %>% 
+  #select(-z,-se) %>% 
+  #arrange(abs(est.std)) %>% 
   # Just the loadings
-  filter(op != '~1')
-
-```
-
-```{r}
-lavaanPlot(
-  model = fit.po05ss, 
-   labels = labelsforplottingnull,
-  node_options = list(shape = "box", fontname = "Helvetica"), 
-  edge_options = list(color = "grey"), 
-  coefs = TRUE, 
-  stand = TRUE,
-  sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem05.pdf"))
+  #filter(op != '~1')
+
+```
+
+```{r}
+eval=FALSE
+#lavaanPlot(
+  #model = fit.po05ss, 
+   #labels = labelsforplottingnull,
+  #node_options = list(shape = "box", fontname = "Helvetica"), 
+  #edge_options = list(color = "grey"), 
+  #coefs = TRUE, 
+  #stand = TRUE,
+  #sig=0.001
+  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem05.pdf"))
 
@@
 ```{r}
+
+#making bd02 the only variable in donation history
 
@@
 intention =~ it01 + it02 + it03
-blooddonor =~ bd02 + bd03 +bd04 
+blooddonor =~ bd02 
 #bd01 is just used for exclusion of "no"
@@
 
-if (fitmodels) {
-  fit.po03 <- sem(po02.model,  
-                  estimator = "ML", 
-                  data = data.fit2,
-                  bootstrap = nBoots,
-                  se = "boot",
-                  missing = "listwise",
-                  auto.fix.first = FALSE,
-                  std.lv = TRUE,
-                  fixed.x =TRUE
-  )
-  models[['fit.po03']] <- fit.po03
-}
-
-lavaan::summary(fit.po03,standardized = TRUE)
-
-```
-
-```{r}
-fit.po03.best <- parameterEstimates(fit.po03)
-fit.po03.best %>% select(-se,-z) %>% filter(op == "~")
-```
-
-```{r}
-lavaanPlot(
-  model = fit.po03, 
-   labels = labelsforplotting,
-  node_options = list(shape = "box", fontname = "Helvetica"), 
-  edge_options = list(color = "grey"), 
-  coefs = TRUE, 
-  stand = TRUE,
-  sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem03.pdf"))
+#if (fitmodels) {
+  #fit.po03 <- sem(po02.model,  
+                  #estimator = "ML", 
+                  #data = data.fit2,
+                  #bootstrap = nBoots,
+                  #se = "boot",
+                  #missing = "listwise",
+                  #auto.fix.first = FALSE,
+                  #std.lv = TRUE,
+                  #fixed.x =TRUE
+  #)
+ # models[['fit.po03']] <- fit.po03
+#}
+
+#lavaan::summary(fit.po03,standardized = TRUE)
+
+```
+
+```{r}
+#fit.po03.best <- parameterEstimates(fit.po03)
+#fit.po03.best %>% select(-se,-z) %>% filter(op == "~")
+```
+
+```{r}
+#lavaanPlot(
+  #model = fit.po03, 
+   #labels = labelsforplotting,
+  #node_options = list(shape = "box", fontname = "Helvetica"), 
+  #edge_options = list(color = "grey"), 
+  #coefs = TRUE, 
+  #stand = TRUE,
+  #sig=0.001
+  #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem03.pdf"))
 
@@
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem04.pdf"))
+
+```
+
+#Exploratory Analysis (a1*b1 comparison)
+
+```{r}
+
+factor_scores <- lavPredict(fit.po04, type = "lv")
+head(factor_scores)
+summary(factor_scores)
+
+df_scores <- cbind(
+  as.data.frame(factor_scores),
+  de02 = data.fit2$de02, #de02 = age
+  de01 = data.fit2$de01 #de01 = gender
+)
+
+df_scores$de01 <- factor(df_scores$de01, levels = c(1, 2), labels = c("man", "woman"), ordered = TRUE)
+
+#psycown ~ blooddonor
+factor1 <- ggplot(df_scores, aes(x = psycown, y = blooddonor)) +
+  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
+  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
+  labs(title = "Factor scores: ownership x identity",
+       x = "Identity",
+       y = "Donation history",
+       color = "Age",
+       linetype = "Gender") +
+  theme_minimal()
+
+#selfid ~ blooddonor
+factor2 <- ggplot(df_scores, aes(x = selfid, y = blooddonor)) +
+  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
+  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
+  labs(title = "Factor scores: ownership x identity",
+       x = "Identity",
+       y = "Donation history",
+       color = "Age",
+       linetype = "Gender") +
+  theme_minimal()
+
+#selfid ~ psycown
+factor3 <- ggplot(df_scores, aes(x = psycown, y = selfid)) +
+  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
+  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
+  labs(title = "Factor scores: ownership x identity",
+       x = "Psychological Ownership",
+       y = "Identity",
+       color = "Age",
+       linetype = "Gender") +
+  theme_minimal()
+
+#intention ~ blooddonor
+factor4 <- ggplot(df_scores, aes(x = intention, y = blooddonor,
+                      color = as.factor(de01))) +
+  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
+  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
+  labs(title = "Factor scores: ownership x identity",
+       x = "Intention",
+       y = "Donation History",
+       color = "Age",
+       linetype = "Gender") +
+  theme_minimal()
+
+#intention ~ psycown
+factor5 <- ggplot(df_scores, aes(x = psycown, y = intention,
+                      color = as.factor(de01))) +
+  geom_point(alpha = 0.7, size = 2.5) +
+  geom_smooth() +
+  labs(title = "Factor Scores: ownership x intention",
+       x = "Psychological Ownership",
+       y = "Intention",
+       color = "Age",
+       linetype = "Gender") +
+  theme_minimal()
+
+#intention ~ selfid
+factor6 <- ggplot(df_scores, aes(x = psycown, y = intention,
+                      color = as.factor(de01))) +
+  geom_point(alpha = 0.7, size = 2.5) +
+  geom_smooth() +
+  labs(title = "Factor Scores: ownership x intention",
+       x = "Psychological Ownership",
+       y = "Intention",
+       color = "Age",
+       linetype = "Gender") +
+  theme_minimal()
+
+combined <- grid.arrange(factor1, factor2, factor3, factor4, factor5, factor6, ncol = 2)
+
+ggsave(
+  "Ownership and Intention Factor Loadings",
+  combined, width = 20, 
+  height = 10, 
+  device = "pdf")
+
+```
+
+#exploratory analysis (multigroup analysis to explore impact of gender)
+
+```{r}
+
+#making gender a factor and ordering
+
+data.fit2$de01 <- factor(data.fit2$de01, levels = c(1, 2), labels = c("man", "woman"), ordered = TRUE)
+
+data.fit2 %>%
+  group_by(de01) %>%
+  summarise(across(everything(), ~ var(.x, na.rm = TRUE)))
+
+#defining new model that doesn't contained de01
+
+ea2.model <- ' 
+#measurement model
+psycown =~ po01 + po02 + po03
+selfid =~ si01 + si02 + si03
+intention =~ it01 + it02 + it03
+blooddonor =~ bd02
+
+# regressions (separate labels per group)
+psycown ~ c(a1g1, a1g2) * blooddonor
+selfid ~ c(a2g1, a2g2) * blooddonor + c(d21g1, d21g2) * psycown
+intention ~ c(cpg1, cpg2) * blooddonor + c(b1g1, b2g2) * psycown + c(b2g1, b2g2) * selfid
+
+#g1 indirect effects
+ind_eff_g1 := a1g1 * d21g1 * b2g1
+ind_eff1_g1:= a1g1 * b1g1
+ind_eff2_g1:= a2g1 * b2g1
+
+#g2 indirect effects
+ind_eff_g2:= a1g2 * d21g2 * b2g2
+ind_eff1_g2:= a1g2 * b1g2
+ind_eff2_g2:= a2g2 * b2g2
+'
+
+ea2.free <- sem(
+  ea2.model,
+  data = data.fit2,
+  group = "de01",
+  estimator = "MLR",
+  std.lv = TRUE,
+  missing = "fiml",
+  fixed.x = FALSE,
+  auto.fix.first = FALSE,
+  group.equal = c("loadings","intercepts”)
+)
+
+ea2.fixed <- sem(
+ea2.model,
+data = data.fit2,
+group = "de01",
+estimator = "MLR",
+std.lv = TRUE,
+missing = "fiml",
+fixed.x = FALSE,
+auto.fix.first = FALSE,
+group.equal = c("loadings", ”intercepts”, ”regressions”),
+)
+
+# this is for model comparison
+lavTestLRT(ea2.free, ea2.fixed)
+
+#ea2 <- sem(
+  #ea2.model,
+  #data = data.fit2,
+  #group = "de01",
+  #estimator = "MLR",
+  #std.lv = TRUE,
+  #missing = "fiml",
+  #fixed.x = FALSE,
+  #auto.fix.first = FALSE
+#)
+
+#summary(ea2, standardized = TRUE)
+
+#end of exploratory analysis
+knitr::knit_exit()
 

### 2025-06-10 manuscript_code.Rmd
factor score plots

@@
 eval=FALSE
+#if (fitmodels) {
+  #fit.po02ss <- sem(po02ss.model,  
+                    #estimator = "MLR", 
+                    #missing = "listwise",
+                    #data = data.fit2,
+                    #auto.fix.first = FALSE, 
+                    #std.lv = TRUE,
+                    #fixed.x = TRUE
+
+ # )
+
+  #models[["fit.po02ss"]] <- fit.po02ss
+#}
+
+#lavaan::summary(fit.po02ss,standardized = TRUE)
+
+```
+
+```{r}
+eval=FALSE
+#lavaan::inspect(fit.po02ss,"r2")
+```
+
+```{r}
+eval=FALSE
+#fit.po02ss.est  <-standardizedSolution(fit.po02ss)
+#fit.po02ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+eval=FALSE
+#fit.po02ss.est  <-standardizedSolution(fit.po02ss)
+#fit.po02ss.est %>% 
+  #filter(abs(est.std) < 0.3) %>% 
+  #select(-z,-se) %>% 
+  #arrange(abs(est.std)) %>% 
+  # Just the loadings
+  #filter(op != '~1')
+
+```
+
+```{r}
+labelsforplottingnull <- c(
+  "ss_psycown" = "Psychological ownership",
+  "ss_selfid"  = "Self-identity",
+  "ss_intention" = "Intention",
+  "ss_donation" = "Blood donation history"
+)
+
+#lavaanPlot(
+  #model = fit.po02ss, 
+   #labels = labelsforplottingnull,
+  #node_options = list(shape = "box", fontname = "Helvetica"), 
+  #edge_options = list(color = "grey"), 
+  #coefs = TRUE, 
+  #stand = TRUE,
+  #sig=0.001
+  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
+```
+
+###  b) Sum score model with bootstrapping and listwise deletion
+
+```{r}
+
+nBoots <- 1000
+
 if (fitmodels) {
-  fit.po02ss <- sem(po02ss.model,  
-                    estimator = "MLR", 
+  fit.po03ss <- sem(po02ss.model,  
+                    estimator = "ML",  
                     missing = "listwise",
                     data = data.fit2, 
+                    bootstrap = nBoots,
+                    se = "boot",
                     auto.fix.first = FALSE,
@@
                     fixed.x = TRUE
-
-  )
-
-  models[["fit.po02ss"]] <- fit.po02ss
+  )
+  models[['fit.po03ss']] <- fit.po03ss
 }
 
-lavaan::summary(fit.po02ss,standardized = TRUE)
-
-```
-
-```{r}
-eval=FALSE
-lavaan::inspect(fit.po02ss,"r2")
-```
-
-```{r}
-eval=FALSE
-fit.po02ss.est  <-standardizedSolution(fit.po02ss)
-fit.po02ss.est %>% select(-label,se,z) %>%  filter(op == '~')
-```
-
-```{r}
-eval=FALSE
-fit.po02ss.est  <-standardizedSolution(fit.po02ss)
-fit.po02ss.est %>% 
+lavaan::summary(fit.po03ss,standardized = TRUE)
+
+```
+
+```{r}
+fit.po03ss.best  <-standardizedSolution(fit.po03ss)
+fit.po03ss.best %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+fit.po03ss.best %>% 
   filter(abs(est.std) < 0.3) %>% 
@@
 ```{r}
-eval=FALSE
-labelsforplottingnull <- c(
-  "ss_psycown" = "Psychological ownership",
-  "ss_selfid"  = "Self-identity",
-  "ss_intention" = "Intention",
-  "ss_donation" = "Blood donation history"
-)
-
 lavaanPlot(
-  model = fit.po02ss, 
+  model = fit.po03ss, 
    labels = labelsforplottingnull,
@@
   sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
-```
-
-###  b) Sum score model with bootstrapping and listwise deletion
-
-```{r}
-eval=FALSE
-#nBoots <- 1000
-
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))
+
+```
+
+###  c) Sum score model with robust estimator and fiml
+
+```{r}
+eval=FALSE
 #if (fitmodels) {
-  #fit.po03ss <- sem(po02ss.model,  
-                    #estimator = "ML",  
-                    #missing = "listwise",
+  #fit.po04ss <- sem(
+    #po02ss.model,  
+    #estimator = "MLR", 
     #data = data.fit2, 
-                    #bootstrap = nBoots,
-                    #se = "boot",
     #auto.fix.first = FALSE,
     #std.lv = TRUE,
-                    #fixed.x = TRUE
+    #missing = "fiml",
+    #fixed.x = FALSE 
   #)
-  #models[['fit.po03ss']] <- fit.po03ss
+  #models[['fit.po04ss']] <- fit.po04ss
 #}
 
-#lavaan::summary(fit.po03ss,standardized = TRUE)
-
-```
-
-```{r}
-#eval=FALSE
-#fit.po03ss.best  <-standardizedSolution(fit.po03ss)
-#fit.po03ss.best %>% select(-label,se,z) %>%  filter(op == '~')
-```
-
-```{r}
-eval=FALSE
-#fit.po03ss.best %>% 
+#lavaan::summary(fit.po04ss,standardized = TRUE)
+
+```
+
+```{r}
+eval=FALSE
+#fit.po04ss.est  <-standardizedSolution(fit.po04ss)
+#fit.po04ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+eval=FALSE
+#fit.po04ss.est %>% 
   #filter(abs(est.std) < 0.3) %>% 
@@
 #lavaanPlot(
-  #model = fit.po03ss, 
+  #model = fit.po04ss, 
    #labels = labelsforplottingnull,
@@
   #sig=0.001
-  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))
-
-```
-
-###  c) Sum score model with robust estimator and fiml
-
-```{r}
-eval=FALSE
-if (fitmodels) {
-  fit.po04ss <- sem(
-    po02ss.model,  
-    estimator = "MLR", 
-    data = data.fit2, 
-    auto.fix.first = FALSE,
-    std.lv = TRUE,
-    missing = "fiml",
-    fixed.x = FALSE 
-  )
-  models[['fit.po04ss']] <- fit.po04ss
-}
-
-lavaan::summary(fit.po04ss,standardized = TRUE)
-
-```
-
-```{r}
-eval=FALSE
-fit.po04ss.est  <-standardizedSolution(fit.po04ss)
-fit.po04ss.est %>% select(-label,se,z) %>%  filter(op == '~')
-```
-
-```{r}
-eval=FALSE
-fit.po04ss.est %>% 
-  filter(abs(est.std) < 0.3) %>% 
-  select(-z,-se) %>% 
-  arrange(abs(est.std)) %>% 
-  # Just the loadings
-  filter(op != '~1')
-
-```
-
-```{r}
-eval=FALSE
-lavaanPlot(
-  model = fit.po04ss, 
-   labels = labelsforplottingnull,
-  node_options = list(shape = "box", fontname = "Helvetica"), 
-  edge_options = list(color = "grey"), 
-  coefs = TRUE, 
-  stand = TRUE,
-  sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))
+  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))
 
@@
 ```{r}
-
-if (fitmodels) {
-  fit.po02 <- sem(po02.model,  
-                  estimator = "MLR",
-                  missing = "listwise",
-                  data = data.fit2,
-                  auto.fix.first = FALSE,
-                  std.lv = TRUE,
-                  fixed.x = TRUE
-  )
-  models[["fit.po02"]] <- fit.po02
-}
-lavaan::summary(fit.po02,standardized = TRUE)
-
-```
-
-```{r}
-lavaan::inspect(fit.po02,"r2")
-```
-
-```{r}
-fit.po02.est  <-standardizedSolution(fit.po02)
-fit.po02.est %>% select(-label,se,z) %>%  filter(op == '~')
-```
-
-```{r}
-
-fit.po02.est %>% 
-  filter(abs(est.std) < 0.3) %>% 
-  select(-z,-se) %>% 
-  arrange(abs(est.std)) %>% 
+eval=FALSE
+#if (fitmodels) {
+  #fit.po02 <- sem(po02.model,  
+                  #estimator = "MLR",
+                  #missing = "listwise",
+                  #data = data.fit2,
+                  #auto.fix.first = FALSE,
+                  #std.lv = TRUE,
+                  #fixed.x = TRUE
+  #)
+  #models[["fit.po02"]] <- fit.po02
+#}
+#lavaan::summary(fit.po02,standardized = TRUE)
+
+```
+
+```{r}
+eval=FALSE
+#lavaan::inspect(fit.po02,"r2")
+```
+
+```{r}
+eval=FALSE
+#fit.po02.est  <-standardizedSolution(fit.po02)
+#fit.po02.est %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+eval=FALSE
+#fit.po02.est %>% 
+  #filter(abs(est.std) < 0.3) %>% 
+  #select(-z,-se) %>% 
+  #arrange(abs(est.std)) %>% 
   # Just the loadings
-  filter(op != '~1')
-
-```
-
-```{r}
-
+  #filter(op != '~1')
+
+```
+
+```{r}
 labelsforplotting <- c(
@@
 
-lavaanPlot(
-  model = fit.po02, 
-   labels = labelsforplotting,
-  node_options = list(shape = "box", fontname = "Helvetica"), 
-  edge_options = list(color = "grey"), 
-  coefs = TRUE, 
-  stand = TRUE,
-  sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))
+#lavaanPlot(
+  #model = fit.po02, 
+   #labels = labelsforplotting,
+  #node_options = list(shape = "box", fontname = "Helvetica"), 
+  #edge_options = list(color = "grey"), 
+  #coefs = TRUE, 
+  #stand = TRUE,
+  #sig=0.001
+  #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))
 ```
@@
 
-factor_scores <- lavPredict(fit.po04, type = "lv")
+#factor_scores <- lavPredict(fit.po04, type = "lv") #taking factor scores from CFA 
+factor_scores <- lavPredict(fit.po01.cfa, type = "lv")
 head(factor_scores)
@@
 #end of exploratory analysis
-knitr::knit_exit()
+#knitr::knit_exit()
 
@@
 
-if (fitmodels) {
-  fit.po05 <- sem(
-    po02.model,  
-    estimator = "ML",  
-    data = data.fit2, 
-    bootstrap = nBoots,
-    se = "boot",
-    auto.fix.first = FALSE,
-    std.lv = TRUE,
-    missing = "fiml",
-    fixed.x = FALSE 
-  )
-  models[['fit.po05']] <- fit.po05
-}
-
-lavaan::summary(fit.po05,standardized = TRUE)
-
-```
-
-```{r}
-fit.po05.est  <-standardizedSolution(fit.po05)
-fit.po05.est %>% select(-label,se,z) %>%  filter(op == '~')
-```
-
-```{r}
-fit.po05.est %>% 
-  filter(abs(est.std) < 0.3) %>% 
-  select(-z,-se) %>% 
-  arrange(abs(est.std)) %>% 
+#if (fitmodels) {
+  #fit.po05 <- sem(
+    #po02.model,  
+    #estimator = "ML",  
+    #data = data.fit2, 
+    #bootstrap = nBoots,
+    #se = "boot",
+    #auto.fix.first = FALSE,
+    #std.lv = TRUE,
+    #missing = "fiml",
+    #fixed.x = FALSE 
+  #)
+  #models[['fit.po05']] <- fit.po05
+#}
+
+#lavaan::summary(fit.po05,standardized = TRUE)
+
+```
+
+```{r}
+#fit.po05.est  <-standardizedSolution(fit.po05)
+#fit.po05.est %>% select(-label,se,z) %>%  filter(op == '~')
+```
+
+```{r}
+#fit.po05.est %>% 
+  #filter(abs(est.std) < 0.3) %>% 
+  #select(-z,-se) %>% 
+  #arrange(abs(est.std)) %>% 
   # Just the loadings
-  filter(op != '~1')
-
-```
-
-```{r}
-lavaanPlot(
-  model = fit.po05, 
-   labels = labelsforplotting,
-  node_options = list(shape = "box", fontname = "Helvetica"), 
-  edge_options = list(color = "grey"), 
-  coefs = TRUE, 
-  stand = TRUE,
-  sig=0.001
-  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem05.pdf"))
+  #filter(op != '~1')
+
+```
+
+```{r}
+#lavaanPlot(
+  #model = fit.po05, 
+   #labels = labelsforplotting,
+  #node_options = list(shape = "box", fontname = "Helvetica"), 
+  #edge_options = list(color = "grey"), 
+  #coefs = TRUE, 
+  #stand = TRUE,
+  #sig=0.001
+  #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem05.pdf"))
 

### 2025-06-10 manuscript_code.Rmd

@@
 library(naniar)
+
 library(gridExtra)
@@
 #library(tidySEM) # https://github.com/yrosseel/lavaan/issues/359 lavaan does not work if tidySEM is loaded!
-#library(griExtra) #used for combining multiple plots
-
 # renv::status() to check 
@@
 eval=FALSE
+
 gg_miss_var(
@@
 ```{r}
-eval=FALSE
+
+eval=FALSE
+
 gg_miss_upset(data.fit2,nsets = 10,nintersects=40)
@@
 ```{r}
-eval=FALSE
+
+eval=FALSE
+
 gg_miss_case_cumsum(
@@
 ```{r}
-eval=FALSE
+
+eval=FALSE
+
 gg_miss_fct(
@@
 ```{r}
-eval=FALSE
+
+eval=FALSE
+
 data.fit3 <- data.fit2 %>%
@@
 ```{r}
+
 #making bd02 the only variable in donation history
@@
 ```{r}
-eval=FALSE
+
+eval=FALSE
+
 #https://stats.stackexchange.com/questions/340857/serial-mediation-in-r-how-to-setup-the-model
@@
 ```
+
 ```{r}
@@
 ```{r}
+
 eval=FALSE
@@
 ```{r}
+
 eval=FALSE
 #lavaan::inspect(fit.po02ss,"r2")
+
 ```
@@
 #fit.po02ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+
 ```
@@
 ```{r}
+
 labelsforplottingnull <- c(
@@
 ```{r}
-
 nBoots <- 1000
@@
 #fit.po04ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+
 ```
@@
 #fit.po05ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+
 ```
@@
 blooddonor =~ bd02
+
 #bd01 is just used for exclusion of "no"
@@
 #lavaan::inspect(fit.po02,"r2")
+
 ```
@@
 #fit.po02.est %>% select(-label,se,z) %>%  filter(op == '~')
+
 ```
@@
   #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))
+
 ```
@@
 #fit.po03.best %>% select(-se,-z) %>% filter(op == "~")
+
 ```
@@
 factor_scores <- lavPredict(fit.po01.cfa, type = "lv")
+
+dim(factor_scores)
+dim(data.fit2)
+
+factor_scores <- lavPredict(fit.po04, type = "lv")
 head(factor_scores)
@@
 #knitr::knit_exit()
+=======
+  de02 = data.fit2$de02,
+  de01 = data.fit2$de01
+)
+
+ggplot(df_scores, aes(x = psycown, y = intention,
+                      color = as.factor(de01))) +
+  geom_point(alpha = 0.7, size = 2.5) +
+  labs(title = "Latent Factor Scores by de01 and de02",
+       x = "Psychological Ownership",
+       y = "Donation Intention",
+       color = "gender") +
+  theme_minimal()
+
+ggplot(df_scores, aes(x = p, y = intention,
+                      color = as.factor(de01))) +
+  geom_point(alpha = 0.7, size = 2.5) +
+  labs(title = "Latent Factor Scores by de01 and de02",
+       x = "Psychological Ownership",
+       y = "Donation Intention",
+       color = "gender") +
+  theme_minimal()
+
+ggplot(df_scores, aes(x = psycown, y = intention,
+                      color = as.factor(de02),
+                      shape = as.factor(de01))) +
+  geom_point(alpha = 0.7, size = 2.5) +
+  labs(title = "Latent Factor Scores by de01 and de02",
+       x = "Psychological Ownership",
+       y = "Donation Intention",
+       color = "age",
+       shape = "gender") +
+  theme_minimal()
+
+>>>>>>> ee3050c55c3e75c3dd19c1a8d96fa7c2f0d5d570
 
@@
 #fit.po05.est %>% select(-label,se,z) %>%  filter(op == '~')
+
 ```

### 2025-06-10 manuscript_code.Rmd
multigroup analysis

@@
   geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor scores: ownership x identity",
+  labs(title = "Factor scores: psycown ~ blooddonor",
+       x = "Ownership",
+       y = "Donation history",
+       color = "Age",
+       linetype = "Gender") +
+  theme_minimal()
+
+#selfid ~ blooddonor
+factor2 <- ggplot(df_scores, aes(x = selfid, y = blooddonor)) +
+  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
+  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
+  labs(title = "Factor scores: selfid ~ blooddonor",
        x = "Identity",
@@
 
-#selfid ~ blooddonor
-factor2 <- ggplot(df_scores, aes(x = selfid, y = blooddonor)) +
-  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
-  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor scores: ownership x identity",
-       x = "Identity",
-       y = "Donation history",
-       color = "Age",
-       linetype = "Gender") +
-  theme_minimal()
-
 #selfid ~ psycown
@@
   geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor scores: ownership x identity",
-       x = "Psychological Ownership",
+  labs(title = "Factor scores: selfid ~ psycown",
+       x = "Ownership",
        y = "Identity",
@@
   geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor scores: ownership x identity",
+  labs(title = "Factor scores: intention ~ blooddonor",
        x = "Intention",
@@
                       color = as.factor(de01))) +
-  geom_point(alpha = 0.7, size = 2.5) +
-  geom_smooth() +
-  labs(title = "Factor Scores: ownership x intention",
-       x = "Psychological Ownership",
+  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
+  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
+  labs(title = "Factor Scores: psycown ~ intention",
+       x = "Ownership",
        y = "Intention",
@@
 #intention ~ selfid
-factor6 <- ggplot(df_scores, aes(x = psycown, y = intention,
+factor6 <- ggplot(df_scores, aes(x = selfid, y = intention,
                       color = as.factor(de01))) +
-  geom_point(alpha = 0.7, size = 2.5) +
-  geom_smooth() +
-  labs(title = "Factor Scores: ownership x intention",
-       x = "Psychological Ownership",
+  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
+  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
+  labs(title = "Factor Scores: selfid ~ intention",
+       x = "Identity",
        y = "Intention",
@@
 selfid ~ c(a2g1, a2g2) * blooddonor + c(d21g1, d21g2) * psycown
-intention ~ c(cpg1, cpg2) * blooddonor + c(b1g1, b2g2) * psycown + c(b2g1, b2g2) * selfid
+intention ~ c(cpg1, cpg2) * blooddonor + c(b1g1, b1g2) * psycown + c(b2g1, b2g2) * selfid
 
@@
   auto.fix.first = FALSE,
-  group.equal = c("loadings","intercepts”)
-)
+  group.equal = c('loadings','intercepts')
+  )
+
+summary(ea2.free)
 
@@
   auto.fix.first = FALSE,
-group.equal = c("loadings", ”intercepts”, ”regressions”),
+  group.equal = c('loadings', 'intercepts', 'regressions')
 )
@@
 # this is for model comparison
+
 lavTestLRT(ea2.free, ea2.fixed)
@@
 #knitr::knit_exit()
-=======
-  de02 = data.fit2$de02,
-  de01 = data.fit2$de01
-)
-
-ggplot(df_scores, aes(x = psycown, y = intention,
-                      color = as.factor(de01))) +
-  geom_point(alpha = 0.7, size = 2.5) +
-  labs(title = "Latent Factor Scores by de01 and de02",
-       x = "Psychological Ownership",
-       y = "Donation Intention",
-       color = "gender") +
-  theme_minimal()
-
-ggplot(df_scores, aes(x = p, y = intention,
-                      color = as.factor(de01))) +
-  geom_point(alpha = 0.7, size = 2.5) +
-  labs(title = "Latent Factor Scores by de01 and de02",
-       x = "Psychological Ownership",
-       y = "Donation Intention",
-       color = "gender") +
-  theme_minimal()
-
-ggplot(df_scores, aes(x = psycown, y = intention,
-                      color = as.factor(de02),
-                      shape = as.factor(de01))) +
-  geom_point(alpha = 0.7, size = 2.5) +
-  labs(title = "Latent Factor Scores by de01 and de02",
-       x = "Psychological Ownership",
-       y = "Donation Intention",
-       color = "age",
-       shape = "gender") +
-  theme_minimal()
-
->>>>>>> ee3050c55c3e75c3dd19c1a8d96fa7c2f0d5d570
 

### 2025-06-10 manuscript_code.Rmd
multigroup analysis

@@
 ---
-title: "Appendix 2: Manuscript and Exploratory Code"
+title: 'Appendix 2: Manuscript and Exploratory Code'
 author: "Abigail Edwards, Lauri Hietajärvi and Mikko Arvas"
@@
 output:
+  html_document:
+    toc: true
+    df_print: paged
   pdf_document:
-    toc: yes
-
+    toc: true
 always_allow_html: true
@@
 
-```{r}
+```{r eval = FALSE}
 data %>% filter(bd04 < 0)
@@
 
-```{r}
+```{r eval = FALSE}
 
@@
 
-```{r}
+```{r eval=FALSE}
 
@@
 
-```{r}
+```{r eval=FALSE}
 gglikert(
@@
 
-```{r}
-eval=FALSE
+```{r eval=FALSE}
 
@@
 
-```{r}
-
-eval=FALSE
+```{r eval=FALSE}
 
@@
 
-eval=FALSE
+include=FALSE
 
@@
 
-```{r}
+```{r echo = FALSE}
 #Bivariate correlation table for paper
@@
 
-### a) Sum score model with robust estimator and listwise deletion
-
-```{r}
-
-eval=FALSE
+# a) Sum score model with robust estimator and listwise deletion
+
+```{r echo = FALSE}
+
 #if (fitmodels) {
@@
 
-```{r}
-
-eval=FALSE
+```{r echo = FALSE}
+
 #lavaan::inspect(fit.po02ss,"r2")
@@
 
-```{r}
-eval=FALSE
+```{r echo = FALSE}
+
 #fit.po02ss.est  <-standardizedSolution(fit.po02ss)
@@
 
-###  c) Sum score model with robust estimator and fiml
-
-```{r}
-eval=FALSE
+#  c) Sum score model with robust estimator and fiml
+
+```{r echo = FALSE}
 #if (fitmodels) {
@@
 
-```{r}
-eval=FALSE
+```{r echo = FALSE}
+
 #fit.po04ss.est  <-standardizedSolution(fit.po04ss)
@@
 
-```{r}
-eval=FALSE
+```{r echo = FALSE}
+
 #fit.po04ss.est %>% 
@@
 
-```{r}
-eval=FALSE
+```{r echo = FALSE}
+
 #lavaanPlot(
@@
 
-###  d) Sum score model with bootstrapping and fiml
-
-```{r}
-eval=FALSE
+#  d) Sum score model with bootstrapping and fiml
+
+```{r echo = FALSE}
+
 #if (fitmodels) {
@@
 
-```{r}
-eval=FALSE
+```{r echo = FALSE}
+
 #fit.po05ss.est  <-standardizedSolution(fit.po05ss)
@@
 
-```{r}
-eval=FALSE
+```{r echo = FALSE}
+
 #fit.po05ss.est %>% 
@@
 
-```{r}
-
-eval=FALSE
+```{r echo = FALSE}
+
 #lavaanPlot(
@@
 
-### e) Latent factor model with robust estimator and listwise deletion
-
-```{r}
-eval=FALSE
+# e) Latent factor model with robust estimator and listwise deletion
+
+```{r echo = FALSE}
 #if (fitmodels) {
@@
 
-```{r}
-eval=FALSE
+```{r echo = FALSE}
+
 #lavaan::inspect(fit.po02,"r2")
@@
 
-```{r}
-eval=FALSE
+```{r echo = FALSE}
+
 #fit.po02.est  <-standardizedSolution(fit.po02)
@@
 
-```{r}
-eval=FALSE
+```{r echo = FALSE}
+
 #fit.po02.est %>% 
@@
 
-```{r}
+```{r echo = FALSE}
 labelsforplotting <- c(
@@
 
-### f) Latent factor model with boostrapping and listwise deletion
-
-```{r}
+# f) Latent factor model with boostrapping and listwise deletion
+
+```{r echo = FALSE}
 
@@
 
-```{r}
+```{r echo = FALSE}
 #fit.po03.best <- parameterEstimates(fit.po03)
@@
 
-```{r}
+```{r echo = FALSE}
 #lavaanPlot(
@@
 
-#Exploratory Analysis (a1*b1 comparison)
+### Exploratory Analysis (a1*b1 comparison)
 
@@
 
-#exploratory analysis (multigroup analysis to explore impact of gender)
+### exploratory analysis (multigroup analysis to explore impact of gender)
 
@@
 #end of exploratory analysis
-#knitr::knit_exit()
+knitr::knit_exit()
 
@@
 
-```{r}
+```{r echo = FALSE}
 
@@
 
-```{r}
+```{r echo = FALSE}
 #fit.po05.est  <-standardizedSolution(fit.po05)
@@
 
-```{r}
+```{r echo = FALSE}
 #fit.po05.est %>% 

### 2025-06-11 preregistration_code.Rmd
changing labels for bd04

@@
 ```{r}
-
 table1 <-
@@
       DonationCountLast2years = bd03,
-      DaysFromLastDonation = bd04 
+      WeeksFromLastDonation = bd04 
     ),
@@
       DonationEligibility,
-      DaysFromLastDonation,
+      WeeksFromLastDonation,
       DonationCount,

### 2025-06-11 manuscript_code.Rmd
additional exploratory analysis, fixing bd02 mistake

@@
 output:
+  pdf_document:
+    toc: true
   html_document:
@@
     df_print: paged
-  pdf_document:
-    toc: true
 always_allow_html: true
@@
 
-```{r eval = FALSE}
+```{r}
 data %>% filter(bd04 < 0)
@@
 )
+
+data <- data %>% 
+  filter(bd02 >= 1)
+
+data <- data %>% 
+  filter(bd03 >= 1)
 
@@
 
-```{r eval = FALSE}
+```{r}
 
@@
 
+## exploratory demographics table
+
+```{r}
+library(gtsummary)
+library(dplyr)
+library(kableExtra)
+
+edata_clean <- data %>%
+  filter(de01 == 'man' | de01 == "woman") %>%
+  rename(
+    Gender = de01,
+    Age = de02,
+    DonationEligibility = bd01,
+    DonationCount = bd02,
+    DonationCountLast2years = bd03,
+    WeeksFromLastDonation = bd04
+  )
+
+# Create summary table
+etable1 <- tbl_summary(
+  data = edata_clean,
+  include = c(Age, DonationEligibility, WeeksFromLastDonation, DonationCount, DonationCountLast2years),
+  by = Gender,
+  statistic = list(
+    all_continuous() ~ "{mean} ({sd})",   # Use mean (SD) for continuous vars
+    all_categorical() ~ "{n} ({p}%)"
+  )
+) %>%
+  add_n()  # Adds total N for each group
+
+# Output as LaTeX
+as_kable_extra(etable1, format = "latex")
+
+```
+
 ## Age by sex
@@
 ```
+##unstandarsied means for reporting
+
+```{r}
+
+data2 <- data %>%
+mutate(across(c(bd02, bd03, bd04), as.numeric))
+
+#lavaan does not understand empty factor levels
+data.fit.stan <- data2 %>%
+  #lavaan cannot deal with unordered factor with more then 2 levels
+  filter(de01 == 'man' | de01 == "woman") %>% # EXCLUSION
+  filter(bd01 != 'no') %>% # EXCLUSION
+  #Original paper used sum scores instead of FA. 
+  mutate(
+    ss_psycown = 
+      across(starts_with("po"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE),
+    ss_selfid = 
+      across(starts_with("si"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE),
+    ss_intention = 
+      across(starts_with("it"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE),
+    ss_control = 
+      across(starts_with("ac"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE),
+    ss_knowledge = 
+      across(starts_with("ai"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE),
+    ss_investment = 
+      across(starts_with("iv"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE)
+  )
+
+#descriptives
+
+data.fit.stan %>%
+  summarise(across(
+    c(de02, ss_control, ss_knowledge, ss_investment, bd03,
+      ss_psycown, ss_selfid, ss_intention),
+    list(mean = ~mean(., na.rm = TRUE),
+         sd = ~sd(., na.rm = TRUE))
+  ))
+
+```
+```
 
@@
 isfactor <- data %>%  summarise(across(everything(), ~ is.factor(.x))) %>%  as.logical() %>% which()
-
-data <- data %>% 
-  filter(bd02 >= 1)
-
-data <- data %>% 
-  filter(bd03 >= 1)
 
@@
       scale() %>% as.numeric(),
-    #ss_donation as only measuring bd02
+    #ss_donation as only measuring bd03
     ss_donation = 
-      across(all_of(c("bd02")), ~as.numeric(.)) %>%
+      across(all_of(c("bd03")), ~as.numeric(.)) %>%
       rowSums(na.rm = TRUE)
@@
 corrplot(cor(data.fit2,use="pairwise.complete.obs"))
+
+```
+
+## Exploratory analysis: missingness on ei01 variable
+
+```{r}
+#, labels = c('miss', 'not'), ordered = TRUE
+
+data.fit.miss <- data.fit2 %>%
+  mutate(
+    ss_control = 
+      across(starts_with("ac"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) %>% 
+      scale() %>% as.numeric(),
+    ss_knowledge = 
+      across(starts_with("ai"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) %>% 
+      scale() %>% as.numeric(),
+    ss_investment = 
+      across(starts_with("iv"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) %>% 
+      scale() %>% as.numeric(),
+    )
+
+data.fit.miss$ie01bin <- ifelse(is.na(data.fit2$ie01), 1, 2)
+data.fit.miss$ie01bin <- factor(data.fit.miss$ie01bin, levels = c(1, 2), labels = c("miss", "not"), ordered = TRUE)
+summary(data.fit.miss)
+
+data.fit.miss <- data.fit.miss %>% 
+  mutate(
+    ie01bin = as.ordered(ie01bin)
+  ) %>% # and neither any factors at all
+mutate(
+  across(everything(),as.numeric)
+)  # but this also good for plotting correlations
+
+data.fit.miss <- data.fit.miss %>%
+  select(
+        #ie01bin,
+         #ac04,
+         de01,
+         de02,
+         #bd02,
+         #bd04,
+         ss_control,
+         ss_knowledge,
+         ss_investment,
+         bd03,
+         ss_psycown,
+         ss_selfid, 
+         ss_intention)
+
+library(psych)
+corrplot(cor(data.fit.miss,use="pairwise.complete.obs"), method ='number')
+
+cor_test <- corr.test(data.fit.miss, use = "pairwise", method = "pearson")
+
+# Combine correlation and p-values into APA-style strings
+apa_matrix <- matrix("", nrow=ncol(data.fit.miss), ncol=ncol(data.fit.miss))
+rownames(apa_matrix) <- colnames(data.fit.miss)
+colnames(apa_matrix) <- colnames(data.fit.miss)
+
+for (i in 1:ncol(data.fit.miss)) {
+  for (j in 1:ncol(data.fit.miss)) {
+    r_val <- round(cor_test$r[i, j], 2)
+    p_val <- cor_test$p[i, j]
+
+    # Format p-value with APA rules
+    p_str <- ifelse(p_val < .001, "< .001", paste0("= ", round(p_val, 3)))
+
+    # Combine r and p
+    apa_matrix[i, j] <- paste0(r_val, ", _p_ ", p_str)
+  }
+}
+
+# View table
+apa_matrix
+
 ```
@@
 
-#making bd02 the only variable in donation history
+#making bd03 the only variable in donation history
 
@@
                 iv =~ iv01 + iv02 + iv03
-                bd =~ bd02
+                bd =~ bd03
 
@@
 ind_eff := a1* d21 * b2 
+ind_eff1 := a1*b1
+ind_eff2 := a2*b2
 '
@@
 intention =~ it01 + it02 + it03
-blooddonor =~ bd02
+blooddonor =~ bd03
 
@@
 ind_eff := a1* d21 * b2
+ind_eff1 := a1*b1
+ind_eff2 := a2*b2
 '
@@
 
-### Exploratory Analysis (a1*b1 comparison)
+# Exploratory Analysis 
+## Factor Scores: a1*b1 comparisos
 
@@
 
-### exploratory analysis (multigroup analysis to explore impact of gender)
+# Multigroup analysis to explore impact of gender
 
@@
 
-#defining new model that doesn't contained de01
-
-ea2.model <- ' 
+#Lauri email code
+
+ea2.free <- '
 #measurement model
@@
 
-ea2.free <- sem(
-  ea2.model,
+ea2.free.fit <- sem(
+  ea2.free,
   data = data.fit2,
@@
 
-summary(ea2.free)
-
-ea2.fixed <- sem(
-  ea2.model,
+ea2.fixed <- '
+#measurement model
+psycown =~ po01 + po02 + po03
+selfid =~ si01 + si02 + si03
+intention =~ it01 + it02 + it03
+blooddonor =~ bd03
+
+# regressions (separate labels per group)
+psycown ~ a1 * blooddonor
+selfid ~ a2 * blooddonor + d21 * psycown
+intention ~ cp * blooddonor + b1 * psycown + b2 * selfid
+
+# indirect effects
+ind_eff := a1 * d21 * b2
+ind_eff1:= a1 * b1
+ind_eff2:= a2 * b2
+'
+
+ea2.fixed.fit <- sem(
+  ea2.fixed,
   data = data.fit2,
@@
   auto.fix.first = FALSE,
-  group.equal = c('loadings', 'intercepts', 'regressions')
+  group.equal = c('loadings', 'intercepts')
 )
@@
 # this is for model comparison
-
-lavTestLRT(ea2.free, ea2.fixed)
-
-#ea2 <- sem(
-  #ea2.model,
-  #data = data.fit2,
-  #group = "de01",
-  #estimator = "MLR",
-  #std.lv = TRUE,
-  #missing = "fiml",
-  #fixed.x = FALSE,
-  #auto.fix.first = FALSE
-#)
-
-#summary(ea2, standardized = TRUE)
+lavTestLRT(ea2.free.fit, ea2.fixed.fit)
 

### 2025-06-13 modelcomparison_code.Rmd
Adding additional file for model comparison with Australia data

--- 2025-06-10 manuscript_code.Rmd
@@
 
+data <- data %>% 
+  filter(bd02 >= 1)
+
+data <- data %>% 
+  filter(bd03 >= 1)
+
 #afterdata changing negative blood donation history to missing data for bd04 from 3 participants
@@
 
+## exploratory demographics table
+
+```{r}
+library(gtsummary)
+library(dplyr)
+library(kableExtra)
+
+edata_clean <- data %>%
+  filter(de01 == 'man' | de01 == "woman") %>%
+  rename(
+    Gender = de01,
+    Age = de02,
+    DonationEligibility = bd01,
+    DonationCount = bd02,
+    DonationCountLast2years = bd03,
+    WeeksFromLastDonation = bd04
+  )
+
+# Create summary table
+etable1 <- tbl_summary(
+  data = edata_clean,
+  include = c(Age, DonationEligibility, WeeksFromLastDonation, DonationCount, DonationCountLast2years),
+  by = Gender,
+  statistic = list(
+    all_continuous() ~ "{mean} ({sd})",   # Use mean (SD) for continuous vars
+    all_categorical() ~ "{n} ({p}%)"
+  )
+) %>%
+  add_n()  # Adds total N for each group
+
+# Output as LaTeX
+as_kable_extra(etable1, format = "latex")
+
+```
+
 ## Age by sex
 
-```{r eval=FALSE}
+```{r}
 
@@
 
-```{r eval=FALSE}
+```{r}
 gglikert(
@@
 
-```{r eval=FALSE}
+```{r}
 
@@
 # Model
-
 ## Format data for modelling
@@
 corrplot(cor(data.fit2,use="pairwise.complete.obs"))
-```
-
-## Missing data analysis
-
-Check proportion of missing in all variables
-
-```{r eval=FALSE}
+
+```
+
+```{r}
 
@@
 #regression
-ss_intention ~  de01 + de02 + ie01 +  cp * ss_donation + b1 * ss_psycown + b2 * ss_selfid
+ss_intention ~  cp * ss_donation + b1 * ss_psycown + b2 * ss_selfid
 
 ind_eff := a1* d21 * b2 
+ind_eff1 := a1*b1
+ind_eff2 := a2*b2
 '
@@
 
-```{r echo = FALSE}
-
-#if (fitmodels) {
-  #fit.po02ss <- sem(po02ss.model,  
-                    #estimator = "MLR", 
-                    #missing = "listwise",
-                    #data = data.fit2,
-                    #auto.fix.first = FALSE, 
-                    #std.lv = TRUE,
-                    #fixed.x = TRUE
-
- # )
-
-  #models[["fit.po02ss"]] <- fit.po02ss
-#}
-
-#lavaan::summary(fit.po02ss,standardized = TRUE)
-
-```
-
-```{r echo = FALSE}
-
-#lavaan::inspect(fit.po02ss,"r2")
-
-```
-
-```{r echo = FALSE}
-
-#fit.po02ss.est  <-standardizedSolution(fit.po02ss)
-#fit.po02ss.est %>% select(-label,se,z) %>%  filter(op == '~')
-
-```
-
-```{r echo = FALSE}
-
-#fit.po02ss.est  <-standardizedSolution(fit.po02ss)
-#fit.po02ss.est %>% 
-  #filter(abs(est.std) < 0.3) %>% 
-  #select(-z,-se) %>% 
-  #arrange(abs(est.std)) %>% 
+```{r}
+
+if (fitmodels) {
+  fit.po02ss <- sem(po02ss.model,  
+                    estimator = "MLR", 
+                    missing = "listwise",
+                    data = data.fit2,
+                    auto.fix.first = FALSE, 
+                    std.lv = TRUE,
+                    fixed.x = TRUE
+
+  )
+
+  models[["fit.po02ss"]] <- fit.po02ss
+}
+
+lavaan::summary(fit.po02ss,standardized = TRUE)
+
+```
+
+```{r}
+
+lavaan::inspect(fit.po02ss,"r2")
+
+```
+
+```{r}
+
+fit.po02ss.est  <-standardizedSolution(fit.po02ss)
+fit.po02ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+
+```
+
+```{r}
+
+fit.po02ss.est  <-standardizedSolution(fit.po02ss)
+fit.po02ss.est %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
   # Just the loadings
-  #filter(op != '~1')
+  filter(op != '~1')
 
@@
 
-#lavaanPlot(
-  #model = fit.po02ss, 
+lavaanPlot(
+  model = fit.po02ss, 
    #labels = labelsforplottingnull,
-  #node_options = list(shape = "box", fontname = "Helvetica"), 
-  #edge_options = list(color = "grey"), 
-  #coefs = TRUE, 
-  #stand = TRUE,
-  #sig=0.001
-  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
-```
-
-###  b) Sum score model with bootstrapping and listwise deletion
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
+```
+
+#  b) Sum score model with bootstrapping and listwise deletion
 
@@
   filter(abs(est.std) < 0.3) %>% 
-  select(-z,-se) %>% 
+  select(-z) %>%   
   arrange(abs(est.std)) %>% 
@@
 
+#including se
 ```
@@
 
-```{r echo = FALSE}
-#if (fitmodels) {
-  #fit.po04ss <- sem(
-    #po02ss.model,  
-    #estimator = "MLR", 
-    #data = data.fit2, 
-    #auto.fix.first = FALSE,
-    #std.lv = TRUE,
-    #missing = "fiml",
-    #fixed.x = FALSE 
-  #)
-  #models[['fit.po04ss']] <- fit.po04ss
-#}
-
-#lavaan::summary(fit.po04ss,standardized = TRUE)
-
-```
-
-```{r echo = FALSE}
-
-#fit.po04ss.est  <-standardizedSolution(fit.po04ss)
-#fit.po04ss.est %>% select(-label,se,z) %>%  filter(op == '~')
-
-```
-
-```{r echo = FALSE}
-
-#fit.po04ss.est %>% 
-  #filter(abs(est.std) < 0.3) %>% 
-  #select(-z,-se) %>% 
-  #arrange(abs(est.std)) %>% 
+```{r}
+if (fitmodels) {
+  fit.po04ss <- sem(
+    po02ss.model,  
+    estimator = "MLR", 
+    data = data.fit2, 
+    auto.fix.first = FALSE,
+    std.lv = TRUE,
+    missing = "fiml",
+    fixed.x = FALSE 
+  )
+  models[['fit.po04ss']] <- fit.po04ss
+}
+
+lavaan::summary(fit.po04ss,standardized = TRUE)
+
+```
+
+```{r}
+
+fit.po04ss.est  <-standardizedSolution(fit.po04ss)
+fit.po04ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+
+```
+
+```{r}
+
+fit.po04ss.est %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
   # Just the loadings
-  #filter(op != '~1')
-
-```
-
-```{r echo = FALSE}
-
-#lavaanPlot(
-  #model = fit.po04ss, 
-   #labels = labelsforplottingnull,
-  #node_options = list(shape = "box", fontname = "Helvetica"), 
-  #edge_options = list(color = "grey"), 
-  #coefs = TRUE, 
-  #stand = TRUE,
-  #sig=0.001
-  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))
+  filter(op != '~1')
+
+```
+
+```{r}
+
+lavaanPlot(
+  model = fit.po04ss, 
+  labels = labelsforplottingnull,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))
 
@@
 
-```{r echo = FALSE}
-
-#if (fitmodels) {
-  #fit.po05ss <- sem(
-    #po02ss.model,  
-    #estimator = "ML", 
-    #bootstrap = nBoots,
-    #se = "boot",
-    #data = data.fit2, 
-    #auto.fix.first = FALSE,
-    #std.lv = TRUE,
-    #missing = "fiml",
-    #fixed.x = FALSE 
-  #)
-  #models[['fit.po05ss']] <- fit.po05ss
-#}
-
-#lavaan::summary(fit.po05ss,standardized = TRUE)
-
-```
-
-```{r echo = FALSE}
-
-#fit.po05ss.est  <-standardizedSolution(fit.po05ss)
-#fit.po05ss.est %>% select(-label,se,z) %>%  filter(op == '~')
-
-```
-
-```{r echo = FALSE}
-
-#fit.po05ss.est %>% 
-  #filter(abs(est.std) < 0.3) %>% 
-  #select(-z,-se) %>% 
-  #arrange(abs(est.std)) %>% 
+```{r}
+
+if (fitmodels) {
+  fit.po05ss <- sem(
+    po02ss.model,  
+    estimator = "ML", 
+    bootstrap = nBoots,
+    se = "boot",
+    data = data.fit2, 
+    auto.fix.first = FALSE,
+    std.lv = TRUE,
+    missing = "fiml",
+    fixed.x = FALSE 
+  )
+  models[['fit.po05ss']] <- fit.po05ss
+}
+
+lavaan::summary(fit.po05ss,standardized = TRUE)
+
+```
+
+```{r}
+
+fit.po05ss.est  <-standardizedSolution(fit.po05ss)
+fit.po05ss.est %>% select(-label,se,z) %>%  filter(op == '~')
+
+```
+
+```{r}
+
+fit.po05ss.est %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
   # Just the loadings
-  #filter(op != '~1')
-
-```
-
-```{r echo = FALSE}
-
-#lavaanPlot(
-  #model = fit.po05ss, 
-   #labels = labelsforplottingnull,
-  #node_options = list(shape = "box", fontname = "Helvetica"), 
-  #edge_options = list(color = "grey"), 
-  #coefs = TRUE, 
-  #stand = TRUE,
-  #sig=0.001
-  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem05.pdf"))
+  filter(op != '~1')
+
+```
+
+```{r}
+
+lavaanPlot(
+  model = fit.po05ss, 
+  labels = labelsforplottingnull,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem05.pdf"))
 
@@
 intention =~ it01 + it02 + it03
-blooddonor =~ bd02
+donation =~ bd03
 
@@
 #mediation
-psycown ~ a1 *  blooddonor
-selfid ~ a2 * blooddonor + d21 * psycown
+psycown ~ a1 *  donation
+selfid ~ a2 * donation + d21 * psycown
 
 #regression
-intention ~  de01 + de02 + ie01 +  cp * blooddonor + b1 * psycown + b2 * selfid
+intention ~  cp * donation + b1 * psycown + b2 * selfid
 ind_eff := a1* d21 * b2
+ind_eff1 := a1*b1
+ind_eff2 := a2*b2
 '
@@
 
-```{r echo = FALSE}
-#if (fitmodels) {
-  #fit.po02 <- sem(po02.model,  
-                  #estimator = "MLR",
-                  #missing = "listwise",
-                  #data = data.fit2,
-                  #auto.fix.first = FALSE,
-                  #std.lv = TRUE,
-                  #fixed.x = TRUE
-  #)
-  #models[["fit.po02"]] <- fit.po02
-#}
-#lavaan::summary(fit.po02,standardized = TRUE)
-
-```
-
-```{r echo = FALSE}
-
-#lavaan::inspect(fit.po02,"r2")
-
-```
-
-```{r echo = FALSE}
-
-#fit.po02.est  <-standardizedSolution(fit.po02)
-#fit.po02.est %>% select(-label,se,z) %>%  filter(op == '~')
-
-```
-
-```{r echo = FALSE}
-
-#fit.po02.est %>% 
-  #filter(abs(est.std) < 0.3) %>% 
-  #select(-z,-se) %>% 
-  #arrange(abs(est.std)) %>% 
+```{r}
+if (fitmodels) {
+  fit.po02 <- sem(po02.model,  
+                  estimator = "MLR",
+                  missing = "listwise",
+                  data = data.fit2,
+                  auto.fix.first = FALSE,
+                  std.lv = TRUE,
+                  fixed.x = TRUE
+  )
+  models[["fit.po02"]] <- fit.po02
+}
+lavaan::summary(fit.po02,standardized = TRUE)
+
+```
+
+```{r}
+
+lavaan::inspect(fit.po02,"r2")
+
+```
+
+```{r}
+
+fit.po02.est  <-standardizedSolution(fit.po02)
+fit.po02.est %>% select(-label,se,z) %>%  filter(op == '~')
+
+```
+
+```{r}
+
+fit.po02.est %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
   # Just the loadings
-  #filter(op != '~1')
-
-```
-
-```{r echo = FALSE}
+  filter(op != '~1')
+
+```
+
+```{r}
 labelsforplotting <- c(
@@
   "intention" = "Intention",
-  "blooddonor" = "Blood donation history"
-)
-
-#lavaanPlot(
-  #model = fit.po02, 
-   #labels = labelsforplotting,
-  #node_options = list(shape = "box", fontname = "Helvetica"), 
-  #edge_options = list(color = "grey"), 
-  #coefs = TRUE, 
-  #stand = TRUE,
-  #sig=0.001
-  #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))
+  "donation" = "Blood donation history"
+)
+
+lavaanPlot(
+  model = fit.po02, 
+  labels = labelsforplotting,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))
 
@@
 
-```{r echo = FALSE}
-
-#if (fitmodels) {
-  #fit.po03 <- sem(po02.model,  
-                  #estimator = "ML", 
-                  #data = data.fit2,
-                  #bootstrap = nBoots,
-                  #se = "boot",
-                  #missing = "listwise",
-                  #auto.fix.first = FALSE,
-                  #std.lv = TRUE,
-                  #fixed.x =TRUE
-  #)
- # models[['fit.po03']] <- fit.po03
-#}
-
-#lavaan::summary(fit.po03,standardized = TRUE)
-
-```
-
-```{r echo = FALSE}
-#fit.po03.best <- parameterEstimates(fit.po03)
-#fit.po03.best %>% select(-se,-z) %>% filter(op == "~")
-
-```
-
-```{r echo = FALSE}
-#lavaanPlot(
-  #model = fit.po03, 
-   #labels = labelsforplotting,
-  #node_options = list(shape = "box", fontname = "Helvetica"), 
-  #edge_options = list(color = "grey"), 
-  #coefs = TRUE, 
-  #stand = TRUE,
-  #sig=0.001
-  #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem03.pdf"))
+```{r}
+
+if (fitmodels) {
+  fit.po03 <- sem(po02.model,  
+                  estimator = "ML", 
+                  data = data.fit2,
+                  bootstrap = nBoots,
+                  se = "boot",
+                  missing = "listwise",
+                  auto.fix.first = FALSE,
+                  std.lv = TRUE,
+                  fixed.x =TRUE
+  )
+  models[['fit.po03']] <- fit.po03
+}
+
+lavaan::summary(fit.po03,standardized = TRUE)
+
+```
+
+```{r}
+fit.po03.best <- parameterEstimates(fit.po03)
+fit.po03.best %>% select(-se,-z) %>% filter(op == "~")
+
+```
+
+```{r}
+lavaanPlot(
+  model = fit.po03, 
+  labels = labelsforplotting,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem03.pdf"))
 
@@
 
-### Exploratory Analysis (a1*b1 comparison)
-
-```{r}
-
-#factor_scores <- lavPredict(fit.po04, type = "lv") #taking factor scores from CFA 
-factor_scores <- lavPredict(fit.po01.cfa, type = "lv")
-
-dim(factor_scores)
-dim(data.fit2)
-
-factor_scores <- lavPredict(fit.po04, type = "lv")
-head(factor_scores)
-summary(factor_scores)
-
-df_scores <- cbind(
-  as.data.frame(factor_scores),
-  de02 = data.fit2$de02, #de02 = age
-  de01 = data.fit2$de01 #de01 = gender
-)
-
-df_scores$de01 <- factor(df_scores$de01, levels = c(1, 2), labels = c("man", "woman"), ordered = TRUE)
-
-#psycown ~ blooddonor
-factor1 <- ggplot(df_scores, aes(x = psycown, y = blooddonor)) +
-  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
-  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor scores: psycown ~ blooddonor",
-       x = "Ownership",
-       y = "Donation history",
-       color = "Age",
-       linetype = "Gender") +
-  theme_minimal()
-
-#selfid ~ blooddonor
-factor2 <- ggplot(df_scores, aes(x = selfid, y = blooddonor)) +
-  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
-  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor scores: selfid ~ blooddonor",
-       x = "Identity",
-       y = "Donation history",
-       color = "Age",
-       linetype = "Gender") +
-  theme_minimal()
-
-#selfid ~ psycown
-factor3 <- ggplot(df_scores, aes(x = psycown, y = selfid)) +
-  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
-  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor scores: selfid ~ psycown",
-       x = "Ownership",
-       y = "Identity",
-       color = "Age",
-       linetype = "Gender") +
-  theme_minimal()
-
-#intention ~ blooddonor
-factor4 <- ggplot(df_scores, aes(x = intention, y = blooddonor,
-                      color = as.factor(de01))) +
-  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
-  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor scores: intention ~ blooddonor",
-       x = "Intention",
-       y = "Donation History",
-       color = "Age",
-       linetype = "Gender") +
-  theme_minimal()
-
-#intention ~ psycown
-factor5 <- ggplot(df_scores, aes(x = psycown, y = intention,
-                      color = as.factor(de01))) +
-  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
-  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor Scores: psycown ~ intention",
-       x = "Ownership",
-       y = "Intention",
-       color = "Age",
-       linetype = "Gender") +
-  theme_minimal()
-
-#intention ~ selfid
-factor6 <- ggplot(df_scores, aes(x = selfid, y = intention,
-                      color = as.factor(de01))) +
-  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
-  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
-  labs(title = "Factor Scores: selfid ~ intention",
-       x = "Identity",
-       y = "Intention",
-       color = "Age",
-       linetype = "Gender") +
-  theme_minimal()
-
-combined <- grid.arrange(factor1, factor2, factor3, factor4, factor5, factor6, ncol = 2)
-
-ggsave(
-  "Ownership and Intention Factor Loadings",
-  combined, width = 20, 
-  height = 10, 
-  device = "pdf")
-
-```
-
-### exploratory analysis (multigroup analysis to explore impact of gender)
-
-```{r}
-
-#making gender a factor and ordering
-
-data.fit2$de01 <- factor(data.fit2$de01, levels = c(1, 2), labels = c("man", "woman"), ordered = TRUE)
-
-data.fit2 %>%
-  group_by(de01) %>%
-  summarise(across(everything(), ~ var(.x, na.rm = TRUE)))
-
-#defining new model that doesn't contained de01
-
-ea2.model <- ' 
-#measurement model
-psycown =~ po01 + po02 + po03
-selfid =~ si01 + si02 + si03
-intention =~ it01 + it02 + it03
-blooddonor =~ bd02
-
-# regressions (separate labels per group)
-psycown ~ c(a1g1, a1g2) * blooddonor
-selfid ~ c(a2g1, a2g2) * blooddonor + c(d21g1, d21g2) * psycown
-intention ~ c(cpg1, cpg2) * blooddonor + c(b1g1, b1g2) * psycown + c(b2g1, b2g2) * selfid
-
-#g1 indirect effects
-ind_eff_g1 := a1g1 * d21g1 * b2g1
-ind_eff1_g1:= a1g1 * b1g1
-ind_eff2_g1:= a2g1 * b2g1
-
-#g2 indirect effects
-ind_eff_g2:= a1g2 * d21g2 * b2g2
-ind_eff1_g2:= a1g2 * b1g2
-ind_eff2_g2:= a2g2 * b2g2
-'
-
-ea2.free <- sem(
-  ea2.model,
+###  h) Latent factor model with bootstrapping and fiml
+
+```{r}
+
+if (fitmodels) {
+  fit.po05 <- sem(
+    po02.model,  
+    estimator = "ML",  
     data = data.fit2, 
-  group = "de01",
-  estimator = "MLR",
+    bootstrap = nBoots,
+    se = "boot",
+    auto.fix.first = FALSE,
     std.lv = TRUE,
     missing = "fiml",
-  fixed.x = FALSE,
-  auto.fix.first = FALSE,
-  group.equal = c('loadings','intercepts')
-  )
-
-summary(ea2.free)
-
-ea2.fixed <- sem(
-  ea2.model,
-  data = data.fit2,
-  group = "de01",
-  estimator = "MLR",
-  std.lv = TRUE,
-  missing = "fiml",
-  fixed.x = FALSE,
-  auto.fix.first = FALSE,
-  group.equal = c('loadings', 'intercepts', 'regressions')
-)
-
-# this is for model comparison
-
-lavTestLRT(ea2.free, ea2.fixed)
-
-#ea2 <- sem(
-  #ea2.model,
-  #data = data.fit2,
-  #group = "de01",
-  #estimator = "MLR",
-  #std.lv = TRUE,
-  #missing = "fiml",
-  #fixed.x = FALSE,
-  #auto.fix.first = FALSE
-#)
-
-#summary(ea2, standardized = TRUE)
-
-#end of exploratory analysis
-knitr::knit_exit()
-
-```
-
-###  h) Latent factor model with bootstrapping and fiml
-
-```{r echo = FALSE}
-
-#if (fitmodels) {
-  #fit.po05 <- sem(
-    #po02.model,  
-    #estimator = "ML",  
-    #data = data.fit2, 
-    #bootstrap = nBoots,
-    #se = "boot",
-    #auto.fix.first = FALSE,
-    #std.lv = TRUE,
-    #missing = "fiml",
-    #fixed.x = FALSE 
-  #)
-  #models[['fit.po05']] <- fit.po05
-#}
-
-#lavaan::summary(fit.po05,standardized = TRUE)
-
-```
-
-```{r echo = FALSE}
-#fit.po05.est  <-standardizedSolution(fit.po05)
-#fit.po05.est %>% select(-label,se,z) %>%  filter(op == '~')
-
-```
-
-```{r echo = FALSE}
-#fit.po05.est %>% 
-  #filter(abs(est.std) < 0.3) %>% 
-  #select(-z,-se) %>% 
-  #arrange(abs(est.std)) %>% 
+    fixed.x = FALSE 
+  )
+  models[['fit.po05']] <- fit.po05
+}
+
+lavaan::summary(fit.po05,standardized = TRUE)
+
+```
+
+```{r}
+fit.po05.est  <-standardizedSolution(fit.po05)
+fit.po05.est %>% select(-label,se,z) %>%  filter(op == '~')
+
+```
+
+```{r}
+fit.po05.est %>% 
+  filter(abs(est.std) < 0.3) %>% 
+  select(-z,-se) %>% 
+  arrange(abs(est.std)) %>% 
   # Just the loadings
-  #filter(op != '~1')
-
-```
-
-```{r echo = FALSE}
-#lavaanPlot(
-  #model = fit.po05, 
-   #labels = labelsforplotting,
-  #node_options = list(shape = "box", fontname = "Helvetica"), 
-  #edge_options = list(color = "grey"), 
-  #coefs = TRUE, 
-  #stand = TRUE,
-  #sig=0.001
-  #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem05.pdf"))
+  filter(op != '~1')
+
+```
+
+```{r}
+lavaanPlot(
+  model = fit.po05, 
+   labels = labelsforplotting,
+  node_options = list(shape = "box", fontname = "Helvetica"), 
+  edge_options = list(color = "grey"), 
+  coefs = TRUE, 
+  stand = TRUE,
+  sig=0.001
+  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem05.pdf"))
 
@@
 ```{r , fig.height=8}
+
+#adding data from previous Australia study 
+##note estimates created using lavaan differ very slightly from estimates using process macro (original paper)
+manual_data <- tibble(
+  est = c(0.343, 0.382, -0.137, 0.657, 0.322, -0.024, 0.073, -0.047, 0.25),
+  ci.lower = c(0.227, 0.308, -0.215, 0.531, 0.226, -0.102, 0.035, -0.08, 0.186),
+  ci.upper = c(0.459, 0.455, -0.06, 0.782, 0.418, 0.054, 0.11, -0.014, 0.315),
+  label = c("manuala1", "manuala2", "manualb1", "manualb2", "manuald21", "manualcp", "manualeff", "manualeff1", "manualeff2"),
+  lhs = c("psycown", "selfid", "intention", "intention", "selfid", "intention", "ind_eff", "ind_eff1", "ind_eff2"),
+  rhs = c("donation", "donation", "psycown", "selfid", "psycown", "donation", "a1*d21*b2", "a1*b1", "a2*b2"),
+  op = c("~", "~", "~", "~", "~", "~", ":=", ":=", ":="),
+  term = c("a1:psycown~donation", "a2:selfid~donation", "b1:intention~psycown", "b2:intention~selfid", "d21:selfid~psycown", "cp:intention~donation", "ind_eff:=a1*d21*b2", "ind_eff1:ind_eff1:=a1*b1", "ind_eff2:ind_eff2:=a2*b2"),
+  #estmet = factor("Manual", levels = levels(data.forest$estmet)),
+  estmet = "Bootstrapped",
+  missing = "Listwise",
+  linetype = "Sum score",
+  sample = "Australia",
+  #mtype = factor("Manual-Manual-Manual", levels = levels(data.forest$mtype)),
+  #color = factor("Manual and Manual", levels = levels(data.forest$color)),
+  hollow_group = NA
+)
 
@@
     estmet = "Bootstrapped",
-    missing = "Listwise"
+    missing = "Listwise",
+    sample = "Finland"
          ) %>%  
@@
       estmet = "Robust",
-      missing = "Listwise"
+      missing = "Listwise",
+      sample = "Finland"
       ) %>% 
@@
       estmet = "Robust",
-      missing = "FIML"
+      missing = "FIML",
+      sample = "Finland"
       ) %>% 
@@
       estmet = "Boostrapped",
-      missing = "FIML"
+      missing = "FIML",
+      sample = "Finland"
       ) %>% 
@@
   arrange(estmet, term)
+
+data.forest <- bind_rows(data.forest, manual_data)
 
@@
     estmet = "Bootstrapped",
-    missing = "Listwise"
+    missing = "Listwise",
+    sample = "Finland"
     ) %>%  
@@
     term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
-    term =str_replace_all(term,"ss_donation","blooddonor:"),
-    term =str_replace_all(term,"blooddonor:","blooddonor"),
+    term =str_replace_all(term,"ss_donation","donation:"),
+    term =str_replace_all(term,"donation","donation"),
     term =str_replace_all(term,"^:",""),
@@
 
+#adding manual data
+
 pos <- position_nudge(y=as.numeric(as.factor(data.forest$mtype))/10 -0.5)
@@
 pf <- data.forest %>% 
-  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=color,linetype = linetype))+
+  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=color,linetype=linetype))+
   geom_vline(xintercept=0,color="grey") +
   geom_linerange(position=pos)+
-  geom_point(shape=21,size=3,position=pos,alpha=0.7)+
+  geom_point(aes(shape = sample),size=3,position=pos,alpha=0.7)+
   scale_fill_discrete(na.value=NA, guide ="none") + 
   labs(x="Standardised effect size",y="Relationship") +
-  labs(color="Model type", linetype="Variable type") +
+  labs(shape="Sample", color="Model type", linetype="Variable type") +
   geom_stripped_rows(color=NA) + theme_classic()

### 2025-06-16 manuscript_code.Rmd
tidying including additional metrics

@@
 library(kableExtra)
+
+#additional packages added (plotting, cronbach's alpha)
 library(naniar)
-
+library(psych)
 library(gridExtra)
+
 #library(semPlot) # update plots to semPlot later
@@
 ```
-```
 
@@
 
-library(psych)
 corrplot(cor(data.fit.miss,use="pairwise.complete.obs"), method ='number')
@@
 
+## Cronbach's Alpha
+
+```{r}
+
+scalepo <- data.fit2 %>% select(starts_with("po"))
+scaleid <- data.fit2 %>% select(starts_with("si"))
+scaleint <- data.fit2 %>% select(starts_with("it"))
+scaleco <- data.fit2 %>% select(starts_with("ac"))
+scalekn <- data.fit2 %>% select(starts_with("ai"))
+scaleiv <- data.fit2 %>% select(starts_with("iv"))
+
+alpha(scalepo)
+alpha(scaleid)
+alpha(scaleint)
+alpha(scaleco)
+alpha(scalekn)
+alpha(scaleiv)
+
+```
+
 ## Confirmatory factor analysis
@@
 
-###  b) Sum score model with bootstrapping and listwise deletion
+#  b) Sum score model with bootstrapping and listwise deletion
 
@@
 
-#end of exploratory analysis
-knitr::knit_exit()
-
 ```

### 2025-06-16 modelcomparison_code.Rmd
fixing title

@@
 ---
-title: 'Appendix 2: Manuscript and Exploratory Code'
+title: 'Appendix 4: Model Comparison Code For Manuscript'
 author: "Abigail Edwards, Lauri Hietajärvi and Mikko Arvas"
@@
         SSBL=fit.po03ss,
-        SSRF = fit.po04ss,
-        SSBL = fit.po05ss,
+        SSRF=fit.po04ss,
+        SSBL=fit.po05ss,
         LFRL=fit.po02,

### 2025-06-19 codebook_data.Rmd
closed data with open questions

@@
 ---
-title: "Codebook created on 2025-05-21 at 2025-05-21 15:18:06"
+title: "Codebook created on 2025-06-19 at 2025-06-19 09:49:51"
 output: github_document
@@
 ## Dataset description
-The data contains 634 cases and 28 variables.
+The data contains 633 cases and 29 variables.
 

### 2025-06-19 prepare_data.R

@@ -79,7 +79,7 @@ data <- data %>%
     #Days between answering and last donation, done already
     #bd04 = as.numeric(as.Date(dttma) - bd04)
     
-  )  %>% select(-X,-consent,-open)
+  )  %>% select(-X,-consent)
 
 summary(data)
 #####################################################################################################################################

### 2025-06-19 codebook_data.Rmd
manual translation error

@@
 ---
-title: "Codebook created on 2025-06-19 at 2025-06-19 09:49:51"
+title: "Codebook created on 2025-06-19 at 2025-06-19 10:42:57"
 output: github_document
@@
 ## Dataset description
-The data contains 633 cases and 29 variables.
+The data contains 634 cases and 29 variables.
 

### 2025-06-19 preregistration_code.Rmd
fix typo

@@
         SSRF = fit.po04ss,
-        SSBL = fit.po05ss,
+        SSBF = fit.po05ss,
         LFRL=fit.po02,
@@
 gfi
+
+#after data (fix typo SSBL -> SSBF)
 ```

### 2025-06-19 modelcomparison_code.Rmd
fixing typo and removing exploratory analysis

@@
   load(file=here("prereg_models.rdata"))
+
   fit.po02ss <- models$fit.po02ss
@@
 ```{r}
-library(gtsummary)
-library(dplyr)
-library(kableExtra)
-
-edata_clean <- data %>%
-  filter(de01 == 'man' | de01 == "woman") %>%
-  rename(
-    Gender = de01,
-    Age = de02,
-    DonationEligibility = bd01,
-    DonationCount = bd02,
-    DonationCountLast2years = bd03,
-    WeeksFromLastDonation = bd04
-  )
+#library(gtsummary)
+#library(dplyr)
+#library(kableExtra)
+
+#edata_clean <- data %>%
+  #filter(de01 == 'man' | de01 == "woman") %>%
+  #rename(
+    #Gender = de01,
+    #Age = de02,
+    #DonationEligibility = bd01,
+    #DonationCount = bd02,
+    #DonationCountLast2years = bd03,
+    #WeeksFromLastDonation = bd04
+  #)
 
 # Create summary table
-etable1 <- tbl_summary(
-  data = edata_clean,
-  include = c(Age, DonationEligibility, WeeksFromLastDonation, DonationCount, DonationCountLast2years),
-  by = Gender,
-  statistic = list(
-    all_continuous() ~ "{mean} ({sd})",   # Use mean (SD) for continuous vars
-    all_categorical() ~ "{n} ({p}%)"
-  )
-) %>%
-  add_n()  # Adds total N for each group
+#etable1 <- tbl_summary(
+  #data = edata_clean,
+  #include = c(Age, DonationEligibility, WeeksFromLastDonation, DonationCount, DonationCountLast2years),
+  #by = Gender,
+  #statistic = list(
+    #all_continuous() ~ "{mean} ({sd})",   # Use mean (SD) for continuous vars
+    #all_categorical() ~ "{n} ({p}%)"
+  #)
+#) %>%
+  #add_n()  # Adds total N for each group
 
 # Output as LaTeX
-as_kable_extra(etable1, format = "latex")
+#as_kable_extra(etable1, format = "latex")
 
@@
 ```
+```{r}
+summary(fit.po03ss, standardized=TRUE, rsquare=TRUE)
+lavaan::fitmeasures(fit.po03ss,c("npar","chisq","df","cfi","rmsea","srmr"))
+```
 
@@
   filter(abs(est.std) < 0.3) %>% 
-  select(-z,-se) %>% 
+  select(-z) %>% 
   arrange(abs(est.std)) %>% 
@@
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem04.pdf"))
+
+```
+
+```{r}
+summary(fit.po04, standardized=TRUE, rsquare=TRUE)
 
@@
         SSRF=fit.po04ss,
-        SSBL=fit.po05ss,
+        #typo fixed (SSBL -> SSBF)
+        SSBF=fit.po05ss,
         LFRL=fit.po02,
@@
 if (fitmodels) {
-  save(models,file=here("prereg_models.rdata"))
+  save(models,file=here("manuscript_models.rdata"))
 

### 2025-09-03 manuscript_code.Rmd
Incorporating additional Australian data

@@
 output:
-  pdf_document:
-    toc: true
   html_document:
@@
     df_print: paged
+  pdf_document:
+    toc: true
 always_allow_html: true
@@
 data.fit3 <- data.fit2 %>%
-  select(-ss_psycown,
+  select(-ie01,
+         -dttma,
+        -ss_psycown,
          -ss_selfid, 
@@
 
+sum(is.na(data.fit3))
+sum(!is.na(data.fit3))
+
 corrplot(cor(data.fit3,use="pairwise.complete.obs"))
-mcar_test(data.fit3 %>% select(-dttma))
+mcar_test(data.fit3)
 
@@
 lavTestLRT(ea2.free.fit, ea2.fixed.fit)
+
+knitr::knit_exit()
 

### 2025-09-03 modelcomparison_code.Rmd
Incorporating additional Australian data

@@
 #adding data from previous Australia study
+
+#Previously Reported Effects (Sum score, bootstrapped, listwise)
 ##note estimates created using lavaan differ very slightly from estimates using process macro (original paper)
-manual_data <- tibble(
+manual_data1 <- tibble(
   est = c(0.343, 0.382, -0.137, 0.657, 0.322, -0.024, 0.073, -0.047, 0.25),
@@
   ci.upper = c(0.459, 0.455, -0.06, 0.782, 0.418, 0.054, 0.11, -0.014, 0.315),
-  label = c("manuala1", "manuala2", "manualb1", "manualb2", "manuald21", "manualcp", "manualeff", "manualeff1", "manualeff2"),
+  label = c("manual1a1", "manual1a2", "manual1b1", "manual1b2", "manual1d21", "manual1cp", "manual1eff", "manual1eff1", "manual1eff2"),
   lhs = c("psycown", "selfid", "intention", "intention", "selfid", "intention", "ind_eff", "ind_eff1", "ind_eff2"),
@@
 
+#best practice model (latent, robust, fiml)
+manual_data2 <- tibble(
+  est = c(0.369, 0.373, -0.19, 0.825, 0.341, -0.089, 0.104, -0.07, 0.308),
+  ci.lower = c(0.25, 0.279, -0.298, 0.696, 0.22, -0.183, 0.048, -0.117, 0.23),
+  ci.upper = c(0.487, 0.467, -0.081, 0.954, 0.461, 0.004, 0.159, -0.023, 0.386),
+  label = c("manual2a1", "manual2a2", "manual2b1", "manual2b2", "manual2d21", "manual2cp", "manual2eff", "manual2eff1", "manual2eff2"),
+  lhs = c("psycown", "selfid", "intention", "intention", "selfid", "intention", "ind_eff", "ind_eff1", "ind_eff2"),
+  rhs = c("donation", "donation", "psycown", "selfid", "psycown", "donation", "a1*d21*b2", "a1*b1", "a2*b2"),
+  op = c("~", "~", "~", "~", "~", "~", ":=", ":=", ":="),
+  term = c("a1:psycown~donation", "a2:selfid~donation", "b1:intention~psycown", "b2:intention~selfid", "d21:selfid~psycown", "cp:intention~donation", "ind_eff:=a1*d21*b2", "ind_eff1:ind_eff1:=a1*b1", "ind_eff2:ind_eff2:=a2*b2"),
+  #estmet = factor("Manual", levels = levels(data.forest$estmet)),
+  estmet = "Robust",
+  missing = "FIML",
+  linetype = "Latent factor",
+  sample = "Australia",
+  #mtype = factor("Manual-Manual-Manual", levels = levels(data.forest$mtype)),
+  #color = factor("Manual and Manual", levels = levels(data.forest$color)),
+  hollow_group = NA
+)
+
 data.forest  <- fit.po03.best %>% 
@@
 
-data.forest <- bind_rows(data.forest, manual_data)
+data.forest <- bind_rows(data.forest, manual_data1, manual_data2)
 

### 2025-10-30 modelcomparison_code.Rmd
adding de01 and de02 back into model
Editing figure

@@
 #regression
-ss_intention ~  cp * ss_donation + b1 * ss_psycown + b2 * ss_selfid
+ss_intention ~  de01 + de02 + cp * ss_donation + b1 * ss_psycown + b2 * ss_selfid
 
@@
 
+```
+```{r}
+lavaan::inspect(fit.po03ss,"r2")
 ```
@@
   geom_vline(xintercept=0,color="grey") +
-  geom_linerange(position=pos)+
+  geom_linerange(aes(size = sample), position=pos) +
+  scale_size_manual(values = c("Australia" = 1.2, "Finland" = 0.8)) +
   geom_point(aes(shape = sample),size=3,position=pos,alpha=0.7)+
   scale_fill_discrete(na.value=NA, guide ="none") + 
-  labs(x="Standardised effect size",y="Relationship") +
-  labs(shape="Sample", color="Model type", linetype="Variable type") +
+  labs(x="Standardised Effect Size",y="Relationship") +
+  labs(color="Model type", linetype="Variable type", shape="Sample") + 
+  guides(size = "none") +
   geom_stripped_rows(color=NA) + theme_classic()

### 2025-10-30 modelcomparison_code.Rmd
Adding labels to figure

@@
   scale_fill_discrete(na.value=NA, guide ="none") + 
+  scale_y_discrete(
+    name = "Relationship", 
+    labels = c(
+      "ind_eff2:ind_eff2:=a2*b2" = "Indirect Effect: a2*b2", 
+      "ind_eff1:ind_eff1:=a1*b1" = "Indirect Effect: a1*b1",
+      "ind_eff:=a1*d21*b2" = "Indirect Effect: a1*d21*b1",
+      "d21:selfid~psycown" = "Identity*Ownership (d21)",
+      "cp:intention~donation" = "Intention*Donation (cp)",
+      "b2:intention~selfid" = "Intention*Identity (b2)",
+      "b1:intention~psycown" = "Intention*Owenrship (b1)",
+      "a2:selfid~donation" = "Identity*Donation (a2)",
+      "a1:psycown~donation" = "Ownership*Donation(a1)"
+    )
+  ) +
   labs(x="Standardised Effect Size",y="Relationship") +

### 2025-11-10 modelcomparison_code.Rmd
demographics and cor plot and stars

@@
 library(corrplot)
-library(ggstats)
 library(lavaan)
@@
 library(kableExtra)
-library(naniar)
-
+library(naniar) # for missing data
+library(patchwork)
 library(gridExtra)
+library(gt) # for table as plot
+library(ggpmisc) 
+library(ggcorrplot) #for corrplot
+library(ggstats) # for likert plots
+
 #library(semPlot) # update plots to semPlot later
@@
 #standarsing behaviour variables
-data <- data %>%
+data.fit <- data %>%
 mutate(across(c(bd02, bd03, bd04), ~scale(.) %>% as.numeric()))
@@
       scale() %>% as.numeric(),
+    ss_control = across(starts_with("ac"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) %>% 
+      scale() %>% as.numeric(),
+    ss_knowledge = across(starts_with("ai"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) %>% 
+      scale() %>% as.numeric(),
+    ss_selfinv = across(starts_with("iv"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) %>% 
+      scale() %>% as.numeric(),
     #ss_donation as only measuring bd03
@@
 data.fit2 <- data.fit %>% 
+  #standarsing behaviour variables
+  mutate(across(c(bd02, bd03, bd04), ~scale(.) %>% as.numeric())) %>% 
+  #remove these as they are not used in modelling
+  select(
+    -ss_control,
+    -ss_knowledge,
+    ss_selfinv
+    ) %>% 
   mutate(
@@
 
+```
+
+## Demographics with correlation
+
+```{r}
+
+tmp <- data.fit %>% 
+  select(
+        de01,
+         de02,
+         ss_control,
+         ss_knowledge,
+         ss_selfinv,
+         bd03, 
+         ss_psycown,
+         ss_selfid,
+         ss_intention 
+         ) %>%
+  select(ncol(.):1) %>% 
+  mutate(
+    across(everything(),as.numeric)
+  ) %>% 
+  rename(
+        "Gender"=de01,
+         "Age (years)" = de02,
+         "Control" = ss_control,
+         "Knowledge" =  ss_knowledge,
+         "Self-investment" = ss_selfinv,
+         "Past donations in 2 years" = bd03, 
+         "Psychological ownership" = ss_psycown,
+         "Self-identity" = ss_selfid,
+         "Intention" = ss_intention 
+           )
+
+pt2 <- ggcorrplot(
+  #compute correlatins from normalised values
+  cor(tmp,use="pairwise.complete.obs"),
+  show.diag = FALSE,
+  outline.color = "white",
+  p.mat = cor_pmat(tmp,use="pairwise.complete.obs"),
+  sig.level=0.01,
+  #insig = "",
+  lab=TRUE,
+  lab_size = 3,
+  show.legend = TRUE,
+  tl.cex = 10,
+  digits = 1
+ )
+
+#compute mean and sd on original scale
+tmp <- data %>%
+  #lavaan cannot deal with unordered factor with more then 2 levels
+  filter(de01 == 'man' | de01 == "woman") %>% # EXCLUSION
+  filter(bd01 != 'no') %>% # EXCLUSION
+  mutate(
+    across(isfactor, ~ fct_drop(.x)
+    ) %>% 
+      mutate(
+        de01 = as.numeric(de01), # for counting correlation
+       )
+  ) %>% 
+  #Original paper used sum scores instead of FA. 
+  mutate(
+    ss_psycown = 
+      across(starts_with("po"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE),
+    ss_selfid = 
+      across(starts_with("si"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) ,
+    ss_intention = 
+      across(starts_with("it"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) ,
+    ss_control = across(starts_with("ac"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) ,
+    ss_knowledge = across(starts_with("ai"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE),
+    ss_selfinv = across(starts_with("iv"),~as.numeric(.)) %>%  
+      rowMeans(na.rm = TRUE) ,
+  ) %>% select(
+    de01,de02,ss_control,ss_knowledge,ss_selfinv,bd03, ss_psycown,ss_selfid,ss_intention 
+  ) %>% pivot_longer(cols = everything(), names_to = "Variable", values_to = "val") %>%
+  group_by(Variable) %>% 
+  summarize(mean = mean(val,na.rm=TRUE), sd = sd(val,na.rm=TRUE)
+  ) %>% mutate(
+    Mean = as.character(round(mean,2)),
+    SD = as.character(round(sd,2))
+  ) %>% select(-mean,-sd) %>%  
+  mutate(
+    Variable = factor(
+      Variable,
+      labels = c(
+          "Gender",
+         "Age (years)",
+         "Control",
+         "Knowledge",
+         "Self-investment",
+         "Past donations in 2 years", 
+         "Psychological ownership",
+         "Self-identity",
+         "Intention" 
+      ),
+      levels = c(
+        "de01",
+        "de02",
+        "ss_control",
+                 "ss_knowledge",
+                 "ss_selfinv",
+                 "bd03", 
+                 "ss_psycown",
+                 "ss_selfid",
+                 "ss_intention"),
+      ordered = TRUE
+    ),
+    x=1,
+  )  %>% arrange(Variable)
+
+tmp[tmp$Variable == 'de01','Mean'] <- ' ' 
+tmp[tmp$Variable == 'de01','SD'] <- ' '
+
+#Plot tehm
+tab1 <-  gt(tmp %>% select(-x)) %>% tab_options(table.font.size = 12,data_row.padding = px(6))
+pt2 + wrap_table(tab1,space="free_y")
+```
+
+```{r}
+ggsave(filename = here("figures/table1asFig.pdf"),units = "cm",width = 18,height = 10)
+ggsave(filename = here("figures/table1asFig.png"),units = "cm",width = 18,height = 10)
 ```
@@
          -ss_intention,
-         -ss_donation)
-
-corrplot(cor(data.fit3,use="pairwise.complete.obs"))
+         -ss_donation,
+         -ss_selfinv)
+
 mcar_test(data.fit3 %>% select(-dttma))
@@
 #afterdata removing sum score donation variable (ss_donation)
+```
+
+```{r}
+corrplot(cor(data.fit3,use="pairwise.complete.obs"))
 ```
@@
   "ss_intention" = "Intention",
-  "ss_donation" = "Blood donation history"
+  "ss_donation" = "Past donation behaviour"
 )
@@
   stand = TRUE,
-  sig=0.001
+  #sig=0.001,
+  stars = "regress"
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
@@
   stand = TRUE,
-  sig=0.001
+  #sig=0.001,
+  stars = "regress"
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))
@@
   stand = TRUE,
-  sig=0.001
+  #sig=0.001,
+  stars = "regress"
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))
@@
   stand = TRUE,
-  sig=0.001
+  #sig=0.001,
+  stars = "regress"
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem05.pdf"))
@@
   "intention" = "Intention",
-  "donation" = "Blood donation history"
+  "donation" = "Past donation behaviour"
 )
@@
   stand = TRUE,
-  sig=0.001
+  #sig=0.001,
+   stars = "regress"
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))
@@
   stand = TRUE,
-  sig=0.001
+  #sig=0.001,
+  stars = "regress"
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem03.pdf"))
@@
   stand = TRUE,
-  sig=0.001
+  stars = "regress"
+  #sig=0.001,
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem04.pdf"))
@@
   stand = TRUE,
-  sig=0.001
+  #sig=0.001
+  stars = "regress"
   ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem05.pdf"))

### 2025-11-12 modelcomparison_code.Rmd
2 decimals to corfig

@@
   lab=TRUE,
-  lab_size = 3,
+  lab_size = 2,
   show.legend = TRUE,
   tl.cex = 10,
-  digits = 1
+  digits = 2
  )

### 2025-11-20 modelcomparison_code.Rmd
header formatting

@@
 
-##data cleaning
+## Data cleaning
 
@@
 as_kable_extra(table1, format = "latex")
-```
-
-## exploratory demographics table
-
-```{r}
-#library(gtsummary)
-#library(dplyr)
-#library(kableExtra)
-
-#edata_clean <- data %>%
-  #filter(de01 == 'man' | de01 == "woman") %>%
-  #rename(
-    #Gender = de01,
-    #Age = de02,
-    #DonationEligibility = bd01,
-    #DonationCount = bd02,
-    #DonationCountLast2years = bd03,
-    #WeeksFromLastDonation = bd04
-  #)
-
-# Create summary table
-#etable1 <- tbl_summary(
-  #data = edata_clean,
-  #include = c(Age, DonationEligibility, WeeksFromLastDonation, DonationCount, DonationCountLast2years),
-  #by = Gender,
-  #statistic = list(
-    #all_continuous() ~ "{mean} ({sd})",   # Use mean (SD) for continuous vars
-    #all_categorical() ~ "{n} ({p}%)"
-  #)
-#) %>%
-  #add_n()  # Adds total N for each group
-
-# Output as LaTeX
-#as_kable_extra(etable1, format = "latex")
-
 ```
@@
 
+## Missing data
+
 ```{r}
@@
 
-# a) Sum score model with robust estimator and listwise deletion
+### a) Sum score model with robust estimator and listwise deletion
 
@@
 
-#  b) Sum score model with bootstrapping and listwise deletion
+###  b) Sum score model with bootstrapping and listwise deletion
 
@@
 
-#  c) Sum score model with robust estimator and fiml
+###  c) Sum score model with robust estimator and fiml
 
@@
 
-#  d) Sum score model with bootstrapping and fiml
+###  d) Sum score model with bootstrapping and fiml
 
@@
 
-# e) Latent factor model with robust estimator and listwise deletion
+### e) Latent factor model with robust estimator and listwise deletion
 
@@
 
-# f) Latent factor model with boostrapping and listwise deletion
+### f) Latent factor model with boostrapping and listwise deletion
 
